---
title: "Codebook metadada"
author: "Manoel"
date: "2025-06-30"
output: pdf_document
---
  
# ---------- OVERALL DESCRIPTION ----------

"This dataset supports the article 'The rank effect of trade on Brazil’s foreign policy political alignment with China'.  It combines annual bilateral-trade statistics, UN General Assembly (UNGA) ideal-point estimates, and auxiliary national indicators to test whether China’s rise to Brazil’s top trading-partner position in 2009 had a detectable causal effect on Brazilian voting behaviour.  The unit of observation is country–year.  Key variables include: (1) China’s share of each country’s total exports and imports; (2) a dummy marking the year a partner becomes a country’s largest trader (the 'rank-reversal' treatment); (3) absolute ideal-point distance between each country and China (Bailey, Strezhnev & Voeten 2017); (4) covariates for GDP, power measure, socioeconomic indicators and geographic variables. Data sources are the Gravity Dataset, Bailey–Strezhnev–Voeten UNGA data, Diplometric and the World Bank’s WDI.  All variable construction steps are documented in the replication pipeline (_targets_)."
  
  
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(labelled)
library(codebook)
```


```{r trade-data, include=FALSE, eval=TRUE}
library(knitr)
library(tibble)

codebook_md <- tribble(
  ~variable,            ~description,                                                        ~source,
  "iso3c",              "Three-letter country code (ISO-3166-1)",                            "Dynamic Gravity, USITC",
  "year",               "Calendar year",                                                    "UNGA ideal-point dataset",
  "pop",                "Population (millions)",                                            "Dynamic Gravity, USITC",
  "abs_distance_china", "Absolute distance to China in UNGA ideal-point space",             "UNGA ideal-point dataset",
  "gdp_cur",            "GDP, current USD",                                                 "Dynamic Gravity, USITC",
  "abs_distance_usa",   "Absolute distance to United States in UNGA ideal-point space",     "UNGA ideal-point dataset",
  "gpi",                "Global Power Index",                                               "Diplometric – Pardee Inst.",
  "us_power_gap",       "U.S. GPI minus country GPI",                                       "Computed from GPI",
  "us_ideal",           "U.S. ideal-point estimate",                                        "UNGA ideal-point dataset",
  "trade_with_china",   "Goods trade with China (USD)",                                     "Dynamic Gravity, USITC",
  "trade_with_us",      "Goods trade with United States (USD)",                             "Dynamic Gravity, USITC",
  "total_trade",        "Total goods trade (USD)",                                          "Dynamic Gravity, USITC",
  "distance_us",        "Distance capital ↔ Washington DC (km)",                            "CEPII GeoDist via DG",
  "distance_china",     "Distance capital ↔ Beijing (km)",                                  "CEPII GeoDist via DG",
  "exchange_rate",      "Market exchange rate LCUs per USD (annual avg.)",                  "World Bank WDI (PA.NUS.FCRF)",
  "latin_america",      "1 = Latin America & Caribbean; 0 otherwise",                       "Author coding (UN regions)",
  "pci_cur",            "GDP per capita, current USD",                                      "Dynamic Gravity, USITC",
  "perc_trade_with_china", "Trade with China / total trade",                                "Author calc.",
  "perc_trade_with_us",   "Trade with U.S. / total trade",                                  "Author calc."
)

knitr::kable(codebook_md, format = "pipe")
```

