---
title: "The Foreign Policy Impact of Trade-Based Status Gains: When China Overtakes the US as Top Export Destination"
author: "Manoel Galdino (University of São Paulo; mgaldino@usp.br)"
date: "`r Sys.Date()`"
output:
  bookdown::pdf_document2:
    number_sections: TRUE
    toc: FALSE
    latex_engine: xelatex

header-includes:
- \usepackage{setspace}\doublespacing
- \setstretch{1.5}
- \usepackage{parskip}
- \setlength{\parskip}{6pt}
- \setlength{\parindent}{0pt}
- \usepackage{float}
- \newcommand{\argmin}{\mathop{\mathrm{argmin}}\limits}
fontsize: 12pt
abstract: "\\singlespacing  How do trade-based status gains affect UNGA voting convergence toward China? I argue that rank reversals in trade hierarchies create politically usable status cues: when China becomes a country's top export destination and remains there, policymakers can more easily justify selective diplomatic adjustment toward Beijing. I test the argument in Brazil, where China overtook the United States as the largest export destination in 2009, using Synthetic Difference-in-Differences on UNGA ideal-point distance. Brazil's ideal-point distance to China fell by 0.26 units, about 41 percent of its pre-treatment mean, and vote-level diagnostics show that the post-2009 movement is clearest in human-rights resolutions where China and the United States diverged. Brazilian newspaper evidence shows increased China-related trade salience after the rank reversal. Cross-country estimates from an interactive fixed-effects panel model show the same directional pattern among countries where China persistently becomes the top export destination. The evidence suggests that discrete and durable trade-status milestones can shape observable multilateral alignment beyond smooth trade growth."
geometry: margin=2.5cm
spacing: double
mainfont: Times New Roman
bibliography: "synth-trade-china.bib"
link-citations: true
---
\pagebreak

```{r setup, include = FALSE}
library(tidyverse)
library(targets)
knitr::opts_chunk$set(echo = FALSE)
source(here::here("scripts", "functions.R"))

caption_note <- function(unit = NULL, se = NULL, cluster = NULL, window = NULL) {
  fields <- c(
    if (!is.null(unit) && nzchar(unit)) unit,
    if (!is.null(se) && nzchar(se)) paste0("SE: ", se),
    if (!is.null(cluster) && nzchar(cluster)) paste0("Cluster: ", cluster),
    if (!is.null(window) && nzchar(window)) paste0("Window: ", window)
  )
  if (length(fields) == 0) {
    return("")
  }
  paste0("Note: ", paste(fields, collapse = ". "), ".")
}

resolve_existing_file <- function(candidates) {
  hits <- candidates[file.exists(candidates)]
  if (length(hits) == 0) {
    stop("Missing file. Tried: ", paste(candidates, collapse = " | "))
  }
  hits[[1]]
}

word_count_text <- function(input = knitr::current_input(), subtract_n = 20L) {
  wc_raw <- system2("wc", c("-w", input), stdout = TRUE)
  wc_n <- as.integer(strsplit(trimws(wc_raw), "\\s+")[[1]][1])
  format(wc_n - subtract_n, scientific = FALSE, big.mark = ",", trim = TRUE)
}
```


Word count: `r word_count_text()`

```{r basic-num, include = FALSE}
estimate <- as.numeric(tar_read(synth_fit)[1])
se_estimate <- as.numeric(tar_read(se_synth))
ci_low <- estimate - 1.96 * se_estimate
ci_high <- estimate + 1.96 * se_estimate
p_estimate <- 2 * pnorm(-abs(estimate / se_estimate))
mean_br <- tar_read(synth_data) %>%
  dplyr::filter(iso3c == "BRA", year < 2009) %>%
  dplyr::summarise(mean_abs = mean(abs_distance_china, na.rm = TRUE)) %>%
  dplyr::pull(mean_abs)
perc_change <- abs(estimate) / mean_br * 100
human_rights_vote_change <- tar_read(brazil_china_unvotes_resolution_2005_2012) %>%
  dplyr::filter(issue_family == "Human rights") %>%
  dplyr::mutate(period = dplyr::if_else(year < 2009, "pre", "post")) %>%
  dplyr::summarise(
    identical_vote_share = 100 * mean(identical_vote, na.rm = TRUE),
    .by = period
  ) %>%
  tidyr::pivot_wider(names_from = period, values_from = identical_vote_share) %>%
  dplyr::mutate(change = post - pre) %>%
  dplyr::pull(change)
sample_size_countries <- tar_read(synth_data) %>%
  dplyr::summarise(num_countries = dplyr::n_distinct(iso3c)) %>%
  dplyr::pull(num_countries)
```



# Introduction

“China overtakes US as Brazil’s top trade partner”. That was a headline in many Brazilian newspapers in 2009 [@rodrigues2009]. The BBC ran a similar headline in 2021, when China surpassed the United States as the EU’s biggest trading partner [@bbc2021]. Sensing the symbolic weight of the milestone in Brazil, President Lula da Silva promptly announced a high-profile trip to Beijing, flanked by an unusually large business delegation, where he mentioned that China was Brazil's largest export destination [@lula2009].

Why do media outlets and political leaders dwell on discrete commercial milestones when the underlying flows are measured continuously? Export statistics are updated monthly, and journalists, diplomats, firms, and policymakers follow them long before one partner crosses a formal threshold. The puzzle, then, is not whether actors possess information about trade volumes. It is why a categorical change—China becomes number one—can matter politically despite adding little new information about the economic trend itself.

Answering this question rigorously is difficult because trade hierarchies are not randomly assigned. Existing work on economic statecraft and foreign-policy alignment usually treats economic exposure as continuous: as trade, lending, investment, aid, or coercive capacity accumulate, alignment is expected to shift through material leverage, changing interests, soft power, or domestic pressure [@flores-macias_kreps2013; @urdinez_etal2016; @kastner_pearson2021]. A continuous-interdependence account would therefore expect foreign-policy alignment to move gradually with trade exposure, leverage, and economic dependence rather than at a discrete rank threshold. Domestic-political accounts, by contrast, would attribute foreign-policy change in cases such as Brazil to leadership ideology, autonomy, South-South cooperation, or broader foreign-policy reorientation [@neto2011; @neto_malamud2015; @rodrigues2019measuring; @schenoniMythsMultipolaritySources2022]. A further selection concern is that politically aligned countries may deepen economic ties with one another, making China’s rise in the trade hierarchy a consequence rather than a cause of prior diplomatic affinity [@gowa1995; @davis_pratt2021; @struver2016]. If alignment reflects only accumulated exposure, domestic-political reorientation, or prior diplomatic affinity, however, no specific symbolic milestone such as becoming a top trade partner should matter.

I argue that such milestones matter because they mark trade-based status gains. The argument builds on research that treats status as socially recognized standing within international hierarchies, clubs, and local status orders [@paul_larson_wohlforth2014; @wolfRespectDisrespect2011; @duque_2018; @renshon2017; @gotz_2021; @macdonald_parent2021; @roren_2024; @roren2025_power_recognition; @wolfTakingInteractionSeriously2019]. By trade-based status, I mean the recognized standing of a foreign power within a country’s hierarchy of economically consequential relationships. Trade volume is continuous, but status is categorical: a partner can move from being important to being understood as central. Once that reclassification becomes public, the relationship is no longer merely larger than before. It becomes easier for political actors to invoke as a national priority, to defend as strategically consequential, and to use as a justification for diplomatic adjustment.

Much of the China and rising-powers branch of this literature asks how rising states seek recognition, major-power status, or institutional standing, and how misrecognition can produce contestation or revisionism [@deng2008_status; @larson_shevchenko2010_status_seekers; @ward2017_status_challenge; @murray2018_recognition; @mukherjee2022_ascending]. This argument relocates status from the ambitions of the rising power to the recognition practices of states. China’s trade-based status gain occurs when another country publicly recategorizes China within its own hierarchy of external economic relationships -- from an important partner to a central one. The mechanism is therefore status-theoretic, with Brazil as the audience and status-conferring actor.

Brazil is the strongest case for the theory. China displaced the United States, a politically salient incumbent and hegemonic benchmark; the rank reversal was publicly visible; and the Brazilian press allows the attention implication to be measured. The case is also substantively important because Lula-era Brazilian foreign policy is often described through autonomy, South-South activism, multipolarity, and Brazil's triangular position between the United States and China [@neto2011; @neto_malamud2015; @rodrigues2019measuring; @schenoniMythsMultipolaritySources2022; @guilhon2014; @schenoni_leiva2021; @brun_covarrubias2024; @coutinho_rodriguez2024]. 

To provide a causal credible estimate, I leverage the discreteness in China's status change to largest Brazilian export destination to estimate its effect in the Brazilian foreign policy alignment using synthetic difference-in-differences (SDiD) on annual UNGA ideal-point distance to China. The estimate is an average post-2009 gap. Relative to Brazil's pre-treatment mean distance, the estimated reduction is `r sprintf('%.1f', perc_change)` percent. UNGA voting is a standard measure to study alignment with China and other powers, linking votes to bilateral ties, ideology, sanctions, aid, Belt and Road membership, arms transfers, or Chinese economic returns [@potrafke2009; @dreher_jensen2013; @struver2016; @yan_zhou2021; @garcia_herrero_etal2024; @steinert_weyrauch2024; @he_zheng_chen2025; @lektzian_biglaiser2023].

To assess whether the pattern is specific to Brazil, I extend the analysis to a cross-country panel in which China enters and holds the top export-destination position to provide evidence that the rank-threshold pattern travels beyond the theory-generating case. The cross-country estimates are smaller in magnitude, consistent with a more heterogeneous sample, and point in the same direction as the Brazilian case. The panel evidence therefore strengthens the core claim that rank thresholds matter beyond the theory-generating case, while the Brazilian evidence shows the mechanism in detail.

The paper’s contribution is to recast economic influence as a status process. Existing work usually treats trade exposure as a continuous source of leverage, interests, or domestic pressure. I show that economic ties matter through accumulated exposure and through moments when they turn a foreign power into a country’s leading trade partner. In Brazil, China’s rise to the top export-destination position made an already expanding commercial relationship politically usable as evidence of China’s new centrality. That status gain helps explain why convergence appears selectively in UNGA voting. More broadly, the paper shows that status gains can emerge from economic hierarchies and shape observable diplomatic behavior.


The next section develops the status-threshold argument and derives observable implications. I then present the Brazil SDiD design and timing diagnostics, followed by issue-area and media evidence on the mechanism. The final empirical section evaluates whether persistent China-top export status is associated with similar movement in a broader cross-country panel.

# Theory and hypotheses

A trade-based status shift differs from smooth growth in economic interdependence because it changes the category through which a foreign power is understood. Actors can observe trade shares, but political attention rarely follows continuous movements one percentage point at a time. It also organizes around coarse public labels: largest export market, number-one partner, top-ranked export destination. These labels simplify complex economic relationships and help journalists, politicians, organized interests, and diplomats coordinate attention [@bordalo_etal2022; @conlon2023; @enke2024; @graeber_etal2025]. The number-one threshold matters because it can reclassify a rising power from one important commercial partner among several into the country's central economic partner. In this sense, export-destination rank operationalizes the trade-based status shift.

The cue should matter most when the new status also displaces a politically prominent incumbent. If a rising major power overtakes a minor partner, the event may remain a commercial fact with limited political resonance. If it overtakes the hegemonic incumbent, the same threshold takes on broader political meaning. It can be read as a visible loss of standing for the incumbent and as a public marker of the challenger's rise. Brazil falls squarely inside this high-salience condition: China became Brazil's largest export destination in 2009, displaced the United States, and contemporary coverage frequently described the event in broader top-partner language.

The mechanism has two steps. First, the status cue should increase attention. Media coverage should make the challenger and the bilateral trade relationship more visible after the threshold is crossed. Second, the cue should affect policy through justification. Once that power is publicly understood as a central economic partner, executives and diplomats can frame limited movement toward it as pragmatic recognition of the country's economic hierarchy rather than as ideological concession or opportunistic bandwagoning. This step can be gradual because UNGA positions are embedded in prior votes, bureaucratic routines, coalition commitments, and reputational constraints. The Brazilian SDiD design estimates the combined reduced-form effect of this status cue on UNGA distance to China.

The justification mechanism also implies a durability condition. The rank reversal is not expected to reshape foreign policy through a single decision at the moment China becomes number one. Rather, it creates a public category that officials can invoke repeatedly across a sequence of selective adjustments. When China remains the leading export destination, presidents, ministers, diplomats, and organized interests can treat its centrality as a stable fact of the national economic hierarchy and use it to justify incremental diplomatic movement. A one-year China-top episode may attract attention, but it is a weaker treatment regime: it does not provide a durable basis for repeated public justification, bureaucratic recalibration, or cumulative foreign-policy adaptation.

## From Attention to Selective Diplomatic Alignment

Trade-based status changes the terms on which selective accommodation can be justified. Some foreign-policy arenas are especially suitable for this kind of adjustment: they are visible to foreign governments and diplomatic audiences, but less costly domestically than treaties, military commitments, sanctions, or trade policy. In such arenas, governments can signal movement toward a newly central partner while preserving flexibility.

When a foreign power becomes publicly recognized as the country’s leading trade partner, policymakers can frame limited convergence with that power as pragmatic recognition of economic importance rather than as ideological concession or opportunistic bandwagoning. The status gain can therefore lower the reputational cost of selected movement while increasing its diplomatic value.

The domestic pathway works through justification. Media coverage need not cause individual foreign-policy positions, and business groups need not lobby over specific diplomatic decisions. Public recognition of the partner’s new standing changes what diplomats and executives can plausibly defend. Adjustment toward that partner becomes easier to present as alignment with national economic interests, especially in diplomatic settings where positions are visible internationally but relatively low-cost domestically.

This logic implies selective convergence. The mechanism should be most informative in issue areas where the country and the newly elevated partner previously differed and where movement is not costless. In domains tied to prior reputational commitments, adjustment toward the newly central partner requires a stronger rationale than in areas where the two countries already agree. Trade-based status supplies such a rationale by making the partner’s economic centrality publicly available as a justification for diplomatic accommodation.

The main theoretical expectation is therefore that when a rising major power becomes a country’s leading trade partner, the country's foreign policy should move closer to that power relative to its counterfactual path. The effect should be easiest to observe when the status gain displaces a politically prominent incumbent, especially a hegemonic benchmark such as the United States. The theory also implies an attention pattern: media coverage of the rising power and of the bilateral economic relationship should increase around the status shift.

# Data and design

Empirically, the treatment is China becoming a country's largest export destination. I use "top trade partner" only as shorthand for the publicly salient Brazilian event when discussing contemporary framing. The Brazil SDiD and cross-country panel define treatment using largest-export-destination status.

```{r basic-data, message=FALSE, warning=FALSE, echo=FALSE}
library(targets)
library(tidyverse)
df <- tar_read(synth_data)
num_countries <- df %>%
  dplyr::summarise(num_countries = dplyr::n_distinct(iso3c)) %>%
  dplyr::pull(num_countries)

num_years <- df %>%
  dplyr::filter(iso3c == "BRA", year < 2016) %>%
  dplyr::summarise(num_years = dplyr::n()) %>%
  dplyr::pull(num_years)

min_year <- df %>%
  dplyr::summarise(min_year = min(year)) %>%
  dplyr::pull(min_year)

max_year <- df %>%
  dplyr::filter(year < 2016) %>%
  dplyr::summarise(max_year = max(year)) %>%
  dplyr::pull(max_year)
```

The primary outcome is absolute UNGA ideal-point distance to China, based on the annual ideal-point estimates of @bailey_etal2017. Lower values indicate convergence toward China. This measure is preferable to raw annual agreement shares for the main design because it is less sensitive to changes in the UNGA agenda over time. I use resolution-level agreement diagnostics only to interpret where convergence appears strongest.

For Brazil, the treatment indicator equals one from 2009 onward, the first year China became Brazil's largest export destination. SDiD estimates Brazil's average post-2009 gap relative to a weighted synthetic counterfactual, using pre-treatment information to construct unit and time weights [@arkhangelsky_etal2021]. The design is reduced-form. Its credibility rests on pre-treatment fit, rank-versus-volume timing tests, donor-pool sensitivity, and outcome robustness.

The SDiD panel combines UNGA ideal points with covariates for trade exposure, macroeconomic conditions, power, geographic distance, political institutions, and trade-agreement ties. The complete panel covers `r num_countries` countries and `r num_years` Brazil years for the main SDiD window (`r min_year`-`r max_year`). Trade data come from the International Trade and Production Database for Estimation [@borchert_etal2021]. The ITPD-E `trade` variable is measured in millions of current US dollars, and export-destination ranks are computed from exporter-importer flows.

The Brazil variables enter the design in three groups. The outcome is Brazil's annual absolute UNGA ideal-point distance to China; the treatment is an indicator for years after China becomes Brazil's largest export destination. The covariates measure trade exposure to China and the United States, relative power, geographic distance to Washington, leader ideology, macroeconomic stress, regime institutions, and U.S. trade-agreement ties. The cross-country panel uses the same outcome and treatment logic. Table \@ref(tab:table-summary-main) summarizes the country-year panel used to estimate Brazil's synthetic counterfactual. The large number of observations reflects the donor pool and pre/post-treatment country-year structure of the SDiD design.

```{r table-summary-main, message=FALSE, warning=FALSE, echo=FALSE, results='asis'}
summary_vars <- tar_read(synth_data) %>%
  dplyr::select(
    `Treatment` = treatment,
    `UNGA distance to China` = abs_distance_china,
    `UNGA distance to US` = abs_distance_usa,
    `Geographic distance to US` = distance_us,
    `Trade share with China` = perc_trade_with_china,
    `Trade share with US` = perc_trade_with_us,
    `Power index` = gpi,
    `Power gap to US` = us_power_gap,
    `GDP per capita` = pci_cur,
    `Exchange rate` = exachange_rate,
    `HoG left ideology` = hog_left,
    `Current account` = CA_GDP,
    `Government deficit` = govdef_GDP,
    `Parliamentary system` = dplyr::any_of("inst_parliamentary"),
    `Military executive` = dplyr::any_of("inst_military_exec"),
    `US trade agreement` = dplyr::any_of("us_trade_agreement")
  )

summary_tbl <- purrr::map_dfr(names(summary_vars), function(var_name) {
  x <- summary_vars[[var_name]]
  tibble::tibble(
    Variable = var_name,
    N = sum(!is.na(x)),
    Mean = mean(x, na.rm = TRUE),
    SD = stats::sd(x, na.rm = TRUE),
    Min = min(x, na.rm = TRUE),
    Max = max(x, na.rm = TRUE)
  )
}) %>%
  dplyr::mutate(
    dplyr::across(c(Mean, SD, Min, Max), ~sprintf("%.2f", .x))
  )

knitr::kable(
  summary_tbl,
  format = "latex",
  booktabs = TRUE,
  caption = "Descriptive statistics for the country-year panel used in the Brazil SDiD design.",
  align = "lccccc"
) %>%
  kableExtra::kable_styling(
    latex_options = c("HOLD_position", "scale_down"),
    font_size = 8
  ) %>%
  kableExtra::footnote(
    general = paste0(
      sub(
        "^Note: ",
        "",
        caption_note(
          unit = "Country-year variables from the SDiD analytic panel; continuous covariates are rescaled in the estimation data, UNGA distances are ideal-point distances, and treatment is binary",
          window = sprintf("%d-%d SDiD analytic panel", min_year, max_year)
        )
      ),
      " Observations include Brazil and donor countries used to construct the synthetic counterfactual. The treatment mean is small because only Brazil is treated after 2009."
    ),
    general_title = "Note: ",
    escape = FALSE,
    threeparttable = TRUE
  )
```

For the cross-country evidence beyond Brazil, the main sample is restricted to persistent China-top export-destination transitions. Because the theory concerns status consolidation, the cross-country analysis defines treatment as persistent China-top export-destination status. Treated countries are those where China first becomes the largest export destination in or after 2000, after an observed prior year in which China was not number one, and then remains in that position through the end of the observed panel. Countries where China never becomes the largest export destination remain controls. Countries where China enters the top position only temporarily are excluded from the main cross-country sample and reported as appendix robustness. This operationalization matches the salience logic more closely than a one-year entry definition: the theory expects political visibility, bureaucratic learning, and diplomatic adaptation to be clearer when the new trade hierarchy persists. For this panel, I use the `fect` counterfactual estimator with interactive fixed effects, which allows countries to load differently on common latent shocks [@liu2024practical]. I refer to this as the IFE specification below. I report the no-covariate IFE specification as the main cross-country estimate and use log GDP per capita plus V-Dem freedom of expression as a covariate-adjusted robustness check. Callaway-Sant'Anna (C\&S) absorbing-treatment estimates are reported as a convergent check on the same persistent-treatment sample.

## Identification strategy

The empirical problem is that China's rise in a country's trade hierarchy is not randomly assigned. Countries may move closer to China because trade exposure accumulates continuously, because domestic political coalitions or foreign-policy doctrines change, or because governments already inclined toward China deepen commercial ties before the rank reversal occurs. The design therefore asks the identifying question: whether foreign-policy alignment changes around the discrete status threshold after accounting for pre-treatment trajectories and observed sources of economic, geopolitical, and domestic-political orientation.

In the Brazil design, SDiD constructs a weighted counterfactual from pre-treatment information. The covariates are not included simply to improve fit. They map directly onto the rival explanations. Trade shares with China and the United States absorb smooth changes in economic exposure; relative power, geographic distance to Washington, and U.S. trade-agreement ties capture structural exposure to the incumbent power; leader ideology, regime institutions, and macroeconomic stress account for domestic-political reorientation and contemporaneous shocks. The demand-shock diagnostics further add smooth China-share transformations, export-destination concentration, pre-2009 primary-goods export exposure, and 2008-2009 exposure interactions. The rank-versus-volume timing tests then ask whether earlier trade growth or lower-rank promotions generate the same estimated convergence as the top-rank reversal.

The cross-country panel provides a second check against Brazil-specific timing explanations. Treated countries enter the persistent China-top condition in different years, so a Brazil-specific 2009 shock cannot mechanically generate the pooled result. Interactive fixed effects allow countries to load differently on common latent shocks, while covariate-adjusted specifications probe whether the estimated pattern survives adjustment for development and public-salience conditions.

# Brazil SDiD evidence

The Brazilian analysis estimates the average post-2009 reduced-form effect of China's entry into the top export-destination position on Brazil's UNGA ideal-point distance to China. It asks whether Brazil moved closer to China than a synthetic counterfactual after the 2009 rank reversal.

Figure \@ref(fig:brazil-rank-volume) shows the timing of the rank threshold. China's export share rose before 2009, but China became the top-ranked export destination only in 2009. The signed margin over the relevant competitor also changes sign at that threshold. The SDiD covariate matrix includes trade shares with China and the United States to account for smooth trade exposure. The rank margin is part of the threshold event itself, so it is used as a timing diagnostic rather than as a separate control.

```{r brazil-rank-volume, message=FALSE, warning=FALSE, echo=FALSE, fig.cap=paste0("Brazil rank-versus-volume diagnostic for China's export-destination status. Panels show China's export share, ordinal rank, and signed margin over the relevant competitor. The dashed line marks the 2009 rank reversal. ", caption_note(unit = "Export share in percent; ordinal rank; export-value margin in current USD billions", window = "2000-2012 annual Brazil series")), fig.pos="H", fig.width=7.2, fig.height=7.2}
tar_read(goal3_brazil_rank_volume_plot)
```

Figure \@ref(fig:plot-sdid) presents the main SDiD fit. The pre-treatment fit is close enough to make the post-treatment comparison informative, and the post-2009 gap is negative, indicating convergence toward China. The point estimate is `r sprintf('%1.2f', estimate)` ideal-point units, with placebo-based SE `r sprintf('%1.2f', se_estimate)` and p-value `r sprintf('%.3f', p_estimate)`. Relative to Brazil's pre-treatment mean distance (`r sprintf('%.3f', mean_br)`), this corresponds to a `r sprintf('%.1f', perc_change)` percent reduction. The estimate summarizes the average post-treatment gap. Appendix Figure \@ref(fig:appendix-brazil-china-ideal-points) plots the underlying Brazil and China ideal-point series separately; China is comparatively stable, so the post-2009 reduction in absolute distance is driven mainly by movement in Brazil's position rather than by China moving toward Brazil.

```{r plot-sdid, message=FALSE, warning=FALSE, echo=FALSE, fig.cap=paste0("SDiD fit for Brazil and synthetic control with time weights. Lower values indicate convergence toward China; the post-treatment gap is an average post-2009 reduced-form effect. ", caption_note(unit = "Absolute UNGA ideal-point distance to China", se = "placebo-based SE with 1,000 replications; p-values reported in the table", window = "Brazil annual series")), fig.pos="H"}
tar_read(plot_trend)
```

Table \@ref(tab:outcome-robustness-table) reports the preferred Brazil SDiD estimate alongside three specification checks. Each column estimates the average post-2009 gap in Brazil's UNGA ideal-point distance to China and reports a placebo-based standard error. The baseline without institutional covariates, the specification without time-varying covariates, and the Latin America donor pool preserve the negative sign. Column (3) addresses post-treatment-adjustment concerns directly by holding only pre-treatment or time-invariant covariates fixed across panel years; the estimate remains nearly identical to the main specification. Column (1) remains the main specification because it keeps the theoretically relevant donor pool while adjusting for trade exposure, macroeconomic stress, geographic distance to Washington, and institutional covariates.

```{r outcome-robustness-table, message=FALSE, warning=FALSE, echo=FALSE, results='asis'}
sdid_spec_table <- tar_read(brazil_sdid_spec_table)
sdid_spec_table$table[[1]] <- ifelse(
  sdid_spec_table$table[[1]] == "Distance to US",
  "Geographic distance to US",
  sdid_spec_table$table[[1]]
)

knitr::kable(
  sdid_spec_table$table,
  format = "latex",
  booktabs = TRUE,
  escape = FALSE,
  caption = "Brazil SDiD estimates across main and robustness specifications.",
  align = "lcccc"
) %>%
  kableExtra::kable_styling(
    latex_options = c("HOLD_position", "scale_down"),
    font_size = 8
  ) %>%
  kableExtra::footnote(
    general = paste0(
      "ATT is absolute UNGA ideal-point distance to China; negative values mean convergence. ",
      "Placebo-based SEs with 1,000 replications are in parentheses. Checkmarks show included covariate groups. ",
      "Donor counts exclude Brazil. Column (3) uses pre-treatment or time-invariant covariates held fixed across panel years."
    ),
    general_title = "Note: ",
    escape = FALSE,
    threeparttable = TRUE
  )
```

```{r china-demand-sdid-numbers, include=FALSE}
china_demand_sdid_results <- tryCatch(
  tar_read(china_demand_sdid_diagnostics_table),
  error = function(e) {
    readr::read_csv(
      here::here(
        "quality_reports",
        "china_demand_shock_rank_threshold",
        "table_3_brazil_sdid_expanded_controls.csv"
      ),
      show_col_types = FALSE
    ) %>%
      dplyr::mutate(
        label = dplyr::case_when(
          specification == "manuscript_main" ~ "Manuscript main",
          specification == "plus_smooth_china_share" ~ "Smooth share",
          specification == "plus_destination_concentration" ~ "Concentration",
          specification == "plus_rank_margin_bad_control" ~ "Rank margin",
          specification == "plus_commodity_gfc_exposure" ~ "Commodity/GFC exposure",
          specification == "full_diagnostic" ~ "Full stress test",
          TRUE ~ specification
        ),
        control_family = covariates_added
      )
  }
)
china_demand_main <- china_demand_sdid_results %>%
  dplyr::filter(specification == "manuscript_main") %>%
  dplyr::slice_head(n = 1)
china_demand_commodity <- china_demand_sdid_results %>%
  dplyr::filter(specification %in% c("commodity_gfc_exposure", "plus_commodity_gfc_exposure")) %>%
  dplyr::slice_head(n = 1)
china_demand_full <- china_demand_sdid_results %>%
  dplyr::filter(specification %in% c("full_stress_test", "full_diagnostic")) %>%
  dplyr::slice_head(n = 1)
```

The rank interpretation requires more than showing that trade with China was increasing. Brazil's exports to China grew before 2009, including years in which China rose in the export hierarchy but did not become number one. The timing tests in Table \@ref(tab:table-placebo) therefore ask whether rapid trade growth and lower-rank promotions generate the same estimated UNGA convergence as the top-rank reversal. They do not. The 2003, 2004, and 2005 rows are timing and lower-threshold falsifications for the continuous-growth alternative: 2004 is especially useful because China first reaches rank #2 but is still not Brazil's largest export destination. The 2012 row probes whether the result is driven by a later break after China had already become the top destination.

The same timing logic addresses the main Brazilian political confounder. President Lula da Silva took office in 2003 and pursued a more explicit South-South foreign-policy agenda, a shift emphasized in work on Brazilian autonomy, presidential foreign policy, and international engagement [@neto2011; @mouron_urdinez2014; @neto_malamud2015; @rodrigues2019measuring; @schenoniMythsMultipolaritySources2022]. If that ideological shift were sufficient to explain Brazil-China UNGA convergence, the pseudo-treatment years in 2003 or 2005 should already show comparable movement. They do not. The cross-country panel also estimates the rank-threshold pattern across countries with different political systems and leadership profiles, which reduces the concern that the Brazil estimate is only a Lula-period artifact.

A remaining concern is that 2009 bundled the Brazilian rank reversal with other contemporaneous shocks, including the global financial crisis, the BRICS/G20 moment, and the commodity cycle. Table \@ref(tab:china-demand-sdid-diagnostics) addresses this critique inside the same Brazil SDiD framework. The main estimate is `r sprintf("%.3f", china_demand_main$estimate)`; after adding pre-2009 primary-goods export exposure and 2008-2009 exposure interactions it is `r sprintf("%.3f", china_demand_commodity$estimate)`, and the full stress test returns `r sprintf("%.3f", china_demand_full$estimate)`. The sign and magnitude therefore do not collapse when the categorical China #1 onset is estimated alongside continuous China-share growth, export concentration, commodity exposure, rank-margin stress tests, and global-crisis interactions. This does not prove that the rank cue is experimentally isolated from every contemporaneous political shock. It does show that the result is not an artifact of omitting smooth China-demand exposure from the SDiD design.

The cross-country panel below is a broader scope check, not the main answer to the Brazil-specific demand-shock critique. Brazil is included in the absorbing cross-country sample, but the panel aligns countries by their own persistent China-top entry years, which range from 2003 to 2021 in the main specification. A Brazil-specific 2009 confounder cannot mechanically generate an average event-time effect across countries treated in different calendar years. For such an explanation to account for both the Brazil estimate and the cross-country result, the omitted force would need to coincide with persistent China-top entry across countries, affect UNGA distance to China in the same direction, and remain outside the observed covariates and the latent factors captured by the interactive fixed-effects specification. In sensitivity-analysis terms, it would need enough residual association with both treatment timing and the outcome to overturn estimates that are stable across timing tests, donor-pool checks, estimators, and covariate choices [@cinelli_hazlett2020; @oster2019].

```{r table-placebo, message=FALSE, warning=FALSE, echo=FALSE, results='asis'}
rank_tests <- tar_read(goal3_brazil_placebo_rank_volume_tests) %>%
  dplyr::mutate(
    `Timing year` = nominal_treatment_year,
    `Test role` = dplyr::case_when(
      nominal_treatment_year == 2003L ~ "Growth / lower-rank promotion",
      nominal_treatment_year == 2004L ~ "China #2 threshold",
      nominal_treatment_year == 2005L ~ "Rapid growth without rank 1",
      nominal_treatment_year == 2009L ~ "Actual rank-1 reversal",
      nominal_treatment_year == 2012L ~ "Later-break falsification",
      TRUE ~ timing_test
    ),
    `China rank` = china_rank,
    `Rank-1 reversal` = ifelse(rank1_reversal, "Yes", "No"),
    `Export share (\\%)` = sprintf("%.1f", china_share_pct),
    `Margin (USD bn)` = sprintf("%.1f", china_margin_vs_competitor_usd_billion),
    ATT = sprintf("%.3f", estimate),
    `Placebo SE` = sprintf("%.3f", se_placebo),
    `p-value` = sprintf("%.3f", p_value)
  ) %>%
  dplyr::select(
    `Timing year`, `Test role`, `China rank`, `Rank-1 reversal`,
    `Export share (\\%)`, `Margin (USD bn)`, ATT, `Placebo SE`, `p-value`
  )
rank_tests_note <- paste0(
  sub(
    "^Note: ",
    "",
    caption_note(
      unit = "ATT is absolute UNGA ideal-point distance to China; negative values mean convergence; export margins are current USD billions",
      se = "normal approximation using placebo-based SE with 1,000 replications",
      window = "2003-2005 timing tests truncate the sample before 2009; 2012 is a later-break falsification"
    )
  ),
  " The 2003, 2004, and 2005 rows are timing and lower-threshold falsifications for rapid export growth without rank-1 status."
)

knitr::kable(
  rank_tests,
  format = "latex",
  booktabs = TRUE,
  escape = FALSE,
  caption = "Brazil SDiD estimate and rank-versus-volume timing tests.",
  align = "clccccccc"
) %>%
  kableExtra::kable_styling(
    latex_options = c("HOLD_position", "scale_down"),
    font_size = 8
  ) %>%
  kableExtra::footnote(
    general = rank_tests_note,
    general_title = "Note: ",
    escape = FALSE,
    threeparttable = TRUE
  )
```

```{r china-demand-sdid-diagnostics, message=FALSE, warning=FALSE, echo=FALSE, results='asis'}
china_demand_table <- china_demand_sdid_results %>%
  dplyr::mutate(
    Specification = label,
    ATT = sprintf("%.3f", estimate),
    `Placebo SE` = sprintf("%.3f", se_placebo),
    `p-value` = sprintf("%.3f", p_value),
    `95\\% CI` = paste0("[", sprintf("%.3f", ci_95_low), ", ", sprintf("%.3f", ci_95_high), "]"),
    `Diagnostic added` = control_family
  ) %>%
  dplyr::select(
    Specification,
    `Diagnostic added`,
    ATT,
    `Placebo SE`,
    `95\\% CI`,
    `p-value`
  )

china_demand_note <- paste0(
  sub(
    "^Note: ",
    "",
    caption_note(
      unit = "ATT is absolute UNGA ideal-point distance to China; negative values mean convergence",
      se = "synthdid placebo SE with 1,000 replications",
      window = "1997-2015 annual SDiD panel"
    )
  ),
  " All rows keep the binary 2009 China number-one treatment onset. ",
  "Commodity/GFC exposure uses pre-2009 primary-goods export composition from ITPD-E Agriculture plus Mining and Energy over goods exports, interacted with the 2008-2009 shock window. ",
  "The full stress test includes the rank margin, which is mechanically related to the threshold and should be read as a conservative diagnostic rather than a preferred control."
)

knitr::kable(
  china_demand_table,
  format = "latex",
  booktabs = TRUE,
  escape = FALSE,
  caption = "Brazil SDiD diagnostics for the China-demand-shock critique.",
  align = "llcccc"
) %>%
  kableExtra::kable_styling(
    latex_options = c("HOLD_position", "scale_down"),
    font_size = 8
  ) %>%
  kableExtra::footnote(
    general = china_demand_note,
    general_title = "Note: ",
    escape = FALSE,
    threeparttable = TRUE
  )
```

## Issue-Area Evidence: Where Did Convergence Occur?

```{r selective-unga-diagnostic-numbers, include=FALSE}
selective_vote_models <- tar_read(selective_china_alignment_vote_level_models)
selective_ddd_models <- tar_read(selective_china_alignment_ddd_hr_nonhr_models)
selective_country_placebo <- tar_read(selective_china_alignment_country_placebo_summary)

hr_china_us_diff <- selective_vote_models %>%
  dplyr::filter(
    sample == "Human-rights China-US divergent votes",
    outcome == "distance_china_minus_usa"
  ) %>%
  dplyr::slice_head(n = 1)

hr_china_us_strong_diff <- selective_vote_models %>%
  dplyr::filter(
    sample == "Human-rights strong yes/no China-US divergent votes",
    outcome == "distance_china_minus_usa"
  ) %>%
  dplyr::slice_head(n = 1)

nonhr_china_us_diff <- selective_vote_models %>%
  dplyr::filter(
    sample == "Non-human-rights China-US divergent votes",
    outcome == "distance_china_minus_usa"
  ) %>%
  dplyr::slice_head(n = 1)

ddd_hr_increment <- selective_ddd_models %>%
  dplyr::filter(
    outcome == "distance_china_minus_usa",
    term == "brazil_post_hr",
    stringr::str_detect(model, "country-clustered")
  ) %>%
  dplyr::slice_head(n = 1)

distance_placebo <- selective_country_placebo %>%
  dplyr::filter(outcome == "distance_china_minus_usa") %>%
  dplyr::slice_head(n = 1)
```

The aggregate ideal-point estimate raises a substantive question: in what kinds of votes did Brazil move closer to China? Resolution-level diagnostics suggest that the post-2009 convergence was selective rather than uniform. In several issue areas Brazil and China already voted together at high rates before treatment, leaving little room for additional convergence. The more informative domain is human rights, where pre-treatment agreement was lower and where movement toward China carried greater reputational costs for Brazil, given the domestic and international implications of Brazil's human-rights foreign policy [@milani2015humanrights]. In that domain, Brazil-China identical voting increased by `r sprintf("%.1f", human_rights_vote_change)` percentage points after 2009. This pattern is consistent with the proposed mechanism: the trade-rank cue made China's relationship politically usable, allowing policymakers to justify selective convergence in a visible multilateral arena where Brazil had previously maintained more distance.

This descriptive increase does not by itself establish China-specific alignment, because annual issue-area shares can reflect agenda composition and do not compare Brazil with donor countries. I therefore add a more restrictive vote-level diagnostic. In country-resolution regressions with country and resolution fixed effects, restricted to human-rights votes where China and the United States cast different votes, Brazil moves closer to China after 2009 relative to the donor pool. The Brazil-by-post-2009 coefficient is `r sprintf("%.3f", hr_china_us_diff$estimate)` for distance to China minus distance to the United States (model-based SE = `r sprintf("%.3f", hr_china_us_diff$se)`), and `r sprintf("%.3f", hr_china_us_strong_diff$estimate)` when the sample is restricted to strong yes/no China-US disagreements (model-based SE = `r sprintf("%.3f", hr_china_us_strong_diff$se)`). The corresponding non-human-rights estimate is smaller (`r sprintf("%.3f", nonhr_china_us_diff$estimate)`, model-based SE = `r sprintf("%.3f", nonhr_china_us_diff$se)`), and a DDD specification estimates an additional human-rights increment of `r sprintf("%.3f", ddd_hr_increment$estimate)` (model-based SE = `r sprintf("%.3f", ddd_hr_increment$se)`). Because Brazil is the single treated country in this diagnostic, I use country placebos as the main inferential benchmark: Brazil ranks `r sprintf("%.0f", distance_placebo$brazil_rank_expected_direction)` out of `r sprintf("%.0f", distance_placebo$n_placebo_units)` units in the China-specific direction (directional p = `r sprintf("%.3f", distance_placebo$randomization_p_directional)`; strict donor-only p = `r sprintf("%.3f", distance_placebo$randomization_p_directional_strict_donor)`). These diagnostics support a narrower interpretation: the post-2009 movement is strongest in human-rights votes where China and the United States diverged, rather than uniformly across the UNGA agenda.

```{r issue-area-prepost-table, message=FALSE, warning=FALSE, echo=FALSE, results='asis'}
issue_order <- c(
  "Human rights",
  "Arms/disarmament/nuclear",
  "Economic development",
  "Decolonization",
  "Palestine/Middle East",
  "Other / uncoded"
)

issue_prepost_tbl <- tar_read(brazil_china_unvotes_resolution_2005_2012) %>%
  dplyr::mutate(period = dplyr::if_else(year < 2009, "pre", "post")) %>%
  dplyr::summarise(
    n_resolutions = dplyr::n_distinct(rcid),
    identical_vote_share = 100 * mean(identical_vote, na.rm = TRUE),
    .by = c(issue_family, period)
  ) %>%
  tidyr::pivot_wider(
    names_from = period,
    values_from = c(n_resolutions, identical_vote_share)
  ) %>%
  dplyr::mutate(
    change = identical_vote_share_post - identical_vote_share_pre,
    interpretation = dplyr::case_when(
      issue_family == "Human rights" ~ "Politically costly convergence; mechanism-informative",
      issue_family == "Palestine/Middle East" ~ "High baseline agreement; limited room to move",
      issue_family == "Economic development" ~ "Diagnostic evidence",
      issue_family == "Arms/disarmament/nuclear" ~ "Diagnostic evidence",
      issue_family == "Decolonization" ~ "Diagnostic evidence",
      TRUE ~ "Heterogeneous residual category"
    ),
    issue_order_id = match(issue_family, issue_order)
  ) %>%
  dplyr::arrange(is.na(issue_order_id), issue_order_id, issue_family) %>%
  dplyr::transmute(
    `Issue area` = issue_family,
    `Pre-2009 identical vote share` = sprintf("%.1f", identical_vote_share_pre),
    `Post-2009 identical vote share` = sprintf("%.1f", identical_vote_share_post),
    `Change` = sprintf("%.1f", change),
    `Interpretation` = interpretation
  )

knitr::kable(
  issue_prepost_tbl,
  format = "latex",
  booktabs = TRUE,
  escape = FALSE,
  caption = "Brazil-China identical UNGA votes by issue area, before and after 2009.",
  align = "lrrrl"
) %>%
  kableExtra::kable_styling(
    latex_options = c("HOLD_position", "scale_down"),
    font_size = 8
  ) %>%
  kableExtra::footnote(
    general = sub(
      "^Note: ",
      "",
      caption_note(
        unit = "Percentage of roll-call resolutions with identical Brazil-China votes",
        window = "2005-2008 versus 2009-2012; multi-coded resolutions enter each issue family"
      )
    ),
    general_title = "Note: ",
    escape = FALSE,
    threeparttable = TRUE
  )
```

Figure \@ref(fig:plot-brazil-china-unvotes-issue-year) disaggregates the same diagnostic by year and substantive issue family. Because several areas already had high Brazil-China vote similarity before 2009, the figure describes where additional convergence was possible.

```{r plot-brazil-china-unvotes-issue-year, message=FALSE, warning=FALSE, echo=FALSE, fig.cap=paste0("Brazil-China identical UNGA votes by issue area, 2005-2012. Points report annual shares; the vertical line marks 2009. The clearest pattern expected by the mechanism is selective convergence in human rights. ", caption_note(unit = "Percentage of roll-call resolutions with identical Brazil-China votes", window = "2005-2012 annual series; multi-coded resolutions enter each issue family")), fig.pos="H", fig.width=11, fig.height=8}
tar_read(plot_brazil_china_unvotes_similarity_by_issue_year_2005_2012)
```

This issue-area evidence is a diagnostic of substantive content, not a standalone causal test. The SDiD estimate identifies the average post-2009 gap in Brazil's ideal-point distance to China; the vote-level evidence clarifies where that movement is most visible. The strongest support comes from human-rights resolutions in which China and the United States diverged, where Brazil moves closer to China relative to comparable donor countries after 2009. A DDD comparison shows that the human-rights shift is larger than the analogous non-human-rights shift. Country placebos counsel against language implying that Brazil is unique. The Brazil evidence therefore supports a cautious claim of selective UNGA adjustment toward China in politically visible domains, rather than a claim that the issue-area table alone proves selective convergence.

# Brazilian media salience

To assess whether the rank reversal became publicly salient, I examine coverage of China in Folha de S.Paulo, Brazil's widest-circulation newspaper [@folha2016], from 2000 to 2014. Salience matters because the rank-threshold argument requires the top-rank cue to become politically usable. A change in export hierarchy can shape elite and public attention when it enters public discourse as a simple signal of China's commercial importance, consistent with work on media, agenda setting, and foreign policy [@soroka2003; @baum_potter2008; @walgraveContingencyMassMedias2006].

Headlines are the appropriate unit for this diagnostic because the mechanism is public attention and categorical labeling. This follows corpus-assisted media studies that use headlines because they are tractable in large archives and operate as attention-directing cues [@zhang_yu2024_chinese_dream]. The relevant question is whether Folha increasingly foregrounded China as a trade relationship after the rank reversal. Headlines are the most visible textual cue in the archive, they are the part of coverage most likely to be scanned by broad audiences, and they are where newspapers condense complex developments into short public labels such as "largest trade partner" or "record exports." Full article texts would be useful for a different exercise -- for example, measuring tone or detailed argumentation -- but the salience claim here is about which China-related themes were made visible at the headline level.

I therefore classify China-related headlines into substantive categories and then collapse them into four analytically relevant groups: China-Brazil trade, China's domestic economy, diplomacy and bilateral relations, and non-economic coverage. The classification uses a fixed label set with examples, and the appendix reports the exact prompt and validation procedure. Figure \@ref(fig:plot-media-counterfactual) shows that the post-2009 increase is concentrated in China-Brazil trade coverage rather than in all China-related topics equally. This pattern is consistent with a rank-salience interpretation: coverage concentrates on China's new commercial status in Brazil.

```{r plot-media-counterfactual, message=FALSE, warning=FALSE, echo=FALSE, fig.cap=paste0("Brazilian media salience: disaggregated China headlines in Folha de S.Paulo by category. The sustained post-2009 increase is concentrated in trade-related coverage. ", caption_note(unit = "Headlines per category per year", window = "2000-2014 annual series")), fig.pos="H", fig.asp = 0.7, fig.width = 8}
mc <- tar_read(media_counterfactual)
ggplot(mc, aes(x = year, y = n, fill = broad_category)) +
  geom_area(alpha = 0.7) +
  geom_vline(xintercept = 2009, linetype = "dashed", colour = "grey30") +
  annotate("text", x = 2009.3, y = max(mc$total) * 0.95, label = "Rank reversal",
           hjust = 0, size = 3, colour = "grey30") +
  scale_x_continuous(breaks = seq(2000, 2014, by = 2)) +
  scale_fill_manual(values = c(
    "China-Brazil Trade" = "#d95f02",
    "Chinese Economy" = "#7570b3",
    "Diplomacy & Relations" = "#1b9e77",
    "Non-economic (sports, health, disasters, etc.)" = "#999999"
  ),
  labels = c(
    "China-Brazil Trade" = "China-Brazil trade",
    "Chinese Economy" = "Chinese economy",
    "Diplomacy & Relations" = "Diplomacy/relations",
    "Non-economic (sports, health, disasters, etc.)" = "Other non-economic"
  )) +
  labs(x = "Year", y = "Number of headlines", fill = "Category") +
  guides(fill = guide_legend(nrow = 2, byrow = TRUE)) +
  theme_minimal() +
  theme(
    legend.position = "bottom",
    legend.text = element_text(size = 8),
    legend.title = element_text(size = 8)
  )
```

\clearpage

The media evidence identifies the attention step in the mechanism. After the 2009 status change, China became more visible in Brazilian public discourse, and the increase was concentrated in trade-related coverage. This pattern matters because the theory requires the rank cue to become publicly available. Primary-source evidence then shows that the same cue entered official diplomatic language. Brazilian officials repeatedly invoked China’s new position in Brazil’s trade hierarchy when explaining diplomatic prioritization, bilateral upgrading, and institutional follow-up.

In a written interview on the eve of his May 2009 visit to Beijing, President Lula described China as Brazil’s leading trading partner and framed the trip as an effort to consolidate cooperation across commerce, technology, and global economic governance [@mre_funag_resenha104_2009, 323]. In Beijing three days later, Lula repeated the rank claim in speeches organized around the Brazil-China Strategic Partnership, trade diversification, investment, presidential diplomacy, and the preparation of the 2010–2014 Joint Action Plan [@mre_funag_resenha104_2009, 115, 121–122]. The cue then persisted in Itamaraty language. When announcing Hu Jintao’s 2010 state visit and the signature of the Joint Action Plan, the Ministry of Foreign Affairs again invoked China’s 2009 rise to the top of Brazil’s trade hierarchy [@mre_funag_resenha106_2010, 356]. Dilma Rousseff and Hu Jintao’s 2011 joint communiqué likewise noted that China had become Brazil’s principal trading partner in 2009 and embedded that status in an agenda of intensified dialogue over trade, investment, diversification, and implementation of the Joint Action Plan [@mre_funag_resenha108_2011, 160–161]. Later that year, Foreign Minister Antonio Patriota made the institutional link explicit: in discussing China, he announced a dedicated Itamaraty task force to monitor economic-commercial relations with Brazil’s principal trading partner and to propose ways to move the relationship beyond complementarity, toward diversification and higher value-added trade [@mre_funag_resenha109_2011, 47].

These statements show that the 2009 status cue was absorbed into official language and used to justify diplomatic prioritization and institutional attention. This is precisely the public justificatory mechanism implied by the theory: once China was officially recognized as central in Brazil’s economic hierarchy, selective diplomatic accommodation could be presented as pragmatic recognition of national economic interests rather than as ideological concession or opportunistic bandwagoning.

# Cross-Country Evidence Beyond Brazil

The cross-country analysis serves two purposes. First, it tests whether persistent top-rank trade status has causal consequences beyond the Brazilian case. Second, it is an additional robustness check against explanations tied to Brazil's 2009 calendar year, such as the global financial crisis, the BRICS/G20 moment, or the commodity cycle. If persistent top-rank trade status matters, countries where China enters and holds the largest export-destination position should move closer to China in UNGA ideal-point space when aligned by their own treatment entry. I therefore estimate the same treatment logic in a cross-country absorbing sample: treated countries are those where China becomes the largest export destination after 2000, after an observed prior non-China-top year, and then does not lose that position. Never-treated countries remain controls. This analysis tests whether the rank-threshold pattern travels beyond the theory-generating case while matching the durable treatment observed in Brazil.

```{r fect-results, include=FALSE}
library(dplyr)

fect_ife_fit <- tar_read(fect_ife_china_top)
fect_fe_fit <- tar_read(fect_fe_china_top)
fect_ife_cov_fit <- tar_read(fect_ife_china_top_cov)
abs_panel <- tar_read(china_top_fect_data)

ife_s <- fect_att_summary(fect_ife_fit)
ife_cov_s <- fect_att_summary(fect_ife_cov_fit)
china_top_s <- tar_read(fect_ife_china_top_summary)
china_top_cov_s <- tar_read(fect_ife_china_top_cov_summary)
cs_abs_s <- tar_read(did_china_top_absorbing_summary)
cs_abs_cov_s <- tar_read(did_china_top_absorbing_cov_summary)
short_lived_cross_tbl <- tar_read(cross_country_short_lived_appendix_table)
recent_pretrend_tbl <- tar_read(cross_country_recent_pretrend_f_test_table)

n_countries_abs <- dplyr::n_distinct(abs_panel$iso3c)
n_treated_abs <- dplyr::n_distinct(abs_panel$iso3c[abs_panel$china_top == 1])
n_control_abs <- n_countries_abs - n_treated_abs
panel_min <- min(abs_panel$year)
panel_max <- max(abs_panel$year)

treat_summary <- abs_panel %>%
  dplyr::group_by(iso3c, country_name) %>%
  dplyr::summarise(
    treated_years = sum(china_top),
    entries = sum(china_top == 1L & dplyr::lag(china_top, default = 0L) == 0L),
    exits = sum(china_top == 0L & dplyr::lag(china_top, default = 0L) == 1L),
    .groups = "drop"
  )
placebo_p <- if (!is.null(fect_fe_fit$test.out$placebo.p)) round(fect_fe_fit$test.out$placebo.p, 3) else NA
placebo_equiv_p <- if (!is.null(fect_fe_fit$test.out$placebo.equiv.p)) round(fect_fe_fit$test.out$placebo.equiv.p, 3) else NA
ife_placebo_f_p <- if (!is.null(fect_ife_fit$test.out$f.p)) round(fect_ife_fit$test.out$f.p, 3) else NA
ife_placebo_tost_p <- if (!is.null(fect_ife_fit$test.out$tost.equiv.p)) round(fect_ife_fit$test.out$tost.equiv.p, 3) else NA
format_p <- function(x) {
  if (length(x) == 0 || is.na(x)) {
    return("not available")
  }
  formatC(x, format = "f", digits = 3)
}
fmt3 <- function(x) sprintf("%.3f", unname(x))
fmt1 <- function(x) sprintf("%.1f", unname(x))
fmt_num <- function(x, digits = 3L) {
  ifelse(is.na(x), "NA", formatC(x, format = "f", digits = digits))
}
sig_sentence <- function(att, p) {
  direction <- ifelse(att < 0, "in the theoretically expected negative direction", "positive")
  if (is.na(p)) {
    "The statistical precision of this estimate is not available in the current target output."
  } else if (p < 0.05) {
    paste0("The estimate is ", direction, " and statistically distinguishable from zero at the 5 percent level.")
  } else {
    paste0("The estimate is ", direction, ", but the confidence interval includes zero.")
  }
}
```

The preferred cross-country specification is the no-covariate IFE model on this absorbing sample. The estimate is `r fmt3(china_top_s$att)` ideal-point units (bootstrap SE = `r fmt3(china_top_s$se)`, 95 percent CI [`r fmt3(china_top_s$ci_lo)`, `r fmt3(china_top_s$ci_hi)`], p = `r fmt3(china_top_s$p)`). It corresponds to about `r fmt1(china_top_s$att_rel_pct)` percent of the pre-treatment mean distance to China, or `r fmt3(china_top_s$att_sd_units)` standard deviations of the outcome. `r sig_sentence(china_top_s$att, china_top_s$p)` The covariate-adjusted IFE estimate, with log GDP per capita and V-Dem freedom of expression, is `r fmt3(china_top_cov_s$att)` (p = `r fmt3(china_top_cov_s$p)`). The C\&S estimates in Table \@ref(tab:cross-country-table) use the same absorbing treatment logic as a convergent check. The staggered timing of persistent China-top entries, the absence of near-pre-treatment divergence, the interactive fixed-effects adjustment, and the convergence between IFE and C\&S estimates make a causal interpretation of the cross-country estimates credible. The cross-country evidence is smaller than the Brazil SDiD effect, consistent with Brazil being a most-likely case in which China displaced the United States and the rank reversal was publicly salient. The direction is the same: durable China-top export-destination status reduces UNGA distance to China beyond the Brazilian case. Appendix Table \@ref(tab:incumbent-salience-scope-table) reports a compact scope diagnostic for displaced-incumbent salience, excluding hub/entrepot cases and the most influential treated country from leave-one-out diagnostics.

```{r cross-country-table, message=FALSE, warning=FALSE, echo=FALSE, results='asis'}
cross_tbl <- tar_read(cross_country_absorbing_main_table) %>%
  dplyr::mutate(
    model = dplyr::case_when(
      stringr::str_detect(model, "Main") ~ "Main IFE",
      stringr::str_detect(model, "Robustness") ~ "IFE + covariates",
      stringr::str_detect(model, "covariates") ~ "C\\&S + covariates",
      TRUE ~ "C\\&S check"
    ),
    estimator = dplyr::case_when(
      stringr::str_detect(estimator, "fect") ~ "IFE",
      stringr::str_detect(estimator, "Callaway") ~ "C\\&S",
      TRUE ~ estimator
    ),
    `ATT` = fmt3(att),
    `SE` = fmt3(se),
    `95\\% CI` = paste0("[", fmt3(ci_lo), ", ", fmt3(ci_hi), "]"),
    `p-value` = fmt3(p),
    `Country-years` = as.character(n_obs),
    `Units (T/C)` = paste0(n_treated, "/", n_control),
    `Treatment entries` = as.character(n_entries),
    `Treatment exits` = ifelse(is.na(n_exits), "NA", as.character(n_exits)),
    `Panel window` = paste0(panel_min, "--", panel_max)
  ) %>%
  dplyr::transmute(
    Model = model,
    Estimator = estimator,
    Covariates = covariates,
    ATT,
    SE,
    `95\\% CI`,
    `p-value`,
    `Units (T/C)`,
    `Panel window`,
    `r*` = latent_factors
  )
cross_note <- paste0(
  sub(
    "^Note: ",
    "",
    caption_note(
      unit = "ATT in absolute UNGA ideal-point distance to China; lower values indicate convergence toward China",
      se = "bootstrap for IFE rows; analytical country-clustered SEs for Callaway-Sant'Anna rows",
      window = sprintf("%d--%d main panel; covariate rows use complete cases where noted", china_top_s$panel_min, china_top_s$panel_max)
    )
  ),
  " IFE = interactive fixed effects; CS = Callaway-Sant'Anna; Units (T/C) = treated/control units. The no-covariate IFE row is the main cross-country estimate. The covariate-adjusted IFE row is a robustness check. The CS rows use the same absorbing-treatment logic."
)

knitr::kable(
  cross_tbl,
  format = "latex",
  booktabs = TRUE,
  escape = FALSE,
  caption = "Cross-country estimates for persistent China top export-destination status.",
  align = "lllccccccc"
) %>%
  kableExtra::kable_styling(
    latex_options = c("HOLD_position", "scale_down"),
    font_size = 8
  ) %>%
  kableExtra::footnote(
    general = cross_note,
    general_title = "Note: ",
    escape = FALSE,
    threeparttable = TRUE
)
```

Figure \@ref(fig:cross-country-dynamic) shows the entry-aligned dynamic pattern for the main absorbing-sample IFE specification. The pre-entry estimates are diagnostics for model fit, and the post-entry path indicates whether the panel pattern is consistent with the Brazil result.

```{r cross-country-dynamic, message=FALSE, warning=FALSE, echo=FALSE, fig.cap=paste0("Entry-aligned dynamic effects from the main cross-country IFE model. Negative estimates indicate reduced UNGA distance to China after persistent top-rank entry. ", caption_note(unit = "Effect on absolute UNGA ideal-point distance to China", se = "bootstrap confidence intervals", window = sprintf("%d--%d absorbing panel", china_top_s$panel_min, china_top_s$panel_max))), fig.pos="H", fig.width=7, fig.asp=0.65}
cross_dynamic_plot <- tar_read(plot_fect_ife_gap_china_top)
cross_dynamic_plot$layers <- cross_dynamic_plot$layers[
  !vapply(
    cross_dynamic_plot$layers,
    function(layer) {
      identical(class(layer$geom)[1], "GeomText") &&
        identical(layer$aes_params$label, "14")
    },
    logical(1)
  )
]
cross_dynamic_plot
```

The two designs test nested implications of the argument. Brazil is the high-salience case: China displaced the United States, the hegemonic benchmark, and the attention mechanism can be observed directly in media and official discourse. The cross-country panel tests a broader scope implication: whether persistent China-top export-destination status is followed by movement toward China even when the displaced incumbent is not necessarily the United States and local salience is less directly observable. Exploratory subgroup diagnostics are consistent with this reading. The estimated effect remains negative among cases where China did not displace the United States, which helps rule out the narrow concern that the panel result is simply a Brazil-like hegemonic-displacement effect. Because treated cells are small and non-U.S. incumbents remain heterogeneous, I interpret these subgroup estimates as scope evidence.

The cross-country estimand is persistent China-top export-destination status. This is the treatment implied by the theory: durable status consolidation. Short-lived entries are reported separately in the appendix as a distinct treatment regime with weaker status content.

# Conclusion

The evidence supports a rank-threshold account of UNGA voting convergence toward China. In Brazil, China's 2009 move into the top export-destination position coincided with a publicly visible status cue, and SDiD estimates show that Brazil moved closer to China relative to a synthetic counterfactual. The issue-area and vote-level evidence suggest that this convergence was selective rather than mechanical: the strongest additional support comes from human-rights resolutions where China and the United States diverged, and the human-rights shift is larger than the analogous non-human-rights shift. The media evidence shows why the rank cue was politically usable: China's new commercial status became visible in public discourse. Cross-country estimates from the IFE model then show that the rank-threshold pattern travels beyond Brazil. The mechanism is measured most directly in the Brazilian case, and the panel evidence extends the argument beyond the theory-generating case: durable top-rank trade milestones can matter for multilateral alignment.

The paper therefore shifts attention from trade volume alone to the political meaning of trade hierarchies. Continuous exposure matters, but actors also respond to categorical milestones: becoming first, displacing a hegemonic benchmark, and making an economic relationship publicly legible. These thresholds can create opportunities for selective diplomatic adjustment even when the underlying commercial relationship has been growing for years.

Future research can extend the design in three directions. First, cross-country media data would allow the salience step to be tested beyond Brazil. Second, qualitative work on policymakers and business actors could identify how the rank cue entered diplomatic deliberation. Third, similar rank-threshold designs could be applied to investment, aid, lending, or other economic ties where categorical milestones may become politically usable.

\clearpage

# References {-}

::: {#refs}
:::

\clearpage

# Appendix

We present more details about the performance and checks on the fitted model and how we used ChatGPT to perform the NLP analysis.

## Brazil and China Ideal-Point Series

Figure \@ref(fig:appendix-brazil-china-ideal-points) plots the separate UNGA ideal-point series for Brazil and China. This diagnostic clarifies the interpretation of the absolute-distance outcome used in the main SDiD design: China remains comparatively stable, while Brazil accounts for most of the movement in the Brazil-China distance.

```{r appendix-brazil-china-ideal-points, message=FALSE, warning=FALSE, echo=FALSE, fig.cap=paste0("Brazil and China UNGA ideal-point series. The dashed line marks 2009. The figure separates the two series behind the absolute-distance outcome used in the main SDiD analysis. ", caption_note(unit = "UNGA ideal-point score (Bailey et al. scale)", window = "1990-2022 annual series")), fig.pos="H", fig.width=7, fig.asp=0.55}
ideal_series_brazil_china <- tar_read(unga_data) %>%
  dplyr::filter(iso3c %in% c("BRA", "CHN")) %>%
  dplyr::mutate(
    country = dplyr::case_when(
      iso3c == "BRA" ~ "Brazil",
      iso3c == "CHN" ~ "China",
      TRUE ~ iso3c
    )
  ) %>%
  dplyr::select(year, country, ideal_point_all)

ggplot(ideal_series_brazil_china, aes(x = year, y = ideal_point_all, colour = country)) +
  geom_line(linewidth = 0.6) +
  geom_point(size = 1.6) +
  geom_vline(xintercept = 2009, linetype = "dashed", colour = "grey35") +
  scale_colour_manual(values = c("Brazil" = "darkgreen", "China" = "firebrick")) +
  labs(x = "Year", y = "UNGA ideal point", colour = NULL) +
  theme_minimal() +
  theme(legend.position = "bottom")
```

## Resolution-Level UNGA Vote Diagnostics

To make the Brazil-China movement in UNGA ideal points more concrete, Table \@ref(tab:table-brazil-china-unvotes-year) reports the annual share of resolutions in which Brazil and China cast identical votes from 2005 to 2012. The diagnostic uses roll-call votes from `unvotes` 0.3.0, based on Erik Voeten's UNGA voting data, and excludes missing or absent votes by either Brazil or China.

```{r table-brazil-china-unvotes-year, message=FALSE, warning=FALSE, echo=FALSE, results='asis'}
bc_year_table <- tar_read(brazil_china_unvotes_similarity_by_year_2005_2012) %>%
  dplyr::mutate(
    identical_votes_percent = sprintf("%.1f", identical_votes_percent),
    mean_similarity_score = sprintf("%.3f", mean_similarity_score)
  ) %>%
  dplyr::select(
    year,
    n_resolutions,
    identical_votes_n,
    identical_votes_percent,
    mean_similarity_score
  )

knitr::kable(
  bc_year_table,
  format = "latex",
  booktabs = TRUE,
  caption = "Brazil-China identical UNGA votes by year, 2005-2012.",
  col.names = c(
    "Year",
    "Resolutions",
    "Identical votes",
    "Identical votes (\\%)",
    "Mean similarity score"
  ),
  align = "ccccc"
) %>%
  kableExtra::kable_styling(
    latex_options = c("HOLD_position", "scale_down"),
    font_size = 9
  ) %>%
  kableExtra::footnote(
    general = sub(
      "^Note: ",
      "",
      caption_note(
        unit = "Roll-call resolutions with valid Brazil and China votes; similarity score ranges from 0 to 1",
        window = "2005-2012 annual series"
      )
    ),
    general_title = "Note: ",
    escape = FALSE,
    threeparttable = TRUE
  )
```

## SDiD

### Comparison of DiD, SCM, and SDiD estimating equations

To gain intuition for the SDiD estimator, it is useful to compare it with standard DiD and SCM. A DiD model estimates:

$$
\bigl(\hat{\tau}_{ATT}^{\text{DiD}}, \hat{\mu}, \hat{\alpha}, \hat{\beta}\bigr)
=
\operatorname*{arg\,min}_{\tau_{ATT},\mu, \alpha, \beta}
\sum_{i=1}^{N}\sum_{t=1}^{T}
\bigl(Y_{it}-\mu-\alpha_i -\beta_t-D_{it}\tau_{ATT}\bigr)^2
$$

While SDiD assigns weights to the control group that are more similar to those of the treated unit and in time periods comparable to the pre-treatment period, classic DiD does not use any weights at all.

The SCM model can be written as follows:

$$
\bigl(\hat{\tau}_{ATT}^{\text{SCM}}, \hat{\mu}, \hat{\beta}\bigr)
=
\operatorname*{arg\,min}_{\tau_{ATT}, \mu, \beta}
\sum_{i=1}^{N}\sum_{t=1}^{T}
\bigl(Y_{it}-\mu-\beta_t-D_{it}\tau_{ATT}\bigr)^2\,
\hat{w}^{\text{SCM}}
$$

The equations show there is no unit fixed effect and no time weight in SCM. As a result, SCM cannot use unit-specific unobserved invariant characteristics to approximate the counterfactual outcome, and must consider the entire period when estimating the causal effect. SDiD combines unit weights (like SCM) with unit fixed effects and time weights (like DiD), addressing the limitations of both. The basic SDiD estimating equation is:

$$
\bigl(\hat{\tau}_{ATT}^{\text{SDiD}}, \hat{\mu}, \hat{\alpha}, \hat{\beta} \bigr)
=
\operatorname*{arg\,min}_{\tau_{ATT},\mu,\beta, \alpha}
\sum_{i=1}^{N}\sum_{t=1}^{T}
\bigl(Y_{it}- \mu- \alpha_i - \beta_t-D_{it}\tau_{ATT}\bigr)^2\,
\hat{w}^{\text{SDiD}} \hat{\lambda}^{\text{SDiD}}_t
$$

Note that SDiD retains unit fixed effects $\alpha_i$ (like DiD) while also assigning unit weights $\hat{w}^{\text{SDiD}}$ (like SCM) and time weights $\hat{\lambda}^{\text{SDiD}}_t$ (unique to SDiD).

### SDiD weights and donor pool

Below we present the weights for the top 10 controls used in the main model.

```{r plot-weights, message=FALSE, warning=FALSE, echo=FALSE, fig.cap=paste0("Weights of countries contributing to the synthetic control in the Brazilian SDiD model. ", caption_note(unit = "Synthetic-control unit weights, summing to 1", window = "pre-treatment period ending in 2008")), fig.pos="H"}
tar_read(plot_weights_coef)
```

We also fit a model for a donor pool of Latin American states (excluding Caribbean countries), which is more similar to Brazil than using countries worldwide. One reason to prefer a pool of Latin American countries is that a smaller donor pool helps avoid overfitting. When comparing results, the estimated causal effect is quite similar (a bit larger in absolute terms), which lends credibility to the main estimation. However, the small number of controls introduces noise and increases the uncertainty of the model. It is known that the usage of the placebo method to compute standard errors inflates the uncertainty of the estimation. For this reason, we prefer the model with the larger donor pool. Nonetheless, it is reassuring that with a donor pool composed of Latin American countries, which are arguably more similar to Brazil, the estimate is quite similar. Below is the plot of the model fitted for Latin America.


```{r plot-latam, message=FALSE, warning=FALSE, echo=FALSE, fig.cap=paste0("SDiD fit with a Latin American donor pool. ", caption_note(unit = "Absolute UNGA ideal-point distance to China", se = "placebo-based, reported in text", window = "years shown in the plot")), fig.pos="H"}
tar_read(plot_trend_latam)
b_est_latam <- tar_read(synth_fit_latam)[1]          # LATAM donor-pool estimate
se_latam    <- tar_read(se_synth_latam)
```

The estimate is `r round(b_est_latam,2) ` with a standard error of `r round(se_latam,2) `. In any case, it is interesting to examine the countries with the largest weights. Argentina, Guatemala, Paraguay, and Uruguay -- all but Guatemala are founding members of Mercosur. They contribute the most to synthetic control.

```{r plot-weight-latam, message=FALSE, warning=FALSE, echo=FALSE, fig.cap=paste0("Synthetic-control donor weights for the Latin American SDiD specification. ", caption_note(unit = "Synthetic-control unit weights, summing to 1", window = "pre-treatment period ending in 2008")), fig.pos="H"}
tar_read(plot_weights_coef_latam)
```

## Cross-Country Convergent Check: Absorbing-Treatment Estimator (C&S)

As a convergent check, we also report an absorbing-treatment estimator using @callaway_santanna2021. This specification is built from the same persistent China-top-export-destination sample used by the main IFE analysis. Treated units are countries for which China first becomes the number-one export destination in or after 2000 after an observed untreated year and then never loses the top position through the end of the panel. Countries with short-lived treatment entries are excluded from this C\&S comparison, not treated as controls. The baseline absorbing C\&S estimate uses `r cs_abs_s$n_treated` treated units and `r cs_abs_s$n_control` never-treated controls over `r cs_abs_s$panel_min`--`r cs_abs_s$panel_max`; the covariate-adjusted specification uses `r cs_abs_cov_s$n_treated` treated units and `r cs_abs_cov_s$n_control` controls over the complete-case window `r cs_abs_cov_s$panel_min`--`r cs_abs_cov_s$panel_max`.

This check estimates the same persistent-treatment logic with a different estimator. The covariate-adjusted C\&S result is additionally a complete-case sensitivity diagnostic: small treated groups and covariate adjustment can create overlap/convergence issues, and the changed panel window and treated/control counts mean that comparison with the baseline C\&S column does not isolate covariate adjustment alone.

```{r plot-es-china-top-absorbing, message=FALSE, warning=FALSE, echo=FALSE, fig.cap=paste0("Dynamic treatment effects from the absorbing-treatment C\\&S specification. ", caption_note(unit = "ATT in absolute UNGA ideal-point distance to China", se = "analytical 95 percent confidence intervals", cluster = "country level", window = sprintf("%d-%d balanced panel", cs_abs_s$panel_min, cs_abs_s$panel_max))), fig.pos="H", fig.width=7, fig.asp=0.65}
tar_read(plot_es_china_top_absorbing)
```

## Cross-Country Robustness: Short-Lived Treatment Entries

The theory expects clearer effects when the new trade hierarchy persists long enough to become politically visible, bureaucratically usable, and diplomatically consequential. Short-lived entries are therefore a distinct, weaker treatment regime. Table \@ref(tab:cross-country-short-lived-table) reports IFE estimates for countries where China becomes the largest export destination but later loses that position, with never-treated countries retained as controls. These estimates are smaller and not statistically distinguishable from zero, consistent with the claim that transient rank changes are weaker signals of trade-based status.

```{r cross-country-short-lived-table, message=FALSE, warning=FALSE, echo=FALSE, results='asis'}
short_lived_tbl <- short_lived_cross_tbl %>%
  dplyr::mutate(
    model = dplyr::case_when(
      stringr::str_detect(model, "covariates") ~ "IFE + covariates",
      TRUE ~ "IFE"
    ),
    estimator = dplyr::if_else(stringr::str_detect(estimator, "fect"), "Interactive FE", estimator),
    `ATT` = fmt3(att),
    `SE` = fmt3(se),
    `95\\% CI` = paste0("[", fmt3(ci_lo), ", ", fmt3(ci_hi), "]"),
    `p-value` = fmt3(p),
    `Treated units` = as.character(n_treated),
    `Control units` = as.character(n_control),
    `Panel window` = paste0(panel_min, "--", panel_max)
  ) %>%
  dplyr::transmute(
    Model = model,
    Estimator = estimator,
    Covariates = covariates,
    ATT,
    SE,
    `95\\% CI`,
    `p-value`,
    `Treated units`,
    `Control units`,
    `Panel window`,
    `Latent factors` = latent_factors
  )

knitr::kable(
  short_lived_tbl,
  format = "latex",
  booktabs = TRUE,
  escape = FALSE,
  caption = "Short-lived China top export-destination treatment entries.",
  align = "lllcccccccc"
) %>%
  kableExtra::kable_styling(
    latex_options = c("HOLD_position", "scale_down"),
    font_size = 7
  ) %>%
  kableExtra::footnote(
    general = sub(
      "^Note: ",
      "",
      caption_note(
        unit = "ATT in absolute UNGA ideal-point distance to China; lower values indicate convergence toward China",
        se = "bootstrap with 10,000 replications",
        cluster = "country level"
      )
    ),
    general_title = "Note: ",
    escape = FALSE,
    threeparttable = TRUE
  )
```

## Cross-Country Scope Robustness: Displaced-Incumbent Salience

Table \@ref(tab:incumbent-salience-scope-table) reports a compact scope diagnostic for the main absorbing IFE estimate. It asks whether the main cross-country estimate is sensitive to two salience-related restrictions: excluding treated countries whose displaced incumbent is a hub or entrepot, and excluding the single treated country that most attenuated the preliminary leave-one-out estimate.

```{r incumbent-salience-scope-table, message=FALSE, warning=FALSE, echo=FALSE, results='asis'}
incumbent_scope_tbl <- tar_read(cross_country_incumbent_salience_appendix_table) %>%
  dplyr::mutate(
    `ATT` = fmt3(att),
    `SE` = fmt3(se),
    `95\\% CI` = paste0("[", fmt3(ci_lo), ", ", fmt3(ci_hi), "]"),
    `p` = fmt3(p),
    `Treated countries` = as.character(n_treated),
    `Controls` = as.character(n_control)
  ) %>%
  dplyr::transmute(
    Specification = specification,
    `Excluded treated cases` = excluded_treated_cases,
    ATT,
    SE,
    `95\\% CI`,
    p,
    `Treated countries`,
    Controls
  )

knitr::kable(
  incumbent_scope_tbl,
  format = "latex",
  booktabs = TRUE,
  escape = FALSE,
  caption = "Scope robustness for displaced-incumbent salience in the absorbing cross-country IFE sample.",
  align = "llcccccc"
) %>%
  kableExtra::kable_styling(
    latex_options = c("HOLD_position", "scale_down"),
    font_size = 7
  ) %>%
  kableExtra::footnote(
    general = paste0(
      "All rows use the no-covariate interactive fixed-effects estimator on the ",
      "absorbing China-top sample, with 10,000 bootstrap replications. ",
      "The hub/entrepot row excludes Malaysia and Sierra Leone, whose displaced ",
      "incumbents are Singapore and Belgium. The leave-one-out row excludes ",
      "Solomon Islands, the treated country whose exclusion most attenuated the ",
      "preliminary Callaway-Sant'Anna ATT. These rows are scope diagnostics for salience."
    ),
    general_title = "Note: ",
    escape = FALSE,
    threeparttable = TRUE
  )
```

## Cross-Country Diagnostic: No-Pretrend Equivalence Tests

Following @liu2024practical and @chiu_etal2025, I report equivalence diagnostics for the no-pretrend assumption. Unlike the conventional placebo test (which tests $H_0$: no pre-trend), the equivalence test reverses the null to $H_0$: pre-trends are meaningfully large, providing positive evidence for the absence of pre-trends when rejected. In the absorbing cross-country sample, the broad native IFE F-test over all selected nonpositive event times is not defined because that diagnostic tries to test more pre-entry periods than the treated-unit support can sustain. The native warning is therefore a support limitation, so Table \@ref(tab:recent-pretrend-f-test) reports a supplementary recent-window diagnostic. The diagnostic uses the most recent nonpositive event times, capped at 12 and reduced only when required by degrees of freedom, bootstrap support, or covariance rank. The IFE estimates themselves are unchanged.

```{r recent-pretrend-f-test, message=FALSE, warning=FALSE, echo=FALSE, results='asis'}
recent_pretrend_display <- recent_pretrend_tbl %>%
  dplyr::mutate(
    model = dplyr::case_when(
      stringr::str_detect(model, "covariates") ~ "IFE + covariates",
      stringr::str_detect(model, "Main") ~ "Main IFE",
      TRUE ~ model
    ),
    `F statistic` = fmt_num(f_stat),
    `F p-value` = fmt_num(f_p),
    `Equivalence p-value` = fmt_num(f_equiv_p),
    `TOST p-value` = fmt_num(tost_equiv_p),
    `Selected event times` = selected_periods,
    `Bootstrap draws` = as.character(n_valid_boots),
    `Covariance rank` = as.character(cov_rank)
  ) %>%
  dplyr::transmute(
    Model = model,
    Status = test_status,
    `Selected event times`,
    `q` = q,
    `N-bar` = n_bar,
    `df2` = df2,
    `Bootstrap draws`,
    `Covariance rank`,
    `F statistic`,
    `F p-value`,
    `Equivalence p-value`,
    `TOST p-value`
  )

knitr::kable(
  recent_pretrend_display,
  format = "latex",
  booktabs = TRUE,
  escape = FALSE,
  caption = "Supplementary recent-window IFE pretrend F diagnostics.",
  align = "lllrrrrrrrrr"
) %>%
  kableExtra::kable_styling(
    latex_options = c("HOLD_position", "scale_down"),
    font_size = 7
  ) %>%
  kableExtra::footnote(
    general = paste0(
      "The table reconstructs the IFE F diagnostic using the most recent ",
      "nonpositive event times with enough treated-unit support for a joint test. ",
      "q is the number of jointly tested periods; N-bar is the treated-unit support ",
      "used in the F-test scaling; df2 = N-bar - q. Bootstrap draws are the ",
      "10,000 bootstrap replications nonmissing for all selected event times. ",
      "The broad native full-window F-test is not reported because its selected ",
      "window has insufficient treated-unit support."
    ),
    general_title = "Note: ",
    escape = FALSE,
    threeparttable = TRUE
  )
```

```{r plot-equiv-appendix, message=FALSE, warning=FALSE, echo=FALSE, fig.cap=paste0("No-pretrend equivalence plots for the FE and IFE cross-country models. FE placebo equivalence p = ", format_p(placebo_equiv_p), "; IFE equivalence TOST p = ", format_p(ife_placebo_tost_p), ". Table \\@ref(tab:recent-pretrend-f-test) reports the IFE recent-window F diagnostic. ", caption_note(se = "bootstrap, 10,000 replications", window = sprintf("%d-%d panel", panel_min, panel_max))), fig.pos="H", fig.width=10, fig.height=5}
tar_read(plot_diagnostics_equiv_china_top)
```

\clearpage

## ChatGPT Classification

Below is the coded instruction to the ChatGPT API to classify the subjects of the headlines.

```{r code-chatgpt, echo=TRUE, eval=FALSE}
system_block <- paste(
  'You are a Brazilian expert in International Relations and the political',
  'economy of China–Brazil relations.  Classify the subject of EACH headline',
  'when I say EACH, I mean all of them, even if repetitive or similar. Do classify **all** of them',
  'using **one** label from this list exactly as written:',
  '"diplomacy", "chinese economy", "china-brasil relations",',
  '"disaster and accidents", "chinese politics and policy",',
  '"sports, science, culture, other", "human rights",',
  '"china-brazil trade", "health".\n\n',
  # 13 static few-shot examples — leave them unchanged!
  'Examples:\n',
  '1. Presidente chinês visita o Brasil e participa de reunião com FHC -> china-brasil relations\n',
  '2. Ministro chinês de Defesa faz visita oficial ao Brasil -> china-brasil relations\n',
  '3. Para governo, novo câmbio chinês só beneficia Brasil no médio prazo -> china-brazil trade\n',
  '4. China ultrapassa EUA como maior parceiro comercial do Brasil -> china-brazil trade\n',
  '5. China vai controlar preços de alimentos e combustíveis -> chinese economy\n',
  '6. Vendas de automóveis na China desaceleram em junho -> chinese economy\n',
  '7. China quer devolução das relíquias Yin, patrimônio da Unesco -> sports, science, culture, other\n',
  '8. Dinossauro encontrado na China esclarece origem das aves -> sports, science, culture, other\n',
  '9. OMS confirma mais de 4.600 casos de gripe suína; China registra 2º caso -> health\n',
  '10. Marinha chinesa busca "cansar" barcos japoneses em águas disputadas -> diplomacy\n',
  '11. Acidente após casamento deixa 16 mortos na China -> disaster and accidents\n',
  '12. Jornal chinês sai das bancas por foto da praça Tiananmen -> human rights\n',
  '13. China é parceira estratégica contra crise econômica, diz UE -> chinese economy',
  'return as in the examples, a string with the headline and the classification'
)

```

The classified headline file is treated as an archived derived dataset rather than regenerated during manuscript rendering. This choice reflects the practical behavior of LLM-based coding: early runs sometimes returned incomplete classifications, altered original headlines, or used slightly inconsistent labels, which made exact matching back to the archive difficult. I therefore fixed the prompt, normalized labels, preserved the resulting classified file, and validate the coding below against an independently coded sample. The evidence in the main text is used as a headline-level salience diagnostic.

Below we provide a random sample of the headlines and the classification provided by ChatGPT.

```{r sample-chatgpt, message=FALSE, warning=FALSE, echo=FALSE, fig.cap=paste0("Random sample of China-related Folha de Sao Paulo headlines and ChatGPT classifications. ", caption_note(unit = "headline text and category labels", window = "sample drawn from the 2000-2014 corpus")), fig.pos="H", out.width="\\linewidth"}
knitr::include_graphics(tar_read(sample_headlines_image_file))
```

### Validation of ChatGPT Classification

To assess the reliability of the automated classification, one of the authors independently coded a stratified random sample of 100 headlines (with a minimum of 5 per category). Table \@ref(tab:chatgpt-validation) reports the agreement rate between the ChatGPT labels and the manual coding.

```{r chatgpt-validation, message=FALSE, warning=FALSE, echo=FALSE, results='asis'}
library(knitr)
library(dplyr)

validation_summary <- tar_read(chatgpt_validation_summary)
val_tbl <- validation_summary$table
overall_acc <- validation_summary$overall$accuracy_pct
validation_n <- validation_summary$overall$N
lowest_category <- validation_summary$lowest$category
lowest_acc <- validation_summary$lowest$accuracy_pct
category_acc <- validation_summary$category_accuracy
trade_acc <- category_acc$accuracy_pct[category_acc$category == "china-brazil trade"]
disaster_acc <- category_acc$accuracy_pct[category_acc$category == "disaster and accidents"]

kable(
  val_tbl,
  format = "latex",
  booktabs = TRUE,
  caption = paste0("Validation of ChatGPT classification against manual coding (N = ", validation_n, "). Overall accuracy: ", sprintf("%.1f", overall_acc), " percent."),
  align = "lccc"
) %>%
  kableExtra::kable_styling(
    latex_options = c("HOLD_position", "scale_down"),
    font_size = 9
  ) %>%
  kableExtra::footnote(
    general = sub(
      "^Note: ",
      "",
      caption_note(
        unit = "headline-level counts and accuracy (percent) by category",
        window = "stratified validation sample from the 2000-2014 headline corpus"
      )
    ),
    general_title = "Note: ",
    escape = FALSE,
    threeparttable = TRUE
  )
```

The overall accuracy is `r sprintf("%.1f", overall_acc)` percent. The lowest agreement is in the "`r lowest_category`" category (`r sprintf("%.1f", lowest_acc)` percent), where several headlines were manually reclassified as "diplomacy". The categories most relevant to our analysis --- "china-brazil trade" and "disaster and accidents" --- achieved `r sprintf("%.1f", trade_acc)` and `r sprintf("%.1f", disaster_acc)` percent agreement, respectively.

## Cross-Country: Absorbing Treated Countries

Table \@ref(tab:table-treated-appendix) describes the `r n_treated_abs` absorbing treated countries where China becomes the top export destination in the main cross-country panel and remains in that position. For each country, we report the treatment year and the mean UNGA voting distance to China before and after treatment onset.

```{r table-treated-appendix, message=FALSE, warning=FALSE, echo=FALSE}
library(kableExtra)

treated_tbl <- tar_read(china_top_fect_data) %>%
  dplyr::arrange(iso3c, year) %>%
  dplyr::group_by(iso3c, country_name) %>%
  dplyr::summarise(
    treated_ever = any(china_top == 1L, na.rm = TRUE),
    first_treat_year = if (any(china_top == 1L, na.rm = TRUE)) min(year[china_top == 1L], na.rm = TRUE) else NA_real_,
    entries = sum(china_top == 1L & dplyr::lag(china_top, default = 0L) == 0L, na.rm = TRUE),
    exits = sum(china_top == 0L & dplyr::lag(china_top, default = 0L) == 1L, na.rm = TRUE),
    mean_distance_pre = mean(abs_distance_china[year < first_treat_year], na.rm = TRUE),
    mean_distance_post = mean(abs_distance_china[year >= first_treat_year], na.rm = TRUE),
    .groups = "drop"
  ) %>%
  dplyr::filter(treated_ever) %>%
  dplyr::mutate(diff = mean_distance_post - mean_distance_pre) %>%
  dplyr::select(country = country_name, iso3c, first_treat_year, entries, exits,
                mean_distance_pre, mean_distance_post, diff)

kable(treated_tbl,
      format = "latex",
      booktabs = TRUE,
      digits = 3,
      caption = "Descriptive statistics of absorbing treated countries where China becomes the top export destination.",
      col.names = c("Country", "ISO3c", "Treatment year", "Entries",
                     "Exits", "Mean dist. (pre)", "Mean dist. (post)", "Diff"),
      align = "llcccccr") %>%
  kable_styling(latex_options = c("hold_position", "scale_down"),
                font_size = 9) %>%
  footnote(general = paste0("The main sample keeps treated countries only if China remains the top export destination from first treatment entry through the end of the sample. ",
                            "Mean dist. = absolute distance to China in UNGA ideal points. ",
                            "Diff = post minus pre."),
           general_title = "Note: ",
           escape = FALSE,
           threeparttable = TRUE)
```
