---
title: "The rank effect of trade on Brazil's foreign policy political alignment with China"
author: "Manoel Galino"
date: '2024-06-30'
output:
  pdf_document:
    number_sections: true
    latex_engine: xelatex
header-includes:
  - \usepackage{setspace}\doublespacing
  - \setstretch{1.5}
  - \usepackage{parskip}
  - \setlength{\parskip}{6pt}
  - \setlength{\parindent}{0pt}
  - \newcommand{\argmin}{\mathop{\mathrm{argmin}}\limits}
fontsize: 12pt
abstract: \singlespacing Loren ispon Loren ispon Loren ispon Loren ispon Loren ispon Loren ispon Loren ispon Loren ispon Loren ispon Loren ispon Loren ispon Loren ispon Loren ispon Loren ispon Loren ispon Loren ispon Loren ispon Loren ispon Loren ispon Loren ispon Loren ispon Loren ispon Loren ispon Loren isponLoren ispon Loren ispon Loren ispon Loren ispon Loren ispon Loren ispon Loren ispon Loren isponLoren ispon Loren ispon Loren ispon Loren ispon Loren ispon Loren ispon Loren ispon Loren isponLoren ispon Loren ispon Loren ispon Loren ispon Loren ispon Loren ispon Loren ispon Loren isponLoren ispon Loren ispon Loren ispon Loren ispon Loren ispon Loren ispon Loren ispon Loren isponLoren ispon Loren ispon Loren ispon Loren ispon Loren ispon Loren ispon Loren ispon Loren isponLoren ispon Loren ispon Loren ispon Loren ispon Loren ispon Loren ispon Loren ispon Loren ispon.

geometry: margin=2.5cm
spacing: double
mainfont: "Times New Roman"
bibliography: synth-trade-china.bib
link-citations: yes
---
\pagebreak

# Introduction

In 2004, China overcome Argentina as the second biggest Brazil's trade partner, behind only the US. At that same year, Brazil recognized China with economy market status, a major demand from China since 2001 [@urdinez_2014]. The economic ties between both countries only grew over time. In 2004, Brazil was the 15th largest market for China exports and became the eighth in 2014. The foreign policy of both countries also showed some alignment. Both countries became members of forums such as G5 (Brazil, Mexico, India, China and South Africa), BASIC (Brazil, South Africa, India and China) and the BRIC summit.

A major strand of the literature argues that expanding trade ties with China have significant foreign policy consequences for Brazil and other Latin American countries. Many of such studies, however, fall short of providing credible causal evidence. In a pioneering quantitative study, @flores-macias_kreps2013 show that the more countries trade with China, the more likely they are to side with China on foreign policy, in which the main measure of political alignment is vote similarity at UNGA. In order to to control for endogeneity, the author used a fixed effect models with lagged variables and two-stage least squares. The first approach is just wrong from a causal point of view, since one key identification assumption in fixed effects models is no carry over effects, which precludes the use of lagged dependent variable in the regression [@blackwell_glynn2018; @imai_kim2019; @imai_kim2021]. The instrument used in the latter approach is lagged energy production of a country. This variable, however, fails to meet the exclusion restriction assumption, since energy production is, for example, related to economic development, which may be causal pathway from the instrument to the outcome (political alignment). In general is very hard to justify the exclusion restriction of instruments [@mellon_2021; @bastardoz_2023; @lal_etal2024], diminishing the credibility of the assumptions needed for causal identification.

Qualitative studies have also pointed the linkage between trade and foreign policy. In a study about the prevalence of trade in Brazil's president Lula speech on China [@guilhon2014] shows that in fact that is the thematic emphasis of Brazil foreign policy during Lula's second term. From the other side, @vadell_2011 emphasizes how the diplomatic ties betwen China and South America countries followed the economic interest of the Asiatic power in the region.

Other studies have emphasized the reverse causal order. They contend that complementarity of aligned national interests and shared identities with the emerging discourse [@haibin_2010] and a revisionist stance towards the international order promoted more trade links between countries [@coutinho_rodriguez2024]. Many studies have find evidence that political alignment with China can significantly impact trade flows [@fuchs_klann2013; @aiyar_etal2024; @yan_zhou2021].

The opposite causal pathways make it hard to deal with the problem of reverse causality and omitted variables bias in observational studies. As a result, we still need to understand better the effect of trade on political alignment. Moreover, many of the existing studies do not differentiate exactly what is the effect working: is it its foreign direct investment in the region? The importance of its imports to the economic success of the latin american countries? Ore are other non-economic factors, such as China’s growing soft-power outreach or shared claims for reforming global governance institutions and resistance to Washington-centric norms that are important? In fact, more recently some studies have argued that it is a mixture of many of these forces working at the same time [@struver_2014; @schenoni_leiva2021].

Another limitation of the literature is that all quantitative studies that measured Brazil's foreign policy alignment with UNGA votes used some similarity measure based on percentage of votes at UNGA [@neto2011;  @flores-macias_kreps2013; @mouron_urdinez2014]. However, as @bailey_etal2017 shows, such vote cannot be used in a panel data set because such indicators do not separate changes in the UN's agenda from shifts in state preferences. As a result, many attribution of changes in the policy orientation are in fact change in the content of the votes.

The present study aims to deal with those three order of problems (causal credibility of findings, disentangle the effect of trade from other effects and separate agenda change from state preference reorientation). We do so leveraging a discontinuity in the rank of an otherwise continuous variable, values of exports to a country, couples with a synthethic difference in differences apporach to estimate the causal effecrt of China becoming the second, as opposed to third, major market for Brazilian exports to estimate its causal effect on foreign policy orientation, measured as the absolute distance in UNGA ideal point estimation [@bailey_etal2017] between Brazil and China over 1990-20014. We found that than change in the rank position decreased the distance by xx points.

Our results show that when salient, the trade position of a country has a distinctive causal effect on the foreign policy of a country like Brazil. All the arguments and discussion about China as an emerging power overlooked the importance of becoming salient in a qualitative distinct way to impact the foreign policy of other countries. It is not only a matter of having more trade ties, for instance, but overcoming historical trade partners that matter. This result also point to the need to investigating similar salience, rank effects, in other economic variables (such as Foreign Direct Investment and Loans) and variables along other dimensions (such as diplomatic ties, frequency of bilateral meetings and deals and agreements signed).


The rest of the paper is organized as follows. Next, we discuss the importance of salience effects in the realm of trade effects and political alignment. Then, we evaluate how to measure political alignment, our main outcome variable. In the fourth section of the paper, we present our methodology, data and variables. then, we present our empirical findings. The last section presents our concluding remarks.

# Political effects of rank

Decision makers live in a complex, noise world. In such situations, it is ecologically rational to use cues from the enviroment to chose to what pay attention and react accordingly [citation needed. Gingerezer]. Many studies in international relations on policmakers have emphasized the roles of cues and heuristics when making a foreign policy decision or reorientation.

A recent literature in the intersection of cognitive and social sciences has emphasizing the role herustics in decision-making, like how redundant information nonetheless work by shifting one's attention [@conlon2023], that people neglect correlated structure of information and treat them as independent [@enke_zimermann2017] or the existence of behavioral attenuation, in the sense that, due to information-processing constraints, people failt to corrected update beliefs and make decisions in response to change in the fundamentals of a problem [@enke_etal2024].

In the field of International Relations, the poliheuristics theory is a standard account on how decision makers also behave like regular people in this regard [@mintz2004; @mintz_2005]. On of the main arguments is that, due to information-processing constraints, poliheuristic decision-making process is based on a nonexhaustive search in the sense that “comparisons of policy options are made within a (very) restrictive alternative set and attribute set” (Mintz and Geva 1997, 84). As such, policy dimensions or issues that come to attention beggin to be considered from the perspective of decision makers as they weren't before.

Na capa da edição de 16 de maio de 2004, a Folha de São Paulo, the main newspaper in Brazil, put in the first page of the paper "Lula leva à China número recorde de empresários" [@trevisan2004]. The news piece within the paper says "As vendas tiveram alta de 80% em 2003, o que levou o país ao terceiro lugar entre os principais parceiros comerciais brasileiros, atrás de EUA e Argentina" [@trevisan2004, p. B8].







@anagol_fujiwara2016; @folke_etal2016

# Trade and Political Alignment

A major strand of the literature argues that expanding trade ties with China have significant foreign policy consequences for Brazil and other Latin American countries. Many of such studies, however, fall short of providing credible causal evidence. In a pioneering quantitative study, @flores-macias_kreps2013 shows that the more countries trade with China, the more likely they are to side with China on foreign policy, in which the main measure of political alignment is vote similarity at UNGA. In order to to control for endogeneity, the author used a fixed effect models with lagged variables and two-stage least squares. The first approach is just wrong from a causal point of view, since one key identification assumption in fixed effects models is no carry over effects, which precludes the use of lagged dependent variable in the regression [@blackwell_glynn2018; @imai_kim2019; @imai_kim2021]. The instrument used in the latter approach is lagged energy production of a country. This variable, however, fails to meet the exclusion restriction assumption, since energy production is, for example, related to economic development, which may be causal pathway from the instrument to the outcome (political alignment). In general is very hard to justify the exclusion restriction of instruments [@mellon_2021; @bastardoz_2023; @lal_etal2024], diminishing the credibility of the assumptions needed for causal identification.

Qualitative studies have also pointed the linkage between trade and foreign policy. In a study about the prevalence of trade in Brazil's president Lula speech on China,  [@guilhon2014] shows that in fact that is the thematic emphasis of Brazil foreign policy during Lula's second term. From the other side, @vadell_2011 emphasizes how the diplomatic ties betwen China and South America countries followed the economic interest of the Asiatic power in the region.

Other studies have emphasized the reverse causal order. They contend that complementarity of aligned national interests and shared identities with the emerging discourse [@haibin_2010] and a revisionist stance towards the international order promoted more trade links between countries [@coutinho_rodriguez2024]. Many studies have find evidence that political alignment (or its absence) can significantly impact trade flows. Countries that bucked China’s political preferences – by officially receiving the Dalai Lama – suffered subsequent trade declines [@fuchs_klann2013]. Countries that voted in support of the PRC (political alignment) in 1971 on the matters of Taiwan enjoyed a surge in bilateral trade with China in the immediately following year [@yan_zhou2021]. In a more general quantitative study, @chen_zhou2021 shows that a country tends to import more from partners with whom it has friendly relations, which is specially true for an authoritarian state (such as China). Studies have find a similar effect in other economic variables, such as Foreign Direct Investment [@aiyar_etal2024].

Another strand of the literature has emphasized how the emergence of China as a major power and its diplomatic inroads into Latin America opened up the possibility of more autonomy from the US in the region [@struver_2014]. @schenoni_leiva2021 argue that recent Brazilian foreign policy change is the result of the pull of China as a new great power, resulting in what they call a "dual hegemony". According to the ground break work by @neto2011, as Brazil became a more powerful country, it could distance itself from the US. It is thus expected that as China emerged as an important country, it would attract some alignment of Brazilian foreign policy due to this effect of power. @mouron_urdinez2014 argued that the power gap was the most important factor, and that it was true for the  broader South American countries: the smaller the power gap between a country and the U.S., the less it aligned with the U.S. at the UN.

@bailey_etal2017 developed an alternative measure, based on ideal point estimation, that disentangle the effect of agenda changing from the foreign policy preference change.

# Mehtodology

We explore a discontinuity in the change of China as the third versus second Brazil trade partner to estimate the (local) causal effect of increased importance of China becoming the second largest trade partner of Brazil. This allow us to approximate a difference-in-differences methodology (DiD). The problem with DiD is that it depends on the paralled trend assumption, which is not credible in the present case, leading to bias in the estimate of the causal effect.  In order to avoid this problems, we employ the synthetic difference in differences (SDiD) method,  developed by @arkhangelsky_etal2021. It combines the DiD estimator with the synthetic control method (SCM). It provides a more flexible approach than both methods separately. On one hand, DiD methods do not work well with a single treated unit and they also rely on the parallel trends assumption for identification fo the causal effect. On the ohter hand, SCM tries to match the pre-treatment trend in the outcome based on covariates to estimate the counterfactual outcome. However, it compares the whole time period before and after the intervention and does noy allow the inclusion of unit fixed effect, with can be a cause of bias, specially in the case of slow moving variables. SDID try to solves both problems (non-parallel trends for DiD, time period and absence of fixed effects for SCM) by combining both methods. The specific model is as follows:

Consider a complete panel^[whas is called balanced panel. We prefer the term complete to avoid confusion with balancing of covariates, which is a standard analysis in causal studies]: $N$ countries measured over $T$ years, the outcome variable for country $i$ in time $t$ is $Y_{it}$ and the binairy treatment variable $D_{it} \in \{0,1\}$. Suppose also that the first $N_{\text{countries}}$ do not receive the treatment and the lasdt country $N_{\text{Brazil}}$ do receive the treatment at $t_0$. The job is, like in SCM, to find weights $\hat{w}^{\text{SDiD}}$ such that before $t_0$ treated and control countries match as closely as possible, that is, $\sum_{i=1}^{\text{countries}} \hat{w}^{\text{SDiD}}Y_{it} \approx \sum_{i=1}^{\text{Brazil}} Y_{it}$, for all $t = 1, ..., t_0$. Differently from SCM, we do not stop here. We then find $\hat{\lambda}^{\text{SDiD}}_t$ to balance the time trend before and after $t_0$. Lastly, the average causal effect on the treated $\tau_{ATT}$ is estimated by a two-way fixed effect regression model (as in standard DiD models):

$$
\bigl(\hat{\tau}_{ATT}^{\text{SCM}}, \hat{\mu}, \hat{\beta} \bigr)
=
\operatorname*{arg\,min}_{\tau_{ATT},\mu,\beta, \alpha}
\sum_{i=1}^{N}\sum_{t=1}^{T}
\bigl(Y_{it}- \mu- \alpha_i - \beta_t-D_{it}\tau_{ATT}\bigr)^2\,
\hat{w}^{\text{SDiD}} \hat{\lambda}^{\text{SDiD}}_t
$$
Vale a pena contrastar esse modelo com o DiD clássico e o SCM. O modelo de DiD estima:
$$
\bigl(\hat{\tau}_{ATT}^{\text{SCM}}, \hat{\mu}, \hat{\lambda}, \hat{\beta}\bigr)
=
\operatorname*{arg\,min}_{\tau_{ATT},\mu, \alpha, \beta}
\sum_{i=1}^{N}\sum_{t=1}^{T}
\bigl(Y_{it}-\mu-\alpha_i -\beta_t-D_{it}\tau_{ATT}\bigr)^2
$$
Como se vê, o SDiD coloca pesos em países dos grupos controle que são mais similares ao país do tratamento e mais peso em períodos pré-tratamento que são mais similares ao período pós-tratamento. Por outro lado, o SC estima um modelo como:

$$
\bigl(\hat{\tau}_{ATT}^{\text{SCM}}, \hat{\mu}, \hat{\beta}\bigr)
=
\operatorname*{arg\,min}_{\tau_{ATT}, \mu, \beta}
\sum_{i=1}^{N}\sum_{t=1}^{T}
\bigl(Y_{it}-\mu-\beta_t-D_{it}\tau_{ATT}\bigr)^2\,
\hat{w}^{\text{SDiD}} \hat{\lambda}^{\text{SDiD}}_t
$$
As we can see, there is no unit fixed effect and not time weight. As a result, the SCM cannot use unit unobserved invariant characteristics to approximate the counter factual outcome and will consider the whole time period when estimating the causal effect. This is a problem for our data, because the relationship between China ans Brazil becomes more engangled as times gos by and years much further from the treatment period is proboably not a good comparison to periods longe before the treatment.

Anotther key advantage of the SDID is that it reduces the variance of the estimator, allowing more precise estimation [@arkhangelsky_etal2021]. One problem with this model, however, is that it does not include time-varying covariates that are important to more closely match the pre-treatment time period. Thus, to account  for time-varying covariates and better adjust the model, we include vector of time-varying controls as suggested by [@hirshberg_klosin2024], such that the estimated equation is now:

$$
\bigl(\hat{\tau}_{ATT}^{\text{SCM}}, \hat{\mu}, \hat{\beta}, \hat{\gamma} \bigr)
=
\operatorname*{arg\,min}_{\tau_{ATT},\mu,\beta, \alpha, \gamma}
\sum_{i=1}^{N}\sum_{t=1}^{T}
\bigl(Y_{it}- \mu- \alpha_i - \beta_t-D_{it}\tau_{ATT} -X_{it}\gamma \bigr)^2\,
\hat{w}^{\text{SDiD}} \hat{\lambda}^{\text{SDiD}}_t
$$


## Data and Variables

Since we need to fit a model to a complete panel, we constraint the sample to 105 countries over a period of 21 years, from 1991 to 2011.

Our outcome variable, vote similarity between a country and China at UNGA, is based on the ideal point estimates based on UNGA vote data from @bailey_etal2017. Other variables include per capita GDP from Penn World Table, version 10.01 [@feenstra_etal2015]. Trade data are drawn from UN Comtrade Database, extracted via their API in current FOB dollars. To avoid double counting, I use the value reported by the importer country, since importers have more interested in correcting measuring this value for tax collection reasons. We use trade with China and US as share of the country GDP as our two measures of trade. As our measure of power, we used the Global Powr Index (GPI) from Diplometric -- Pardee Institute, University of Denver. GPI combines demographic, economic, military, technology and diplomatic dimensions and then we computed the power gap between Brazil and the US. We also include distance to US, as countries closer to US will hhave a bigger incentive to align its foreign policy with it and exachange rate, since one complain of China is its devalued currency. A country with a more devalued currency is expected to be more aligned with China than not. The definitions and measurements of key variables are provied in table 1. Table 2 presents the descriptive statistics.



```{r table_summary, message=FALSE, warning=FALSE, echo=FALSE}
library(targets)
library(skimr)
library(knitr)
library(gtsummary)
library(tidyverse)
df <- tar_read(final_df)  %>%
  dplyr::select(-c("us_agree", "china_agree", "year"))

# my_skim <- skim_with(numeric = sfl(n_missing = NULL, complete_rate = NULL, iqr = IQR, p25 = NULL, p75 = NULL, hist=NULL), 
#                      character= sfl(n_missing = NULL, complete_rate = NULL, empty=NULL, whitespace=NULL))
# 
# df %>%
#   my_skim() 
# df %>%
#   my_skim(numeric = sfl(hist = NULL))

df %>%
  dplyr::select(where(is.numeric)) %>%
  tbl_summary(label= list(pop = "Population, millions",
                          gdp_const = "GDP in 2017 USD PPP, millions"),
              type = all_continuous() ~ "continuous2",
              statistic = all_continuous() ~ c("{mean} ({sd})", "{min}, {max}"),
              digits = all_continuous() ~ 2) 


```




```{r plot, message=FALSE, warning=FALSE, echo=FALSE, fig.cap= "caption", fig.pos="h"}
# library(ggplot2)
tar_read(plot)
```

# Descriptive Analysis

In figure bellow, we plot the evolution of UNGA's ideal point of US, Brazil and China over time. As we can see, Brazil is closer to China than to the US since the beginning of the series of my data, in 1990 upt to 2022, during Bolsonaro's government. Even thoud there is a distancing of Brasil under Bolsonaro's term, it is still closer to the China than to the US. This is a testimonement of how much Brazil is structury closer to China at the UNGA.

It also shows that Brazil is moving more over time than China, which is mostly stable over time. Thus, that is an indication that is Brazil who is moving closer to China than the opposite, as expected, given the differential of power and importance of both Countries and the dependence between them. 


```{r plot_unga1, message=FALSE, warning=FALSE, echo=FALSE, fig.cap= "caption", fig.pos="h"}
# library(ggplot2)
tar_read(plot_ideal)
```

Having established that is mostly Brazil moving around than China, we can use as our main variable the absolute distance between Brazil and China's ideal point, with the understand that, a part from the beginning of the nineties, Brazil is almost always closer to the center than China and that Brazil is the one moving around.

```{r plot_unga2, message=FALSE, warning=FALSE, echo=FALSE, fig.cap= "caption", fig.pos="h"}
# library(ggplot2)
tar_read(plot_ideal_distance)
```

The question one wonders after seeing the graphs is: why Brazil was getting closer over time? the main hypothesis discussed in the literature is the gravitation pull of the China economy, peharps as shown in the percentage of exports of Brazil that wen to China in the period considered. The figure bellow shows that around year 2000, the percentage of exports destined to China increase at a roughly constant rate, in a very steep variation, which does not exactly fit the political alignment as measured by UNGA ideal points, although there is some positive correlation.

```{r plot_trade, message=FALSE, warning=FALSE, echo=FALSE, fig.cap= "caption", fig.pos="h"}
# library(ggplot2)
tar_read(trade_plot)
```

# Empirical Results

The graph below presents the evolution of political alignment between Brazil and China at UNGA and a donor pool of countries before and after China becoming the main trade partner of Brazil. The counterfactual estimate from the donor pool allows us to estimate the causal effect of the treatment on similarity of vote at UNGA. The estimate is -.27 points, which means that becoming the main trade partner implied a decrease of .27 points in the absolute distance between the ideal point of Brazil and China after 2006. The result is statistically significant at the 5% level, allowing us to reject the null hypothesis of no increase in the distance between the ideal points of both countries.

```{r plot_sdis, message=FALSE, warning=FALSE, echo=FALSE, fig.cap= "caption", fig.pos="h"}
tar_read(plot_trends)
```

```{r plot_parallel, message=FALSE, warning=FALSE, echo=FALSE, fig.cap= "caption", fig.pos="h"}
tar_read(plot_dif)
```

```{r plot_weigths, message=FALSE, warning=FALSE, echo=FALSE, fig.cap= "caption", fig.pos="h"}
tar_read(plot_weights)
```
# Robustness checks

We already have reasons to not use the SCM method. However, since it is a more established methdoology, it is useful to compare both SCM and SDiD models. As table 3 bellow shows, the results are qualitatively similar. the result of the SCM is not significant, but this is expeced, since the variance is inflated and, with the low number of data point, it is hard to find a significant result. What is important is that the results are qualitatively similar in both methods.


# Conclusion