---
title: "The Foreign Policy Impact of Trade-Based Status Gains: When China Overtakes the US as Top Export Destination"
author: "Manoel Galdino (University of São Paulo; mgaldino@usp.br)"
date: "`r Sys.Date()`"
output:
  bookdown::pdf_document2:
    number_sections: TRUE
    toc: FALSE
    latex_engine: xelatex

header-includes:
- \usepackage{setspace}\doublespacing
- \setstretch{1.5}
- \usepackage{parskip}
- \setlength{\parskip}{6pt}
- \setlength{\parindent}{0pt}
- \usepackage{float}
- \usepackage{longtable}
- \newcommand{\argmin}{\mathop{\mathrm{argmin}}\limits}
fontsize: 12pt
abstract: "\\singlespacing  How do trade-based status gains affect UNGA voting convergence toward China? I argue that rank reversals in trade hierarchies create politically usable status cues: when China becomes a country's top export destination and remains there, policymakers can more easily justify selective diplomatic adjustment toward Beijing. I test the argument in Brazil, where China overtook the United States as the largest export destination in 2009, using Synthetic Difference-in-Differences on UNGA ideal-point distance. The estimated Brazil-minus-synthetic post-2009 gap is -0.26 ideal-point units, about 41 percent of Brazil's pre-treatment mean, and vote-level diagnostics show that the post-2009 movement is clearest in human-rights resolutions where China and the United States diverged. Brazilian newspaper evidence shows increased China-related trade salience after the rank reversal. Cross-country estimates from a goods-only interactive fixed-effects panel model show a compatible directional pattern among countries where China persistently becomes the top goods export destination. The evidence suggests that discrete and durable trade-status milestones can shape observable multilateral alignment beyond smooth trade growth."
geometry: margin=2.5cm
spacing: double
mainfont: Times New Roman
bibliography: "synth-trade-china.bib"
link-citations: true
---
\pagebreak

```{r setup, include = FALSE}
library(tidyverse)
library(targets)
knitr::opts_chunk$set(echo = FALSE)
source(here::here("scripts", "functions.R"))

caption_note <- function(unit = NULL, se = NULL, cluster = NULL, window = NULL) {
  fields <- c(
    if (!is.null(unit) && nzchar(unit)) unit,
    if (!is.null(se) && nzchar(se)) paste0("SE: ", se),
    if (!is.null(cluster) && nzchar(cluster)) paste0("Cluster: ", cluster),
    if (!is.null(window) && nzchar(window)) paste0("Window: ", window)
  )
  if (length(fields) == 0) {
    return("")
  }
  paste0("Note: ", paste(fields, collapse = ". "), ".")
}

resolve_existing_file <- function(candidates) {
  hits <- candidates[file.exists(candidates)]
  if (length(hits) == 0) {
    stop("Missing file. Tried: ", paste(candidates, collapse = " | "))
  }
  hits[[1]]
}

word_count_text <- function(input = knitr::current_input(), subtract_n = 20L) {
  wc_raw <- system2("wc", c("-w", input), stdout = TRUE)
  wc_n <- as.integer(strsplit(trimws(wc_raw), "\\s+")[[1]][1])
  format(wc_n - subtract_n, scientific = FALSE, big.mark = ",", trim = TRUE)
}
```


Word count: `r word_count_text()`

```{r basic-num, include = FALSE}
estimate <- as.numeric(tar_read(synth_fit)[1])
se_estimate <- as.numeric(tar_read(se_synth))
ci_low <- estimate - 1.96 * se_estimate
ci_high <- estimate + 1.96 * se_estimate
p_estimate <- 2 * pnorm(-abs(estimate / se_estimate))
mean_br <- tar_read(synth_data) %>%
  dplyr::filter(iso3c == "BRA", year < 2009) %>%
  dplyr::summarise(mean_abs = mean(abs_distance_china, na.rm = TRUE)) %>%
  dplyr::pull(mean_abs)
perc_change <- abs(estimate) / mean_br * 100
human_rights_vote_change <- tar_read(brazil_china_unvotes_resolution_2005_2012) %>%
  dplyr::filter(issue_family == "Human rights") %>%
  dplyr::mutate(period = dplyr::if_else(year < 2009, "pre", "post")) %>%
  dplyr::summarise(
    identical_vote_share = 100 * mean(identical_vote, na.rm = TRUE),
    .by = period
  ) %>%
  tidyr::pivot_wider(names_from = period, values_from = identical_vote_share) %>%
  dplyr::mutate(change = post - pre) %>%
  dplyr::pull(change)
sample_size_countries <- tar_read(synth_data) %>%
  dplyr::summarise(num_countries = dplyr::n_distinct(iso3c)) %>%
  dplyr::pull(num_countries)
```



# Introduction

“China overtakes US as Brazil’s top trade partner”. That was a headline in many Brazilian newspapers in 2009 [@rodrigues2009]. The BBC ran a similar headline in 2021, when China surpassed the United States as the EU’s biggest trading partner [@bbc2021]. During a high-profile trip to Beijing with an unusually large business delegation, President Lula da Silva invoked the milestone by noting that China was Brazil's largest export destination [@lula2009].

Why do media outlets and political leaders dwell on discrete commercial milestones when the underlying flows are measured continuously? Export statistics are updated monthly, and journalists, diplomats, firms, and policymakers follow them long before one partner crosses a formal threshold. The puzzle, then, is not whether actors possess information about trade volumes. It is why a categorical change—China becomes number one—can matter politically despite adding little new information about the economic trend itself.

Answering this question rigorously is difficult because trade hierarchies are not randomly assigned. Existing work on economic statecraft and foreign-policy alignment usually treats economic exposure as continuous: as trade, lending, investment, aid, or coercive capacity accumulate, alignment is expected to shift through material leverage, changing interests, soft power, or domestic pressure [@flores-macias_kreps2013; @urdinez_etal2016; @kastner_pearson2021]. A continuous-interdependence account would therefore expect foreign-policy alignment to move gradually with trade exposure, leverage, and economic dependence rather than at a discrete rank threshold. Domestic-political accounts, by contrast, would attribute foreign-policy change in cases such as Brazil to leadership ideology, autonomy, South-South cooperation, or broader foreign-policy reorientation [@neto2011; @neto_malamud2015; @rodrigues2019measuring; @schenoniMythsMultipolaritySources2022]. A further selection concern is that politically aligned countries may deepen economic ties with one another, making China’s rise in the trade hierarchy a consequence rather than a cause of prior diplomatic affinity [@gowa1995; @davis_pratt2021; @struver2016]. If alignment reflects only accumulated exposure, domestic-political reorientation, or prior diplomatic affinity, however, no specific symbolic milestone such as becoming a top trade partner should matter.

I argue that such milestones matter because they mark trade-based status gains. The argument builds on research that treats status as socially recognized standing within international hierarchies, clubs, and local status orders [@paul_larson_wohlforth2014; @wolfRespectDisrespect2011; @duque_2018; @renshon2017; @gotz_2021; @macdonald_parent2021; @roren_2024; @roren2025_power_recognition; @wolfTakingInteractionSeriously2019]. By trade-based status, I mean the recognized standing of a foreign power within a country’s hierarchy of economically consequential relationships. Trade volume is continuous, but status is categorical: a partner can move from being important to being understood as central. Once that reclassification becomes public, the relationship is no longer merely larger than before. It becomes easier for political actors to invoke as a national priority, to defend as strategically consequential, and to use as a justification for diplomatic adjustment.

Much of the China and rising-powers branch of this literature asks how rising states seek recognition, major-power status, or institutional standing, and how misrecognition can produce contestation or revisionism [@deng2008_status; @larson_shevchenko2010_status_seekers; @ward2017_status_challenge; @murray2018_recognition; @mukherjee2022_ascending]. This argument relocates status from the ambitions of the rising power to the recognition practices of states. China’s trade-based status gain occurs when another country publicly recategorizes China within its own hierarchy of external economic relationships -- from an important partner to a central one. The mechanism is therefore status-theoretic, with Brazil as the audience and status-conferring actor.

Brazil is the strongest case for the theory. China displaced the United States, a politically salient incumbent and hegemonic benchmark; the rank reversal was publicly visible; and the Brazilian press allows the attention implication to be measured. The case is also substantively important because Lula-era Brazilian foreign policy is often described through autonomy, South-South activism, multipolarity, and Brazil's triangular position between the United States and China [@neto2011; @neto_malamud2015; @rodrigues2019measuring; @schenoniMythsMultipolaritySources2022; @guilhon2014; @schenoni_leiva2021; @brun_covarrubias2024; @coutinho_rodriguez2024]. 

To estimate the Brazilian foreign-policy alignment pattern, I leverage the discreteness in China's status change to largest Brazilian export destination using synthetic difference-in-differences (SDiD) on annual UNGA ideal-point distance to China. The estimate is an average post-2009 gap. Relative to Brazil's pre-treatment mean distance, the estimated reduction is `r sprintf('%.1f', perc_change)` percent. UNGA voting is a standard measure to study alignment with China and other powers, linking votes to bilateral ties, ideology, sanctions, aid, Belt and Road membership, arms transfers, or Chinese economic returns [@potrafke2009; @dreher_jensen2013; @struver2016; @yan_zhou2021; @garcia_herrero_etal2024; @steinert_weyrauch2024; @he_zheng_chen2025; @lektzian_biglaiser2023].

To assess whether the pattern is specific to Brazil, I extend the analysis to a cross-country panel in which China enters and holds the top goods export-destination position. The cross-country estimates are smaller in magnitude, consistent with a more heterogeneous sample, and point in the same direction as the Brazilian case. The panel evidence is compatible with the rank-threshold argument beyond the theory-generating case, while the Brazilian evidence documents the mechanism in detail.

The paper’s contribution is to recast economic influence as a status process. Existing work usually treats trade exposure as a continuous source of leverage, interests, or domestic pressure. I show that accumulated trade growth can become politically consequential when it crosses a durable, publicly legible rank threshold: when China becomes and remains a country’s largest export destination, the new status condition creates a persistent basis for selective diplomatic adjustment. In Brazil, China’s rise to the top export-destination position made an already expanding commercial relationship politically usable as evidence of China’s new centrality. That status gain helps explain why convergence appears selectively in UNGA voting. More broadly, the paper shows that status gains can emerge from economic hierarchies and shape observable diplomatic behavior.


The next section develops the status-threshold argument and derives observable implications. I then present the Brazil SDiD design and timing diagnostics, followed by issue-area and media evidence on the mechanism. The final empirical section evaluates whether durable China-top goods-export status is associated with similar movement in a broader cross-country panel.

# Theory and hypotheses

A trade-based status shift differs from smooth growth in economic interdependence because it changes the category through which a foreign power is understood. Actors can observe trade shares, but political attention rarely follows continuous movements one percentage point at a time. It also organizes around coarse public labels: largest export market, number-one partner, top-ranked export destination. These labels simplify complex economic relationships and help journalists, politicians, organized interests, and diplomats coordinate attention [@bordalo_etal2022; @conlon2023; @enke2024; @graeber_etal2025]. The number-one threshold matters because it can reclassify a rising power from one important commercial partner among several into the country's central economic partner. In this sense, export-destination rank operationalizes the trade-based status shift.

The cue should matter most when the new status also displaces a politically prominent incumbent. If a rising major power overtakes a minor partner, the event may remain a commercial fact with limited political resonance. If it overtakes the hegemonic incumbent, the same threshold takes on broader political meaning. It can be read as a visible loss of standing for the incumbent and as a public marker of the challenger's rise. Brazil falls squarely inside this high-salience condition: China became Brazil's largest export destination in 2009, displaced the United States, and contemporary coverage frequently described the event in broader top-partner language.

The mechanism has two steps. First, the status cue should increase attention. Media coverage should make the challenger and the bilateral trade relationship more visible after the threshold is crossed. Second, the cue should affect policy through justification. Once that power is publicly understood as a central economic partner, executives and diplomats can frame limited movement toward it as pragmatic recognition of the country's economic hierarchy rather than as ideological concession or opportunistic bandwagoning. This step can be gradual because UNGA positions are embedded in prior votes, bureaucratic routines, coalition commitments, and reputational constraints. The Brazilian SDiD design estimates the combined reduced-form effect of this status cue on UNGA distance to China.

The justification mechanism also implies a durability condition. The rank reversal is not expected to reshape foreign policy through a single decision at the moment China becomes number one. Rather, it creates a public category that officials can invoke repeatedly across a sequence of selective adjustments. When China remains the largest export destination, presidents, ministers, diplomats, and organized interests can treat its centrality as a stable fact of the national economic hierarchy and use it to justify incremental diplomatic movement. A one-year China-top episode may attract attention, but it is a weaker treatment regime: it does not provide a durable basis for repeated public justification, bureaucratic recalibration, or cumulative foreign-policy adaptation.

## From Attention to Selective Diplomatic Alignment

Trade-based status changes the terms on which selective accommodation can be justified. Some foreign-policy arenas are especially suitable for this kind of adjustment: they are visible to foreign governments and diplomatic audiences, materially less costly than treaties, military commitments, sanctions, or trade policy, and still capable of generating reputational exposure. In such arenas, governments can signal movement toward a newly central partner while preserving flexibility.

When a foreign power's rise to the top export-destination position is publicly recognized as evidence of centrality in the country's trade hierarchy, policymakers can frame limited convergence with that power as pragmatic recognition of economic importance rather than as ideological concession or opportunistic bandwagoning. The status gain can therefore lower the reputational cost of selected movement while increasing its diplomatic value.

The domestic pathway works through justification. Media coverage need not cause individual foreign-policy positions, and business groups need not lobby over specific diplomatic decisions. Public recognition of the partner’s new standing changes what diplomats and executives can plausibly defend. Adjustment toward that partner becomes easier to present as alignment with national economic interests, especially in diplomatic settings where positions are visible internationally but relatively low-cost domestically.

This logic implies selective convergence. The mechanism should be most informative in issue areas where the country and the newly elevated partner previously differed and where movement is not costless. In domains tied to prior reputational commitments, adjustment toward the newly central partner requires a stronger rationale than in areas where the two countries already agree. Trade-based status supplies such a rationale by making the partner’s economic centrality publicly available as a justification for diplomatic accommodation.

The main theoretical expectation is therefore that when a rising major power becomes a country’s largest export destination and that milestone becomes a public trade-status cue, the country's foreign policy should move closer to that power relative to its counterfactual path. The effect should be easiest to observe when the status gain displaces a politically prominent incumbent, especially a hegemonic benchmark such as the United States. The theory also implies an attention pattern: media coverage of the rising power and of the bilateral economic relationship should increase around the status shift.

# Data and design

Empirically, the treatment is China becoming a country's largest export destination. I use "largest trading partner" only as shorthand for the publicly salient Brazilian event when discussing contemporary framing. The Brazil SDiD and cross-country panel define treatment using largest-export-destination status.

This distinction matters because public rank language is often broader than the operational treatment. Officials and journalists may describe China as the largest trading partner, main commercial partner, or top export market, while the design uses a narrower export-destination rank. I therefore treat public-source evidence as mechanism evidence only when the metric is clear. A supplemental source audit in the appendix uses displaced-incumbent coverage as a recoverability diagnostic: cases where neither China nor the former top export destination appears in local or official sources are weakly observed, not, by themselves, evidence of low salience. Australia illustrates the opposite caveat. The *Australian Financial Review* reported in 2009 that China was Australia's largest aggregate trading partner, combining exports and imports of goods and services, before Australia meets the paper's stricter goods-export-destination treatment criterion. I code that source as evidence that a broad China-as-number-one cue was already publicly salient, not as newspaper uptake of the paper's stricter export-destination treatment.

```{r basic-data, message=FALSE, warning=FALSE, echo=FALSE}
library(targets)
library(tidyverse)
df <- tar_read(synth_data)
num_countries <- df %>%
  dplyr::summarise(num_countries = dplyr::n_distinct(iso3c)) %>%
  dplyr::pull(num_countries)

num_years <- df %>%
  dplyr::filter(iso3c == "BRA", year < 2016) %>%
  dplyr::summarise(num_years = dplyr::n()) %>%
  dplyr::pull(num_years)

min_year <- df %>%
  dplyr::summarise(min_year = min(year)) %>%
  dplyr::pull(min_year)

max_year <- df %>%
  dplyr::filter(year < 2016) %>%
  dplyr::summarise(max_year = max(year)) %>%
  dplyr::pull(max_year)
```

The primary outcome is absolute UNGA ideal-point distance to China, based on the annual ideal-point estimates of @bailey_etal2017. Lower values indicate convergence toward China. This measure is preferable to raw annual agreement shares for the main design because it is less sensitive to changes in the UNGA agenda over time. I use resolution-level agreement diagnostics only to interpret where convergence appears strongest.

For Brazil, the treatment indicator equals one from 2009 onward, the first year China became Brazil's largest export destination. SDiD estimates Brazil's average post-2009 gap relative to a weighted synthetic counterfactual, using pre-treatment information to construct unit and time weights [@arkhangelsky_etal2021]. The design is reduced-form. Its credibility rests on pre-treatment fit, rank-versus-volume timing tests, donor-pool sensitivity, and outcome robustness.

The SDiD panel combines UNGA ideal points with covariates for trade exposure, macroeconomic conditions, power, geographic distance, political institutions, and trade-agreement ties. The complete panel covers `r num_countries` countries and `r num_years` Brazil years for the main SDiD window (`r min_year`-`r max_year`). Trade data come from the International Trade and Production Database for Estimation [@borchert_etal2021]. The ITPD-E `trade` variable is measured in millions of current US dollars, and export-destination ranks are computed from exporter-importer flows.

The Brazil variables enter the design in three groups. The outcome is Brazil's annual absolute UNGA ideal-point distance to China; the treatment is an indicator for years after China becomes Brazil's largest export destination. The covariates measure trade exposure to China and the United States, relative power, geographic distance to Washington, leader ideology, macroeconomic stress, regime institutions, and U.S. trade-agreement ties. The cross-country panel uses the same outcome, but constructs treatment separately from goods-only export-destination ranks and a status-current risk set. Table \@ref(tab:table-summary-main) summarizes the country-year panel used to estimate Brazil's synthetic counterfactual. The large number of observations reflects the donor pool and pre/post-treatment country-year structure of the SDiD design.

```{r table-summary-main, message=FALSE, warning=FALSE, echo=FALSE, results='asis'}
summary_vars <- tar_read(synth_data) %>%
  dplyr::filter(year >= min_year, year <= max_year) %>%
  dplyr::select(
    `Treatment` = treatment,
    `UNGA distance to China` = abs_distance_china,
    `UNGA distance to US` = abs_distance_usa,
    `Geographic distance to US` = distance_us,
    `Trade share with China` = perc_trade_with_china,
    `Trade share with US` = perc_trade_with_us,
    `Power index` = gpi,
    `Power gap to US` = us_power_gap,
    `GDP per capita` = pci_cur,
    `Exchange rate` = exachange_rate,
    `HoG left ideology` = hog_left,
    `Current account` = CA_GDP,
    `Government deficit` = govdef_GDP,
    `Parliamentary system` = dplyr::any_of("inst_parliamentary"),
    `Military executive` = dplyr::any_of("inst_military_exec"),
    `US trade agreement` = dplyr::any_of("us_trade_agreement")
  )

summary_tbl <- purrr::map_dfr(names(summary_vars), function(var_name) {
  x <- summary_vars[[var_name]]
  tibble::tibble(
    Variable = var_name,
    N = sum(!is.na(x)),
    Mean = mean(x, na.rm = TRUE),
    SD = stats::sd(x, na.rm = TRUE),
    Min = min(x, na.rm = TRUE),
    Max = max(x, na.rm = TRUE)
  )
}) %>%
  dplyr::mutate(
    dplyr::across(c(Mean, SD, Min, Max), ~sprintf("%.2f", .x))
  )

knitr::kable(
  summary_tbl,
  format = "latex",
  booktabs = TRUE,
  caption = "Descriptive statistics for the country-year panel used in the Brazil SDiD design.",
  align = "lccccc"
) %>%
  kableExtra::kable_styling(
    latex_options = c("HOLD_position", "scale_down"),
    font_size = 8
  ) %>%
  kableExtra::footnote(
    general = paste0(
      sub(
        "^Note: ",
        "",
        caption_note(
          unit = "Country-year variables from the SDiD analytic panel; continuous covariates are rescaled in the estimation data, UNGA distances are ideal-point distances, and treatment is binary",
          window = sprintf("%d-%d SDiD analytic panel", min_year, max_year)
        )
      ),
      " Observations include Brazil and donor countries used to construct the synthetic counterfactual. The treatment mean is small because only Brazil is treated after 2009."
    ),
    general_title = "Note: ",
    escape = FALSE,
    threeparttable = TRUE
  )
```

For the cross-country evidence beyond Brazil, the main sample is restricted to durable China-top export-destination periods measured with goods exports only. The treatment equals one only in country-years where China is the largest goods export destination and that China-top period lasts at least five observed years in the joint trade-rank and UNGA-outcome panel. Treated countries must first enter this status in or after 2000, after an observed prior year in which China was not number one. Countries that never reach China-top goods-export status remain controls. For treated countries, the main risk set keeps clean pre-entry non-China-top years and qualifying treated years, while excluding post-exit off-status years and short China-top episodes. This operationalization matches the salience logic more closely than either a one-year entry definition or an absorbing treatment rule: the theory expects political visibility, bureaucratic learning, and diplomatic adaptation to be clearer when the new trade hierarchy persists and is currently true. I compute partner ranks from ITPD-E goods sectors only, excluding services before ranks are constructed. For this panel, I use the `fect` counterfactual estimator with interactive fixed effects, which allows countries to load differently on common latent shocks [@liu2024practical]. I refer to this as the IFE specification below.

## Identification strategy

The empirical problem is that China's rise in a country's trade hierarchy is not randomly assigned. Countries may move closer to China because trade exposure accumulates continuously, because domestic political coalitions or foreign-policy doctrines change, or because governments already inclined toward China deepen commercial ties before the rank reversal occurs. The design therefore asks the identifying question: whether foreign-policy alignment changes around the discrete status threshold after accounting for pre-treatment trajectories and observed sources of economic, geopolitical, and domestic-political orientation.

In the Brazil design, SDiD constructs a weighted counterfactual from pre-treatment information. The covariates are not included simply to improve fit. They proxy observable rival explanations and improve balance on measured trade, geopolitical, and domestic-political factors. Trade shares with China and the United States measure smooth changes in economic exposure; relative power, geographic distance to Washington, and U.S. trade-agreement ties proxy structural exposure to the incumbent power; leader ideology, regime institutions, and macroeconomic stress proxy domestic-political reorientation and contemporaneous shocks. These covariates do not eliminate unobserved domestic coalitions, doctrines, or shocks. The demand-shock diagnostics further add smooth China-share transformations, export-destination concentration, pre-2009 primary-goods export exposure, and 2008-2009 exposure interactions. The rank-versus-volume timing tests then ask whether earlier trade growth or lower-rank promotions generate the same estimated convergence as the top-rank reversal.

The cross-country panel provides a second check against Brazil-specific timing explanations. Treated countries enter durable China-top goods-export status in different years, so a Brazil-specific 2009 shock cannot mechanically generate the pooled result. Interactive fixed effects allow countries to load differently on common latent shocks, while the appendix reports duration-threshold and risk-set robustness checks.

# Brazil SDiD evidence

The Brazilian analysis estimates the average post-2009 reduced-form effect of China's entry into the top export-destination position on Brazil's UNGA ideal-point distance to China. It asks whether Brazil moved closer to China than a synthetic counterfactual after the 2009 rank reversal.

Figure \@ref(fig:brazil-rank-volume) shows the timing of the rank threshold. China's export share rose before 2009, but China became the top-ranked export destination only in 2009. The signed margin over the relevant competitor also changes sign at that threshold. The SDiD covariate matrix includes trade shares with China and the United States to account for smooth trade exposure. The rank margin is part of the threshold event itself, so it is used as a timing diagnostic rather than as a separate control.

```{r brazil-rank-volume, message=FALSE, warning=FALSE, echo=FALSE, fig.cap=paste0("Brazil rank-versus-volume diagnostic for China's export-destination status. Panels show China's export share, ordinal rank, and signed margin over the relevant competitor. The dashed line marks the 2009 rank reversal. ", caption_note(unit = "Export share in percent; ordinal rank; export-value margin in current USD billions", window = "2000-2012 annual Brazil series")), fig.pos="H", fig.width=7.2, fig.height=7.2}
tar_read(goal3_brazil_rank_volume_plot)
```

Figure \@ref(fig:plot-sdid) presents the main SDiD fit. As shown in the Appendix comparison between SCM and SDiD, SDiD is more flexible than SCM because it includes unit effects and therefore permits a stable pre-treatment level difference between Brazil and its synthetic comparison. I therefore report an intercept-adjusted RMSPE as the pre-treatment fit metric: the Brazil-minus-synthetic gap is centered by its pre-2009 mean before the RMSPE is computed, making the diagnostic comparable in spirit to the usual SCM RMSPE while matching the SDiD estimating equation. Balance on trade exposure to China remains incomplete, so I evaluate the estimate through adjusted fit, balance, placebo ranks, donor influence, donor exposure, and window sensitivity rather than through the visual comparison alone. The post-2009 gap is negative, indicating convergence toward China. The point estimate is `r sprintf('%1.2f', estimate)` ideal-point units, with placebo-based SE `r sprintf('%1.2f', se_estimate)` and a normal-approximation p-value of `r sprintf('%.3f', p_estimate)`; the single-treated-unit placebo ranks in Table \@ref(tab:brazil-sdid-diagnostic-summary) provide the more appropriate inferential benchmark. Relative to Brazil's pre-treatment mean distance (`r sprintf('%.3f', mean_br)`), this corresponds to a `r sprintf('%.1f', perc_change)` percent reduction. The estimate summarizes the average post-treatment gap. Appendix Figure \@ref(fig:appendix-brazil-china-ideal-points) plots the underlying Brazil and China ideal-point series separately; China is comparatively stable, so the post-2009 reduction in absolute distance is driven mainly by movement in Brazil's position rather than by China moving toward Brazil.

```{r plot-sdid, message=FALSE, warning=FALSE, echo=FALSE, fig.cap=paste0("SDiD fit for Brazil and synthetic control with time weights. Lower values indicate convergence toward China; the post-treatment gap is an average post-2009 reduced-form effect. ", caption_note(unit = "Absolute UNGA ideal-point distance to China", se = "placebo-based SE with 1,000 replications; p-values reported in the table", window = "Brazil annual series")), fig.pos="H"}
tar_read(plot_trend)
```

```{r brazil-sdid-diagnostic-inputs, include=FALSE}
sdid_diagnostics <- tar_read(brazil_sdid_diagnostics_bundle)
sdid_unit_weights <- sdid_diagnostics$unit_weights
sdid_time_weights <- sdid_diagnostics$time_weights
sdid_fit_summary <- sdid_diagnostics$fit_summary
sdid_balance <- sdid_diagnostics$balance
sdid_placebo_rank <- sdid_diagnostics$placebo_rank
sdid_donor_sensitivity <- sdid_diagnostics$donor_sensitivity
sdid_window_sensitivity <- sdid_diagnostics$window_sensitivity
sdid_donor_china_exposure <- sdid_diagnostics$donor_china_exposure
sdid_high_weight_exposure_summary <- sdid_diagnostics$high_weight_exposure_summary
china_demand_timing <- tar_read(goal3_brazil_placebo_rank_volume_tests)

sdid_top_donors_text <- sdid_unit_weights %>%
  dplyr::slice_head(n = 5) %>%
  dplyr::mutate(label = paste0(iso3c, " ", sprintf("%.1f%%", 100 * unit_weight))) %>%
  dplyr::pull(label) %>%
  paste(collapse = ", ")

sdid_high_time_weights_text <- sdid_time_weights %>%
  dplyr::arrange(dplyr::desc(time_weight)) %>%
  dplyr::slice_head(n = 4) %>%
  dplyr::mutate(label = paste0(year, " ", sprintf("%.1f%%", 100 * time_weight))) %>%
  dplyr::pull(label) %>%
  paste(collapse = ", ")

sdid_one_sided_placebo <- sdid_placebo_rank %>%
  dplyr::filter(estimand == "One-sided negative effect") %>%
  dplyr::slice_head(n = 1)
sdid_fit_filtered_placebo <- sdid_placebo_rank %>%
  dplyr::filter(estimand == "One-sided negative effect, fit-filtered") %>%
  dplyr::slice_head(n = 1)
sdid_rmspe_ratio_rank <- sdid_placebo_rank %>%
  dplyr::filter(estimand == "Intercept-adjusted RMSPE ratio rank") %>%
  dplyr::slice_head(n = 1)
sdid_outcome_balance <- sdid_balance %>%
  dplyr::filter(variable == "abs_distance_china") %>%
  dplyr::slice_head(n = 1)
sdid_china_share_balance <- sdid_balance %>%
  dplyr::filter(variable == "perc_trade_with_china") %>%
  dplyr::slice_head(n = 1)
sdid_max_single_donor_change <- sdid_donor_sensitivity %>%
  dplyr::filter(n_removed_donors == 1) %>%
  dplyr::summarise(max_change = max(abs(percent_change_vs_baseline), na.rm = TRUE)) %>%
  dplyr::pull(max_change)
sdid_window_summary <- sdid_window_sensitivity %>%
  dplyr::summarise(
    n_windows = dplyr::n(),
    n_negative = sum(estimate < 0, na.rm = TRUE),
    min_estimate = min(estimate, na.rm = TRUE),
    max_estimate = max(estimate, na.rm = TRUE),
    max_one_sided_p = max(placebo_p_one_sided_negative, na.rm = TRUE)
  )
sdid_rank2_placebo <- china_demand_timing %>%
  dplyr::filter(nominal_treatment_year == 2004L) %>%
  dplyr::slice_head(n = 1)
sdid_actual_timing <- china_demand_timing %>%
  dplyr::filter(nominal_treatment_year == 2009L) %>%
  dplyr::slice_head(n = 1)
```

Table \@ref(tab:brazil-sdid-diagnostic-summary) summarizes the diagnostic package behind the Brazil SDiD estimate. The donor vector is diffuse rather than driven by one country: the top five donors are `r sdid_top_donors_text`, and the complete vector appears in Appendix Table \@ref(tab:sdid-unit-weights-complete). This donor pool is a statistical counterfactual supported by fit, placebo, and sensitivity diagnostics, not a claim that each high-weight donor is substantively similar to Brazil. The time weights concentrate most heavily on 2007-2008, but the appendix reports all pre-treatment weights. These diagnostics do not eliminate all possible 2009 confounding: the global financial crisis, the BRICS/G20 moment, and commodity-cycle politics remain contemporaneous with the Brazilian threshold. They do, however, weaken the narrower China-demand explanation because the point estimate remains stable after adding smooth trade-share, concentration, predetermined commodity-exposure, donor-influence, and window diagnostics, while the 2004 China number-two placebo is much smaller than the 2009 rank-one onset.

```{r brazil-sdid-diagnostic-summary, message=FALSE, warning=FALSE, echo=FALSE, results='asis'}
sdid_diag_summary <- tibble::tribble(
  ~Diagnostic, ~Result, ~Interpretation,
  "Complete donor and time weights",
  paste0(
    "N = ", nrow(sdid_unit_weights), " donors; top five: ",
    sdid_top_donors_text, ". Highest time weights: ",
    sdid_high_time_weights_text, "."
  ),
  "Synthetic comparison is not a single-donor result; late pre-treatment years receive most time weight.",
  "Fit and balance",
  paste0(
    "Intercept-adjusted pre-treatment RMSPE = ", sprintf("%.3f", sdid_fit_summary$rmspe_pre[1]),
    "; post/pre adjusted RMSPE ratio = ", sprintf("%.3f", sdid_fit_summary$rmspe_ratio[1]),
    ". Outcome pre-mean: Brazil ", sprintf("%.3f", sdid_outcome_balance$brazil_pre_mean),
    ", synthetic ", sprintf("%.3f", sdid_outcome_balance$synthetic_pre_mean),
    "; China-share standardized difference = ",
    sprintf("%.3f", sdid_china_share_balance$standardized_difference), "."
  ),
  "The pre-period fit is useful but imperfect, so placebo and sensitivity checks are central to interpretation.",
  "Placebo ranks",
  paste0(
    "One-sided negative rank = ", sprintf("%.0f", sdid_one_sided_placebo$rank),
    "/", sprintf("%.0f", sdid_one_sided_placebo$denominator),
    " (p = ", sprintf("%.3f", sdid_one_sided_placebo$rank_based_p_value),
    "); fit-filtered p = ", sprintf("%.3f", sdid_fit_filtered_placebo$rank_based_p_value),
    "; adjusted RMSPE-ratio rank = ", sprintf("%.0f", sdid_rmspe_ratio_rank$rank),
    "/", sprintf("%.0f", sdid_rmspe_ratio_rank$denominator), "."
  ),
  "The Brazil gap is large relative to placebo assignments, especially when scaled by intercept-adjusted pre-treatment fit.",
  "Donor influence and donor exposure",
  paste0(
    "Largest leave-one-donor change = ",
    sprintf("%.1f%%", sdid_max_single_donor_change),
    "; top-ten donor deletion estimate = ",
    sprintf("%.3f", sdid_donor_sensitivity$estimate[sdid_donor_sensitivity$specification == "Drop top 10 donors by weight"]),
    ". High-weight donors where China ranked in the top two after 2009: ",
    sdid_high_weight_exposure_summary$n_high_weight_top_two_china_post,
    " (", sprintf("%.1f%%", 100 * sdid_high_weight_exposure_summary$weighted_share_high_weight_top_two_china_post),
    " of total weight)."
  ),
  "The sign does not hinge on one high-weight donor; donor exposure to China is a remaining contamination risk, not evidence that the result is driven by one exposed donor.",
  "Window and threshold sensitivity",
  paste0(
    sdid_window_summary$n_negative, "/", sdid_window_summary$n_windows,
    " window checks are negative; estimates range from ",
    sprintf("%.3f", sdid_window_summary$min_estimate), " to ",
    sprintf("%.3f", sdid_window_summary$max_estimate),
    ". China number-two placebo in 2004 = ",
    sprintf("%.3f", sdid_rank2_placebo$estimate),
    " vs. actual 2009 = ",
    sprintf("%.3f", sdid_actual_timing$estimate), "."
  ),
  "The pattern is consistent with a rank-threshold component, while remaining cautious about Brazil-specific shocks in 2009."
)

knitr::kable(
  sdid_diag_summary,
  format = "latex",
  booktabs = TRUE,
  caption = "Compact diagnostic summary for the Brazil SDiD estimate.",
  align = "lll"
) %>%
  kableExtra::kable_styling(
    latex_options = c("HOLD_position", "scale_down"),
    font_size = 8
  ) %>%
  kableExtra::footnote(
    general = paste0(
      "All Brazil SDiD diagnostics and timing-placebo values in this table are produced by the targets pipeline. ",
      "RMSPE diagnostics are intercept-adjusted: the Brazil-minus-synthetic gap is centered by its pre-treatment mean before the RMSPE is computed, because SDiD permits stable level differences whereas pure SCM does not. ",
      "Negative estimates mean reduced absolute UNGA ideal-point distance to China. Donor-deletion and expanded-control rows are point-estimate diagnostics unless placebo inference is explicitly reported."
    ),
    general_title = "Note: ",
    escape = FALSE,
    threeparttable = TRUE
  )
```

Table \@ref(tab:outcome-robustness-table) reports the preferred Brazil SDiD estimate alongside three specification checks. Each column estimates the average post-2009 gap in Brazil's UNGA ideal-point distance to China and reports a placebo-based standard error. The baseline without institutional covariates, the specification without time-varying covariates, and the Latin America donor pool preserve the negative sign. Column (3) addresses post-treatment-adjustment concerns directly by holding only pre-treatment or time-invariant covariates fixed across panel years; the estimate remains nearly identical to the main specification. Column (1) remains the main specification because it keeps the theoretically relevant donor pool while adjusting for trade exposure, macroeconomic stress, geographic distance to Washington, and institutional covariates.

```{r outcome-robustness-table, message=FALSE, warning=FALSE, echo=FALSE, results='asis'}
sdid_spec_table <- tar_read(brazil_sdid_spec_table)
sdid_spec_table$table[[1]] <- ifelse(
  sdid_spec_table$table[[1]] == "Distance to US",
  "Geographic distance to US",
  sdid_spec_table$table[[1]]
)

table3_normal_p <- function(estimate_value, se_value) {
  if (is.na(estimate_value) || is.na(se_value) || se_value == 0) {
    return(NA_real_)
  }
  2 * stats::pnorm(-abs(estimate_value / se_value))
}
table3_stars <- function(p_value) {
  dplyr::case_when(
    is.na(p_value) ~ "",
    p_value < 0.01 ~ "***",
    p_value < 0.05 ~ "**",
    p_value < 0.10 ~ "*",
    TRUE ~ ""
  )
}
table3_format_estimate <- function(estimate_value, se_value) {
  paste0(
    sprintf("%.3f", estimate_value),
    table3_stars(table3_normal_p(estimate_value, se_value))
  )
}
table3_format_se <- function(se_value) {
  paste0("(", sprintf("%.3f", se_value), ")")
}
table3_estimates <- c(
  as.numeric(tar_read(synth_fit)[1]),
  as.numeric(tar_read(synth_fit_baseline)[1]),
  as.numeric(tar_read(synth_fit_no_time_varying_covariates)[1]),
  as.numeric(tar_read(synth_fit_latam)[1])
)
table3_ses <- c(
  as.numeric(tar_read(se_synth)[1]),
  as.numeric(tar_read(se_synth_baseline)[1]),
  as.numeric(tar_read(se_synth_no_time_varying_covariates)[1]),
  as.numeric(tar_read(se_synth_latam)[1])
)
table3_model_cols <- names(sdid_spec_table$table)[-1]
sdid_spec_table$table[1, table3_model_cols] <- as.list(mapply(
  table3_format_estimate,
  table3_estimates,
  table3_ses,
  USE.NAMES = FALSE
))
sdid_spec_table$table[2, table3_model_cols] <- as.list(vapply(
  table3_ses,
  table3_format_se,
  character(1)
))

knitr::kable(
  sdid_spec_table$table,
  format = "latex",
  booktabs = TRUE,
  escape = FALSE,
  caption = "Brazil SDiD estimates across main and robustness specifications.",
  align = "lcccc"
) %>%
  kableExtra::kable_styling(
    latex_options = c("HOLD_position", "scale_down"),
    font_size = 8
  ) %>%
  kableExtra::footnote(
    general = paste0(
      "The outcome is absolute UNGA ideal-point distance to China; ATT is the average post-treatment Brazil-minus-synthetic gap, so negative values mean convergence. ",
      "ATTs are reported with three decimals. Placebo-based SEs with 1,000 replications are in parentheses. ",
      "Stars use two-sided normal-approximation p-values based on placebo SEs: * p < 0.10; ** p < 0.05; *** p < 0.01. ",
      "Checkmarks show included covariate groups. ",
      "Donor counts exclude Brazil. Column (3) uses pre-treatment or time-invariant covariates held fixed across panel years."
    ),
    general_title = "Note: ",
    escape = FALSE,
    threeparttable = TRUE
  )
```

```{r china-demand-sdid-numbers, include=FALSE}
china_demand_sdid_results <- tar_read(china_demand_sdid_diagnostics_table)
china_demand_main <- china_demand_sdid_results %>%
  dplyr::filter(specification == "manuscript_main") %>%
  dplyr::slice_head(n = 1)
china_demand_commodity <- china_demand_sdid_results %>%
  dplyr::filter(specification %in% c("commodity_gfc_exposure", "plus_commodity_gfc_exposure")) %>%
  dplyr::slice_head(n = 1)
china_demand_full <- china_demand_sdid_results %>%
  dplyr::filter(specification %in% c("full_stress_test", "full_diagnostic")) %>%
  dplyr::slice_head(n = 1)
```

The rank interpretation requires more than showing that trade with China was increasing. Brazil's exports to China grew before 2009, including years in which China rose in the export hierarchy but did not become number one. The timing tests in Table \@ref(tab:table-placebo) therefore ask whether rapid trade growth and lower-rank promotions generate the same estimated UNGA convergence as the top-rank reversal. They do not. The 2003, 2004, and 2005 rows are timing and lower-threshold falsifications for the continuous-growth alternative: 2004 is especially useful because China first reaches rank #2 but is still not Brazil's largest export destination. The 2012 row probes whether the result is driven by a later break after China had already become the top destination.

The same timing logic addresses the main Brazilian political confounder. President Lula da Silva took office in 2003 and pursued a more explicit South-South foreign-policy agenda, a shift emphasized in work on Brazilian autonomy, presidential foreign policy, and international engagement [@neto2011; @mouron_urdinez2014; @neto_malamud2015; @rodrigues2019measuring; @schenoniMythsMultipolaritySources2022]. If that ideological shift were sufficient to explain Brazil-China UNGA convergence, the pseudo-treatment years in 2003 or 2005 should already show comparable movement. They do not. The cross-country panel also estimates the rank-threshold pattern across countries with different political systems and leadership profiles, which reduces the concern that the Brazil estimate is only a Lula-period artifact.

A remaining concern is that 2009 bundled the Brazilian rank reversal with other contemporaneous shocks, including the global financial crisis, the BRICS/G20 moment, and the commodity cycle. Table \@ref(tab:china-demand-sdid-diagnostics) addresses this critique inside the same Brazil SDiD framework. The main estimate is `r sprintf("%.3f", china_demand_main$estimate)`; after adding pre-2009 primary-goods export exposure and 2008-2009 exposure interactions it is `r sprintf("%.3f", china_demand_commodity$estimate)`. The point estimate also remains negative in the full stress test (`r sprintf("%.3f", china_demand_full$estimate)`), but that specification combines several highly collinear exposure diagnostics and produces unstable placebo inference. I therefore read the evidence conservatively: the point estimate for the categorical China number-one onset remains stable after adding destination concentration and predetermined commodity/GFC exposure controls, while the over-saturated stress test is useful mainly as a warning about limited support for simultaneously partialling out every continuous rank-related measure. This does not prove that the rank cue is experimentally isolated from every contemporaneous political shock. It does show that the result is not an artifact of omitting smooth China-demand exposure from the SDiD design.

The cross-country panel below is a broader scope check, not the main answer to the Brazil-specific demand-shock critique. Brazil is included in the goods-only cross-country sample, but the panel aligns countries by their own China-top goods-export entry years. A Brazil-specific 2009 confounder cannot mechanically generate an average event-time effect across countries treated in different calendar years. For such an explanation to account for both the Brazil estimate and the cross-country result, the omitted force would need to coincide with durable China-top entry across countries, affect UNGA distance to China in the same direction, and remain outside the latent factors captured by the interactive fixed-effects specification. In sensitivity-analysis terms, it would need enough residual association with both treatment timing and the outcome to overturn estimates that are stable across the Brazil timing tests, donor-pool checks, and the cross-country duration and risk-set checks [@cinelli_hazlett2020; @oster2019].

```{r table-placebo, message=FALSE, warning=FALSE, echo=FALSE, results='asis'}
rank_tests <- tar_read(goal3_brazil_placebo_rank_volume_tests) %>%
  dplyr::mutate(
    `Timing year` = nominal_treatment_year,
    `Test role` = dplyr::case_when(
      nominal_treatment_year == 2003L ~ "Growth / lower-rank promotion",
      nominal_treatment_year == 2004L ~ "China \\#2 threshold",
      nominal_treatment_year == 2005L ~ "Rapid growth without rank 1",
      nominal_treatment_year == 2009L ~ "Actual rank-1 reversal",
      nominal_treatment_year == 2012L ~ "Later-break falsification",
      TRUE ~ timing_test
    ),
    `China rank` = china_rank,
    `Rank-1 reversal` = ifelse(rank1_reversal, "Yes", "No"),
    `Export share (\\%)` = sprintf("%.1f", china_share_pct),
    `Margin (USD bn)` = sprintf("%.1f", china_margin_vs_competitor_usd_billion),
    ATT = sprintf("%.3f", estimate),
    `Placebo SE` = sprintf("%.3f", se_placebo),
    `p-value` = sprintf("%.3f", p_value)
  ) %>%
  dplyr::select(
    `Timing year`, `Test role`, `China rank`, `Rank-1 reversal`,
    `Export share (\\%)`, `Margin (USD bn)`, ATT, `Placebo SE`, `p-value`
  )
rank_tests_note <- paste0(
  sub(
    "^Note: ",
    "",
      caption_note(
      unit = "Outcome is absolute UNGA ideal-point distance to China; ATT is the estimated post-onset Brazil-minus-synthetic gap, so negative values mean convergence; export margins are current USD billions",
      se = "normal approximation using placebo-based SE with 1,000 replications",
      window = "2003-2005 timing tests truncate the sample before 2009; 2012 is a later-break falsification"
    )
  ),
  " The 2003, 2004, and 2005 rows are timing and lower-threshold falsifications for rapid export growth without rank-1 status."
)

knitr::kable(
  rank_tests,
  format = "latex",
  booktabs = TRUE,
  escape = FALSE,
  caption = "Brazil SDiD estimate and rank-versus-volume timing tests.",
  align = "clccccccc"
) %>%
  kableExtra::kable_styling(
    latex_options = c("HOLD_position", "scale_down"),
    font_size = 8
  ) %>%
  kableExtra::footnote(
    general = rank_tests_note,
    general_title = "Note: ",
    escape = FALSE,
    threeparttable = TRUE
  )
```

```{r china-demand-sdid-diagnostics, message=FALSE, warning=FALSE, echo=FALSE, results='asis'}
format_diag_p <- function(p_value, se_value) {
  dplyr::case_when(
    is.na(p_value) ~ "NA",
    is.na(se_value) ~ "NA",
    se_value > 10 ~ "unstable",
    p_value < 0.001 ~ "<0.001",
    TRUE ~ sprintf("%.3f", p_value)
  )
}

format_diag_se <- function(se_value) {
  dplyr::case_when(
    is.na(se_value) ~ "NA",
    se_value > 10 ~ "unstable",
    TRUE ~ sprintf("%.3f", se_value)
  )
}

format_diag_ci <- function(ci_low, ci_high, se_value) {
  dplyr::case_when(
    is.na(ci_low) | is.na(ci_high) ~ "NA",
    is.na(se_value) ~ "NA",
    se_value > 10 ~ "not informative",
    TRUE ~ paste0("[", sprintf("%.3f", ci_low), ", ", sprintf("%.3f", ci_high), "]")
  )
}

china_demand_table <- china_demand_sdid_results %>%
  dplyr::mutate(
    Specification = label,
    ATT = sprintf("%.3f", estimate),
    `Placebo SE` = format_diag_se(se_placebo),
    `p-value` = format_diag_p(p_value, se_placebo),
    `95\\% CI` = format_diag_ci(ci_95_low, ci_95_high, se_placebo),
    `Diagnostic added` = control_family
  ) %>%
  dplyr::select(
    Specification,
    `Diagnostic added`,
    ATT,
    `Placebo SE`,
    `95\\% CI`,
    `p-value`
  )

china_demand_note <- paste0(
  sub(
    "^Note: ",
    "",
      caption_note(
      unit = "Outcome is absolute UNGA ideal-point distance to China; ATT is the average post-treatment Brazil-minus-synthetic gap, so negative values mean convergence",
      se = "synthdid placebo SE with 1,000 replications",
      window = "1997-2015 annual SDiD panel"
    )
  ),
  " All rows keep the binary 2009 China number-one treatment onset. ",
  "Commodity/GFC exposure uses pre-2009 primary-goods export composition from ITPD-E Agriculture plus Mining and Energy over goods exports, interacted with the 2008-2009 shock window. ",
  "The full stress test includes the rank margin, which is mechanically related to the threshold, and several continuous exposure transformations; its placebo SE is numerically unstable, so that row is a conservative support diagnostic rather than a preferred inferential specification."
)

knitr::kable(
  china_demand_table,
  format = "latex",
  booktabs = TRUE,
  escape = FALSE,
  caption = "Brazil SDiD diagnostics for the China-demand-shock critique.",
  align = "llcccc"
) %>%
  kableExtra::kable_styling(
    latex_options = c("HOLD_position", "scale_down"),
    font_size = 8
  ) %>%
  kableExtra::footnote(
    general = china_demand_note,
    general_title = "Note: ",
    escape = FALSE,
    threeparttable = TRUE
  )
```

## Issue-Area Evidence: Where Did Convergence Occur?

```{r selective-unga-diagnostic-numbers, include=FALSE}
selective_vote_models <- tar_read(selective_china_alignment_vote_level_models)
selective_ddd_models <- tar_read(selective_china_alignment_ddd_hr_nonhr_models)
selective_country_placebo <- tar_read(selective_china_alignment_country_placebo_summary)

hr_china_us_diff <- selective_vote_models %>%
  dplyr::filter(
    sample == "Human-rights China-US divergent votes",
    outcome == "distance_china_minus_usa"
  ) %>%
  dplyr::slice_head(n = 1)

hr_china_us_strong_diff <- selective_vote_models %>%
  dplyr::filter(
    sample == "Human-rights strong yes/no China-US divergent votes",
    outcome == "distance_china_minus_usa"
  ) %>%
  dplyr::slice_head(n = 1)

nonhr_china_us_diff <- selective_vote_models %>%
  dplyr::filter(
    sample == "Non-human-rights China-US divergent votes",
    outcome == "distance_china_minus_usa"
  ) %>%
  dplyr::slice_head(n = 1)

ddd_hr_increment <- selective_ddd_models %>%
  dplyr::filter(
    outcome == "distance_china_minus_usa",
    term == "brazil_post_hr",
    stringr::str_detect(model, "country-clustered")
  ) %>%
  dplyr::slice_head(n = 1)

distance_placebo <- selective_country_placebo %>%
  dplyr::filter(outcome == "distance_china_minus_usa") %>%
  dplyr::slice_head(n = 1)
```

The aggregate ideal-point estimate raises a substantive question: in what kinds of votes did Brazil move closer to China? Resolution-level diagnostics suggest that the post-2009 convergence was selective rather than uniform. In several issue areas Brazil and China already voted together at high rates before treatment, leaving little room for additional convergence. The more informative domain is human rights, where pre-treatment agreement was lower and where movement toward China carried greater reputational costs for Brazil, given the domestic and international implications of Brazil's human-rights foreign policy [@milani2015humanrights]. In that domain, Brazil-China identical voting increased by `r sprintf("%.1f", human_rights_vote_change)` percentage points after 2009. This pattern is consistent with the proposed mechanism: the trade-rank cue made China's relationship politically usable, allowing policymakers to justify selective convergence in a visible multilateral arena where Brazil had previously maintained more distance.

This descriptive increase does not by itself establish China-specific alignment, because annual issue-area shares can reflect agenda composition and do not compare Brazil with donor countries. I therefore add a more restrictive vote-level diagnostic. In country-resolution regressions with country and resolution fixed effects, restricted to human-rights votes where China and the United States cast different votes, Brazil moves closer to China after 2009 relative to the donor pool. The Brazil-by-post-2009 coefficient is `r sprintf("%.3f", hr_china_us_diff$estimate)` for distance to China minus distance to the United States (model-based SE = `r sprintf("%.3f", hr_china_us_diff$se)`), and `r sprintf("%.3f", hr_china_us_strong_diff$estimate)` when the sample is restricted to strong yes/no China-US disagreements (model-based SE = `r sprintf("%.3f", hr_china_us_strong_diff$se)`). The corresponding non-human-rights estimate is smaller (`r sprintf("%.3f", nonhr_china_us_diff$estimate)`, model-based SE = `r sprintf("%.3f", nonhr_china_us_diff$se)`), and a DDD specification estimates an additional human-rights increment of `r sprintf("%.3f", ddd_hr_increment$estimate)` (model-based SE = `r sprintf("%.3f", ddd_hr_increment$se)`). Because Brazil is the single treated country in this diagnostic, I use country placebos as the main inferential benchmark: Brazil ranks `r sprintf("%.0f", distance_placebo$brazil_rank_expected_direction)` out of `r sprintf("%.0f", distance_placebo$n_placebo_units)` units in the China-specific direction (directional p = `r sprintf("%.3f", distance_placebo$randomization_p_directional)`; strict donor-only p = `r sprintf("%.3f", distance_placebo$randomization_p_directional_strict_donor)`). These diagnostics support a narrower interpretation: the post-2009 movement is strongest in human-rights votes where China and the United States diverged, rather than uniformly across the UNGA agenda.

```{r issue-area-prepost-table, message=FALSE, warning=FALSE, echo=FALSE, results='asis'}
issue_order <- c(
  "Human rights",
  "Arms/disarmament/nuclear",
  "Economic development",
  "Decolonization",
  "Palestine/Middle East",
  "Other / uncoded"
)

issue_prepost_tbl <- tar_read(brazil_china_unvotes_resolution_2005_2012) %>%
  dplyr::mutate(period = dplyr::if_else(year < 2009, "pre", "post")) %>%
  dplyr::summarise(
    n_resolutions = dplyr::n_distinct(rcid),
    identical_vote_share = 100 * mean(identical_vote, na.rm = TRUE),
    .by = c(issue_family, period)
  ) %>%
  tidyr::pivot_wider(
    names_from = period,
    values_from = c(n_resolutions, identical_vote_share)
  ) %>%
  dplyr::mutate(
    change = identical_vote_share_post - identical_vote_share_pre,
    interpretation = dplyr::case_when(
      issue_family == "Human rights" ~ "Politically costly convergence; mechanism-informative",
      issue_family == "Palestine/Middle East" ~ "High baseline agreement; limited room to move",
      issue_family == "Economic development" ~ "Diagnostic evidence",
      issue_family == "Arms/disarmament/nuclear" ~ "Diagnostic evidence",
      issue_family == "Decolonization" ~ "Diagnostic evidence",
      TRUE ~ "Heterogeneous residual category"
    ),
    issue_order_id = match(issue_family, issue_order)
  ) %>%
  dplyr::arrange(is.na(issue_order_id), issue_order_id, issue_family) %>%
  dplyr::transmute(
    `Issue area` = issue_family,
    `Pre-2009 identical vote share` = sprintf("%.1f", identical_vote_share_pre),
    `Post-2009 identical vote share` = sprintf("%.1f", identical_vote_share_post),
    `Change` = sprintf("%.1f", change),
    `Interpretation` = interpretation
  )

knitr::kable(
  issue_prepost_tbl,
  format = "latex",
  booktabs = TRUE,
  escape = FALSE,
  caption = "Brazil-China identical UNGA votes by issue area, before and after 2009.",
  align = "lrrrl"
) %>%
  kableExtra::kable_styling(
    latex_options = c("HOLD_position", "scale_down"),
    font_size = 8
  ) %>%
  kableExtra::footnote(
    general = paste0(
      sub(
        "^Note: ",
        "",
        caption_note(
          unit = "Percentage of roll-call resolutions with identical Brazil-China votes",
          window = "2005-2008 versus 2009-2012; multi-coded resolutions enter each issue family"
        )
      ),
      " Rows are label-specific diagnostics, not a mutually exclusive decomposition of the aggregate convergence pattern."
    ),
    general_title = "Note: ",
    escape = FALSE,
    threeparttable = TRUE
  )
```

Figure \@ref(fig:plot-brazil-china-unvotes-issue-year) disaggregates the same diagnostic by year and substantive issue family. Because several areas already had high Brazil-China vote similarity before 2009, the figure describes where additional convergence was possible.

```{r plot-brazil-china-unvotes-issue-year, message=FALSE, warning=FALSE, echo=FALSE, fig.cap=paste0("Brazil-China identical UNGA votes by issue area, 2005-2012. Points report annual shares; the vertical line marks 2009. The clearest pattern expected by the mechanism is selective convergence in human rights. ", caption_note(unit = "Percentage of roll-call resolutions with identical Brazil-China votes", window = "2005-2012 annual series; multi-coded resolutions enter each issue family")), fig.pos="H", fig.width=11, fig.height=8}
tar_read(plot_brazil_china_unvotes_similarity_by_issue_year_2005_2012)
```

This issue-area evidence is a diagnostic of substantive content, not a standalone causal test. The SDiD estimate identifies the average post-2009 gap in Brazil's ideal-point distance to China; the vote-level evidence clarifies where that movement is most visible. The strongest support comes from human-rights resolutions in which China and the United States diverged, where Brazil moves closer to China relative to comparable donor countries after 2009. A DDD comparison shows that the human-rights shift is larger than the analogous non-human-rights shift. Country placebos counsel against language implying that Brazil is unique. The Brazil evidence therefore supports a cautious claim of selective UNGA adjustment toward China in politically visible domains, rather than a claim that the issue-area table alone proves selective convergence.

# Brazilian media salience

To assess whether the rank reversal became publicly salient, I examine coverage of China in Folha de S.Paulo, Brazil's widest-circulation newspaper [@folha2016], from 2000 to 2014. Salience matters because the rank-threshold argument requires the top-rank cue to become politically usable. A change in export hierarchy can shape elite and public attention when it appears in a prominent national newspaper as a simple signal of China's commercial importance, consistent with work on media, agenda setting, and foreign policy [@soroka2003; @baum_potter2008; @walgraveContingencyMassMedias2006].

Headlines are the appropriate unit for this diagnostic because the mechanism is public attention and categorical labeling. This follows corpus-assisted media studies that use headlines because they are tractable in large archives and operate as attention-directing cues [@zhang_yu2024_chinese_dream]. The relevant question is whether Folha increasingly foregrounded China as a trade relationship after the rank reversal. Headlines are the most visible textual cue in the archive, they are the part of coverage most likely to be scanned by broad audiences, and they are where newspapers condense complex developments into short public labels such as "largest trade partner" or "record exports." Full article texts would be useful for a different exercise -- for example, measuring tone or detailed argumentation -- but the salience claim here is about which China-related themes were made visible at the headline level.

I therefore classify China-related headlines into substantive categories and then collapse them into four analytically relevant groups: China-Brazil trade, China's domestic economy, diplomacy and bilateral relations, and non-economic coverage. The classification uses a fixed label set with examples, and the appendix reports the exact prompt and validation procedure. Figure \@ref(fig:plot-media-counterfactual) shows that the post-2009 increase is concentrated in China-Brazil trade coverage rather than in all China-related topics equally. This pattern is consistent with a rank-salience interpretation: coverage concentrates on China's new commercial status in Brazil.

```{r plot-media-counterfactual, message=FALSE, warning=FALSE, echo=FALSE, fig.cap=paste0("Brazilian media salience: disaggregated China headlines in Folha de S.Paulo by category. The sustained post-2009 increase is concentrated in trade-related coverage. ", caption_note(unit = "Headlines per category per year", window = "2000-2014 annual series")), fig.pos="H", fig.asp = 0.7, fig.width = 8}
mc <- tar_read(media_counterfactual)
ggplot(mc, aes(x = year, y = n, fill = broad_category)) +
  geom_area(alpha = 0.7) +
  geom_vline(xintercept = 2009, linetype = "dashed", colour = "grey30") +
  annotate("text", x = 2009.3, y = max(mc$total) * 0.95, label = "Rank reversal",
           hjust = 0, size = 3, colour = "grey30") +
  scale_x_continuous(breaks = seq(2000, 2014, by = 2)) +
  scale_fill_manual(values = c(
    "China-Brazil Trade" = "#d95f02",
    "Chinese Economy" = "#7570b3",
    "Diplomacy & Relations" = "#1b9e77",
    "Non-economic (sports, health, disasters, etc.)" = "#999999"
  ),
  labels = c(
    "China-Brazil Trade" = "China-Brazil trade",
    "Chinese Economy" = "Chinese economy",
    "Diplomacy & Relations" = "Diplomacy/relations",
    "Non-economic (sports, health, disasters, etc.)" = "Other non-economic"
  )) +
  labs(x = "Year", y = "Number of headlines", fill = "Category") +
  guides(fill = guide_legend(nrow = 2, byrow = TRUE)) +
  theme_minimal() +
  theme(
    legend.position = "bottom",
    legend.text = element_text(size = 8),
    legend.title = element_text(size = 8)
  )
```

\clearpage

The media evidence documents public availability and official uptake of the attention cue. The status cue became publicly available during the 2009 treatment year, before the full-year trade table was finalized. In early May, contemporary coverage of January--April trade statistics reported a historic shift in Brazil's trade hierarchy: China had become Brazil's largest trading partner and had purchased US$5.627 billion in Brazilian exports, compared with US$4.925 billion purchased by the United States [@agencia_brasil2009_china_supera]. This real-time visibility matters for the mechanism. The empirical design codes treatment annually because the outcome and comparative trade data are annual, but officials, journalists, and firms encountered the rank cue through contemporaneous monthly and year-to-date statistics. Lula's May 2009 Beijing visit therefore provides evidence of official uptake during the treatment year: the new rank was already available as a public category around which diplomatic priorities could be narrated. Brazilian officials then repeatedly invoked China's new position in Brazil's trade hierarchy when explaining diplomatic prioritization, bilateral upgrading, and institutional follow-up.

In a written interview on the eve of his May 2009 visit to Beijing, President Lula described China as Brazil’s leading trading partner and framed the trip as an effort to consolidate cooperation across commerce, technology, and global economic governance [@mre_funag_resenha104_2009, 323]. In Beijing three days later, Lula repeated the rank claim in speeches organized around the Brazil-China Strategic Partnership, trade diversification, investment, presidential diplomacy, and the preparation of the 2010–2014 Joint Action Plan [@mre_funag_resenha104_2009, 115, 121–122]. The cue then persisted in Itamaraty language. When announcing Hu Jintao’s 2010 state visit and the signature of the Joint Action Plan, the Ministry of Foreign Affairs again invoked China’s 2009 rise to the top of Brazil’s trade hierarchy [@mre_funag_resenha106_2010, 356]. Dilma Rousseff and Hu Jintao’s 2011 joint communiqué likewise noted that China had become Brazil’s principal trading partner in 2009 and embedded that status in an agenda of intensified dialogue over trade, investment, diversification, and implementation of the Joint Action Plan [@mre_funag_resenha108_2011, 160–161]. Later that year, Foreign Minister Antonio Patriota made the institutional link explicit: in discussing China, he announced a dedicated Itamaraty task force to monitor economic-commercial relations with Brazil’s principal trading partner and to propose ways to move the relationship beyond complementarity, toward diversification and higher value-added trade [@mre_funag_resenha109_2011, 47].

These statements show that the 2009 status cue was absorbed into official language in ways consistent with justificatory use for diplomatic prioritization and institutional attention. This matches the public mechanism implied by the theory: once China was officially recognized as central in Brazil’s economic hierarchy, selective diplomatic accommodation could be presented as pragmatic recognition of national economic interests rather than as ideological concession or opportunistic bandwagoning.

# Cross-Country Evidence Beyond Brazil

The cross-country analysis serves two purposes. First, it tests whether durable top-rank goods-export status is followed by movement toward China beyond the Brazilian case. Second, it is an additional robustness check against explanations tied to Brazil's 2009 calendar year, such as the global financial crisis, the BRICS/G20 moment, or the commodity cycle. If durable top-rank trade status matters, countries where China enters and holds the largest goods export-destination position should move closer to China in UNGA ideal-point space when aligned by their own treatment entry. I therefore estimate the same treatment logic in a goods-only status-current panel: treated country-years are those where China is rank 1 among goods export destinations and the China-top period lasts at least five observed years. Never-China-top countries remain controls. Short China-top episodes and post-exit off-status years for treated countries are removed from the main risk set. This analysis evaluates whether the rank-threshold pattern is compatible with evidence beyond the theory-generating case while matching the durable treatment observed in Brazil.

```{r fect-results, include=FALSE}
library(dplyr)
m2_model_results <- tar_read(china_top_m2_goods_status_current_model_results)
m2_sector_audit <- tar_read(china_top_m2_goods_sector_audit)
m2_unit_summary <- tar_read(china_top_m2_goods_status_current_unit_summary)
m2_country_audit <- tar_read(china_top_m2_goods_status_current_country_audit)

main_cross_s <- m2_model_results %>%
  dplyr::filter(min_duration_years == 5L, specification == "risk_set_restricted") %>%
  dplyr::slice(1)
clean_cross_s <- m2_model_results %>%
  dplyr::filter(min_duration_years == 5L, specification == "clean_single_spell") %>%
  dplyr::slice(1)
switching_cross_s <- m2_model_results %>%
  dplyr::filter(min_duration_years == 5L, specification == "switching_allowed") %>%
  dplyr::slice(1)

n_treated_abs <- main_cross_s$n_treated
n_control_abs <- main_cross_s$n_control
panel_min <- main_cross_s$panel_min
panel_max <- main_cross_s$panel_max
goods_sector_values <- c("Agriculture", "Mining and Energy", "Manufacturing")
m2_bootstrap_label <- paste0(
  format(
    max(m2_model_results$nboots, na.rm = TRUE),
    big.mark = ",",
    scientific = FALSE,
    trim = TRUE
  ),
  " replications"
)

format_p <- function(x) {
  if (length(x) == 0 || is.na(x)) {
    return("not available")
  }
  if (x < 0.001) {
    return("<0.001")
  }
  formatC(x, format = "f", digits = 3)
}
fmt3 <- function(x) sprintf("%.3f", unname(x))
fmt1 <- function(x) sprintf("%.1f", unname(x))
fmt_num <- function(x, digits = 3L) {
  ifelse(is.na(x), "NA", formatC(x, format = "f", digits = digits))
}
sig_sentence <- function(att, p) {
  direction <- ifelse(att < 0, "in the theoretically expected negative direction", "positive")
  if (is.na(p)) {
    "The statistical precision of this estimate is not available in the current diagnostic output."
  } else if (p < 0.05) {
    paste0("The estimate is ", direction, " and statistically distinguishable from zero at the 5 percent level.")
  } else {
    paste0("The estimate is ", direction, ", but the confidence interval includes zero.")
  }
}
```

The preferred cross-country specification is the no-covariate IFE model on the goods-only restricted risk set. The estimate is `r fmt3(main_cross_s$att)` ideal-point units (bootstrap SE = `r fmt3(main_cross_s$se)`, 95 percent CI [`r fmt3(main_cross_s$ci_lo)`, `r fmt3(main_cross_s$ci_hi)`], p = `r format_p(main_cross_s$p)`). It corresponds to about `r fmt1(main_cross_s$att_rel_pct)` percent of the pre-treatment mean distance to China, or `r fmt3(main_cross_s$att_sd_units)` standard deviations of the outcome. `r sig_sentence(main_cross_s$att, main_cross_s$p)` The clean single-entry robustness estimate is `r fmt3(clean_cross_s$att)` (p = `r format_p(clean_cross_s$p)`). The switching-allowed estimate is smaller, `r fmt3(switching_cross_s$att)` (p = `r format_p(switching_cross_s$p)`), which is consistent with attenuation when short China-top periods and post-exit off-status years are allowed into the comparison structure. The cross-country evidence is smaller than the Brazil SDiD effect, consistent with Brazil being a most-likely case in which China displaced the United States and the rank reversal was publicly salient. The direction is the same: durable China-top goods-export status is associated with reduced UNGA distance to China beyond the Brazilian case. Appendix Tables \@ref(tab:cross-country-duration-table) and \@ref(tab:cross-country-audit-table) report duration-threshold robustness and sample-coding audits.

```{r cross-country-table, message=FALSE, warning=FALSE, echo=FALSE, results='asis'}
cross_tbl <- m2_model_results %>%
  dplyr::filter(min_duration_years == 5L) %>%
  dplyr::mutate(
    Model = specification_label,
    Estimator = "IFE",
    `ATT` = fmt3(att),
    `SE` = fmt3(se),
    `95\\% CI` = paste0("[", fmt3(ci_lo), ", ", fmt3(ci_hi), "]"),
    `p-value` = vapply(p, format_p, character(1)),
    `Units (T/C)` = paste0(n_treated, "/", n_control),
    `Treated periods` = as.character(n_treated_country_years),
    `Panel window` = paste0(panel_min, "--", panel_max)
  ) %>%
  dplyr::transmute(
    Model,
    Estimator,
    ATT,
    SE,
    `95\\% CI`,
    `p-value`,
    `Units (T/C)`,
    `Treated periods`,
    `Panel window`,
    `r*` = r_cv
  )
cross_note <- paste0(
  sub(
    "^Note: ",
    "",
    caption_note(
      unit = "ATT in absolute UNGA ideal-point distance to China; lower values indicate convergence toward China",
      se = paste("bootstrap with", m2_bootstrap_label),
      window = sprintf("%d--%d goods-only cross-country panel", main_cross_s$panel_min, main_cross_s$panel_max)
    )
  ),
  " Goods-only ranks are computed from ITPD-E Agriculture, Mining and Energy, and Manufacturing sectors; Services are excluded before ranks are constructed. ",
  "Units (T/C) = treated/control units. ",
  "For IFE rows, r* is the number of latent factors selected by cross-validation over r = 0:3. ",
  "The restricted risk-set row is the main cross-country estimate. The clean single-entry row excludes treated countries with multiple China-top periods. The switching-allowed row is a status-current robustness check that keeps a balanced panel and changes the comparison structure."
)

knitr::kable(
  cross_tbl,
  format = "latex",
  booktabs = TRUE,
  escape = FALSE,
  caption = "Cross-country IFE estimates for durable China top goods export-destination status.",
  align = "llcccccccc"
) %>%
  kableExtra::kable_styling(
    latex_options = c("HOLD_position", "scale_down"),
    font_size = 8
  ) %>%
  kableExtra::footnote(
    general = cross_note,
    general_title = "Note: ",
    escape = FALSE,
    threeparttable = TRUE
)
```

Figure \@ref(fig:cross-country-dynamic) shows the entry-aligned dynamic pattern for the main goods-only restricted-risk-set IFE specification. The pre-entry estimates are diagnostics for model fit, and the post-entry path indicates whether the panel pattern is consistent with the Brazil result. Event-time tails with few treated units are descriptive; the average ATT and the duration-threshold table are the better-supported cross-country summaries.

```{r cross-country-dynamic, message=FALSE, warning=FALSE, echo=FALSE, fig.cap=paste0("Entry-aligned dynamic effects from the main goods-only cross-country IFE model. Negative estimates indicate reduced UNGA distance to China after entry into durable status as the top goods-export destination. Event-time tails with few treated units should be read descriptively. ", caption_note(unit = "Effect on absolute UNGA ideal-point distance to China", se = "bootstrap confidence intervals", window = sprintf("%d--%d goods-only panel", main_cross_s$panel_min, main_cross_s$panel_max))), fig.pos="H", fig.width=7, fig.asp=0.65}
tar_read(plot_china_top_m2_goods_status_current_dynamic)
```

The two designs test nested implications of the argument. Brazil is the high-salience case: China displaced the United States, the hegemonic benchmark, and the attention mechanism can be observed directly in media and official discourse. The cross-country panel tests a broader scope implication: whether durable status as the top goods-export destination is followed by movement toward China when local salience is less directly observable. The panel estimate should therefore be interpreted as scope evidence rather than as a substitute for the Brazilian mechanism test.

The cross-country treatment is durable status-current China-top goods-export-destination status, the exposure rule implied by the theory's durability condition. The corresponding estimand is the ATT on absolute UNGA ideal-point distance to China for country-years in qualifying China-top periods.

# Conclusion

The evidence supports a rank-threshold account of UNGA voting convergence toward China. In Brazil, China's 2009 move into the top export-destination position coincided with a publicly visible status cue, and SDiD estimates show that Brazil moved closer to China relative to a synthetic counterfactual. The issue-area and vote-level evidence suggest that this convergence was selective rather than mechanical: the strongest additional support comes from human-rights resolutions where China and the United States diverged, and the human-rights shift is larger than the analogous non-human-rights shift. The media evidence shows that the rank cue was publicly available and politically usable: China's new commercial status became visible in Folha headlines and official discourse. Cross-country estimates from the goods-only IFE model then show a compatible directional pattern beyond Brazil. The mechanism is measured most directly in the Brazilian case, and the panel evidence extends the argument beyond the theory-generating case: durable top-rank goods-export milestones can matter for multilateral alignment.

The paper therefore shifts attention from trade volume alone to the political meaning of trade hierarchies. Continuous exposure matters, but actors also respond to categorical milestones: becoming first, displacing a hegemonic benchmark, and making an economic relationship publicly legible. These thresholds can create opportunities for selective diplomatic adjustment even when the underlying commercial relationship has been growing for years.

Future research can extend the design in three directions. First, cross-country media data would allow the salience step to be tested beyond Brazil. Second, qualitative work on policymakers and business actors could identify how the rank cue entered diplomatic deliberation. Third, similar rank-threshold designs could be applied to investment, aid, lending, or other economic ties where categorical milestones may become politically usable.

\clearpage

# References {-}

::: {#refs}
:::

\clearpage

# Appendix

We present more details about the performance and checks on the fitted model and how we used ChatGPT to perform the NLP analysis.

## Brazil and China Ideal-Point Series

Figure \@ref(fig:appendix-brazil-china-ideal-points) plots the separate UNGA ideal-point series for Brazil and China. This diagnostic clarifies the interpretation of the absolute-distance outcome used in the main SDiD design: China remains comparatively stable, while Brazil accounts for most of the movement in the Brazil-China distance.

```{r appendix-brazil-china-ideal-points, message=FALSE, warning=FALSE, echo=FALSE, fig.cap=paste0("Brazil and China UNGA ideal-point series. The dashed line marks 2009. The figure separates the two series behind the absolute-distance outcome used in the main SDiD analysis. ", caption_note(unit = "UNGA ideal-point score (Bailey et al. scale)", window = "1990-2022 annual series")), fig.pos="H", fig.width=7, fig.asp=0.55}
ideal_series_brazil_china <- tar_read(unga_data) %>%
  dplyr::filter(iso3c %in% c("BRA", "CHN")) %>%
  dplyr::mutate(
    country = dplyr::case_when(
      iso3c == "BRA" ~ "Brazil",
      iso3c == "CHN" ~ "China",
      TRUE ~ iso3c
    )
  ) %>%
  dplyr::select(year, country, ideal_point_all)

ggplot(ideal_series_brazil_china, aes(x = year, y = ideal_point_all, colour = country)) +
  geom_line(linewidth = 0.6) +
  geom_point(size = 1.6) +
  geom_vline(xintercept = 2009, linetype = "dashed", colour = "grey35") +
  scale_colour_manual(values = c("Brazil" = "darkgreen", "China" = "firebrick")) +
  labs(x = "Year", y = "UNGA ideal point", colour = NULL) +
  theme_minimal() +
  theme(legend.position = "bottom")
```

## Resolution-Level UNGA Vote Diagnostics

To make the Brazil-China movement in UNGA ideal points more concrete, Table \@ref(tab:table-brazil-china-unvotes-year) reports the annual share of resolutions in which Brazil and China cast identical votes from 2005 to 2012. The diagnostic uses roll-call votes from `unvotes` 0.3.0, based on Erik Voeten's UNGA voting data, and excludes missing or absent votes by either Brazil or China.

```{r table-brazil-china-unvotes-year, message=FALSE, warning=FALSE, echo=FALSE, results='asis'}
bc_year_table <- tar_read(brazil_china_unvotes_similarity_by_year_2005_2012) %>%
  dplyr::mutate(
    identical_votes_percent = sprintf("%.1f", identical_votes_percent),
    mean_similarity_score = sprintf("%.3f", mean_similarity_score)
  ) %>%
  dplyr::select(
    year,
    n_resolutions,
    identical_votes_n,
    identical_votes_percent,
    mean_similarity_score
  )

knitr::kable(
  bc_year_table,
  format = "latex",
  booktabs = TRUE,
  caption = "Brazil-China identical UNGA votes by year, 2005-2012.",
  col.names = c(
    "Year",
    "Resolutions",
    "Identical votes",
    "Identical votes (\\%)",
    "Mean similarity score"
  ),
  align = "ccccc"
) %>%
  kableExtra::kable_styling(
    latex_options = c("HOLD_position", "scale_down"),
    font_size = 9
  ) %>%
  kableExtra::footnote(
    general = sub(
      "^Note: ",
      "",
      caption_note(
        unit = "Roll-call resolutions with valid Brazil and China votes; similarity score ranges from 0 to 1",
        window = "2005-2012 annual series"
      )
    ),
    general_title = "Note: ",
    escape = FALSE,
    threeparttable = TRUE
  )
```

## SDiD

### Comparison of DiD, SCM, and SDiD estimating equations

To gain intuition for the SDiD estimator, it is useful to compare it with standard DiD and SCM. A DiD model estimates:

$$
\bigl(\hat{\tau}_{ATT}^{\text{DiD}}, \hat{\mu}, \hat{\alpha}, \hat{\beta}\bigr)
=
\operatorname*{arg\,min}_{\tau_{ATT},\mu, \alpha, \beta}
\sum_{i=1}^{N}\sum_{t=1}^{T}
\bigl(Y_{it}-\mu-\alpha_i -\beta_t-D_{it}\tau_{ATT}\bigr)^2
$$

While SDiD assigns weights to the control group that are more similar to those of the treated unit and in time periods comparable to the pre-treatment period, classic DiD does not use any weights at all.

The SCM model can be written as follows:

$$
\bigl(\hat{\tau}_{ATT}^{\text{SCM}}, \hat{\mu}, \hat{\beta}\bigr)
=
\operatorname*{arg\,min}_{\tau_{ATT}, \mu, \beta}
\sum_{i=1}^{N}\sum_{t=1}^{T}
\hat{w}^{\text{SCM}}_i
\bigl(Y_{it}-\mu-\beta_t-D_{it}\tau_{ATT}\bigr)^2
$$

Here $\hat{w}^{\text{SCM}}_i$ is the donor-unit weight assigned to unit $i$. SCM uses these donor weights to match the treated unit's pre-treatment outcome path, and effects are then read from post-treatment treated-minus-synthetic gaps. SCM can approximate stable unit differences through donor weights when the treated unit lies in the donor convex hull, but it does not add unit fixed effects or time weights. SDiD combines unit weights (like SCM) with unit fixed effects and time weights (like DiD), addressing the limitations of both. The basic SDiD estimating equation is:

$$
\bigl(\hat{\tau}_{ATT}^{\text{SDiD}}, \hat{\mu}, \hat{\alpha}, \hat{\beta} \bigr)
=
\operatorname*{arg\,min}_{\tau_{ATT},\mu,\beta, \alpha}
\sum_{i=1}^{N}\sum_{t=1}^{T}
\hat{w}^{\text{SDiD}}_i \hat{\lambda}^{\text{SDiD}}_t
\bigl(Y_{it}- \mu- \alpha_i - \beta_t-D_{it}\tau_{ATT}\bigr)^2
$$

Note that SDiD retains unit fixed effects $\alpha_i$ (like DiD) while also assigning unit weights $\hat{w}^{\text{SDiD}}_i$ (like SCM) and time weights $\hat{\lambda}^{\text{SDiD}}_t$ (unique to SDiD).

### SDiD weights and donor pool

Tables \@ref(tab:sdid-unit-weights-complete)-\@ref(tab:sdid-high-weight-donor-exposure) report the Brazil SDiD diagnostic package behind the compact summary in Section 3.1. I report the complete donor-weight vector rather than only the largest donors, all pre-treatment time weights, fit and balance diagnostics, rank-based placebo inference, donor-influence checks, window sensitivity, and donor exposure to China's export expansion.

```{r sdid-unit-weights-complete, message=FALSE, warning=FALSE, echo=FALSE, results='asis'}
sdid_unit_weights_table <- sdid_unit_weights %>%
  dplyr::transmute(
    Rank = as.integer(weight_rank),
    ISO3 = iso3c,
    Donor = country_name,
    Weight = sprintf("%.4f", unit_weight),
    `Cumulative weight` = sprintf("%.4f", cumulative_weight)
  )

knitr::kable(
  sdid_unit_weights_table,
  format = "latex",
  booktabs = TRUE,
  longtable = TRUE,
  caption = "Complete SDiD donor-unit weights for the Brazil synthetic counterfactual.",
  align = "rllrr"
) %>%
  kableExtra::kable_styling(
    font_size = 7
  ) %>%
  kableExtra::footnote(
    general = "Weights sum to one over the donor pool; Brazil is excluded from the donor pool.",
    general_title = "Note: ",
    escape = FALSE,
    threeparttable = TRUE
  )
```

```{r sdid-time-weights, message=FALSE, warning=FALSE, echo=FALSE, results='asis'}
sdid_time_weights_table <- sdid_time_weights %>%
  dplyr::transmute(
    Year = as.integer(year),
    `Time weight` = sprintf("%.4f", time_weight),
    `Weight rank` = as.integer(time_weight_rank),
    `High weight` = dplyr::if_else(high_time_weight, "Yes", "No")
  )

knitr::kable(
  sdid_time_weights_table,
  format = "latex",
  booktabs = TRUE,
  caption = "SDiD pre-treatment time weights for the Brazil baseline.",
  align = "rrrr"
) %>%
  kableExtra::kable_styling(
    latex_options = c("HOLD_position"),
    font_size = 8
  ) %>%
  kableExtra::footnote(
    general = "The weights are assigned over the 1997-2008 pre-treatment period.",
    general_title = "Note: ",
    escape = FALSE,
    threeparttable = TRUE
  )
```

```{r sdid-prefit-summary, message=FALSE, warning=FALSE, echo=FALSE, results='asis'}
sdid_prefit_table <- sdid_fit_summary %>%
  dplyr::transmute(
    Specification = specification,
    ATT = sprintf("%.3f", estimate),
    `Placebo SE` = sprintf("%.3f", placebo_se),
    `Adj. pre RMSPE` = sprintf("%.3f", rmspe_pre),
    `Adj. post RMSPE` = sprintf("%.3f", rmspe_post),
    `Post/pre adj. RMSPE` = sprintf("%.3f", rmspe_ratio),
    `Pre-treatment mean` = sprintf("%.3f", brazil_pre_treatment_mean),
    `% of pre-mean` = sprintf("%.1f", estimate_as_percent_of_pre_mean)
  )

knitr::kable(
  sdid_prefit_table,
  format = "latex",
  booktabs = TRUE,
  caption = "Intercept-adjusted pre-treatment fit summary for the Brazil SDiD baseline.",
  align = "lrrrrrrr"
) %>%
  kableExtra::kable_styling(
    latex_options = c("HOLD_position", "scale_down"),
    font_size = 8
  ) %>%
  kableExtra::footnote(
    general = "ATT is the average post-2009 Brazil-minus-synthetic gap in absolute UNGA ideal-point distance to China. RMSPE diagnostics center the Brazil-minus-synthetic gap by its pre-treatment mean before computing RMSPE, reflecting SDiD's allowance for stable level differences.",
    general_title = "Note: ",
    escape = FALSE,
    threeparttable = TRUE
  )
```

```{r sdid-balance, message=FALSE, warning=FALSE, echo=FALSE, results='asis'}
sdid_balance_table <- sdid_balance %>%
  dplyr::transmute(
    Variable = variable,
    Role = role,
    `Brazil pre-mean` = sprintf("%.3f", brazil_pre_mean),
    `Synthetic pre-mean` = sprintf("%.3f", synthetic_pre_mean),
    Difference = sprintf("%.3f", brazil_minus_synthetic),
    `Std. difference` = sprintf("%.3f", standardized_difference)
  )

knitr::kable(
  sdid_balance_table,
  format = "latex",
  booktabs = TRUE,
  longtable = TRUE,
  caption = "Pre-treatment balance of the outcome and SDiD covariates.",
  align = "llrrrr"
) %>%
  kableExtra::kable_styling(
    font_size = 7
  ) %>%
  kableExtra::footnote(
    general = "Continuous covariates use the scale of the estimation data. The table is a diagnostic, not a claim that all covariates are perfectly balanced.",
    general_title = "Note: ",
    escape = FALSE,
    threeparttable = TRUE
  )
```

```{r sdid-placebo-rank, message=FALSE, warning=FALSE, echo=FALSE, results='asis'}
sdid_placebo_rank_table <- sdid_placebo_rank %>%
  dplyr::transmute(
    Estimand = estimand,
    `Brazil statistic` = sprintf("%.3f", brazil_statistic),
    Rank = as.integer(rank),
    Denominator = as.integer(denominator),
    `Rank p-value` = sprintf("%.3f", rank_based_p_value),
    Note = inference_note
  )

knitr::kable(
  sdid_placebo_rank_table,
  format = "latex",
  booktabs = TRUE,
  caption = "Rank-based placebo inference for the Brazil SDiD estimate.",
  align = "lrrrrl"
) %>%
  kableExtra::kable_styling(
    latex_options = c("HOLD_position", "scale_down"),
    font_size = 7
  ) %>%
  kableExtra::footnote(
    general = "The placebo distribution includes Brazil and valid placebo-in-space assignments.",
    general_title = "Note: ",
    escape = FALSE,
    threeparttable = TRUE
  )
```

```{r sdid-donor-sensitivity, message=FALSE, warning=FALSE, echo=FALSE, results='asis'}
sdid_donor_sensitivity_table <- sdid_donor_sensitivity %>%
  dplyr::transmute(
    Specification = specification,
    ATT = sprintf("%.3f", estimate),
    `Adj. pre RMSPE` = sprintf("%.3f", rmspe_pre),
    `Post/pre adj. RMSPE` = sprintf("%.3f", rmspe_ratio),
    `Removed donors` = as.integer(n_removed_donors),
    `% change` = sprintf("%.1f", percent_change_vs_baseline),
    Stability = substantive_stability
  )

knitr::kable(
  sdid_donor_sensitivity_table,
  format = "latex",
  booktabs = TRUE,
  longtable = TRUE,
  caption = "Sensitivity to influential donors and high-weight donor blocks.",
  align = "lrrrrrl"
) %>%
  kableExtra::kable_styling(
    font_size = 7
  ) %>%
  kableExtra::footnote(
    general = "Placebo SEs are not recomputed for donor-deletion rows; those rows are point-estimate and intercept-adjusted fit diagnostics.",
    general_title = "Note: ",
    escape = FALSE,
    threeparttable = TRUE
  )
```

```{r sdid-window-sensitivity, message=FALSE, warning=FALSE, echo=FALSE, results='asis'}
sdid_window_sensitivity_table <- sdid_window_sensitivity %>%
  dplyr::transmute(
    Window = specification,
    ATT = sprintf("%.3f", estimate),
    `Adj. pre RMSPE` = sprintf("%.3f", rmspe_pre),
    `Post/pre adj. RMSPE` = sprintf("%.3f", rmspe_ratio),
    `One-sided p` = sprintf("%.3f", placebo_p_one_sided_negative),
    `Two-sided p` = sprintf("%.3f", placebo_p_two_sided_abs),
    `% change` = sprintf("%.1f", percent_change_vs_baseline),
    Stability = stability
  )

knitr::kable(
  sdid_window_sensitivity_table,
  format = "latex",
  booktabs = TRUE,
  caption = "Sensitivity to plausible time windows around the 1997-2015 baseline.",
  align = "lrrrrrrl"
) %>%
  kableExtra::kable_styling(
    latex_options = c("HOLD_position", "scale_down"),
    font_size = 8
  ) %>%
  kableExtra::footnote(
    general = "One-sided p-values are rank-based placebo p-values for negative effects when reported; non-baseline window rows are point-estimate and intercept-adjusted fit diagnostics.",
    general_title = "Note: ",
    escape = FALSE,
    threeparttable = TRUE
  )
```

```{r sdid-high-weight-donor-exposure, message=FALSE, warning=FALSE, echo=FALSE, results='asis'}
sdid_donor_china_exposure_table <- sdid_donor_china_exposure %>%
  dplyr::filter(high_weight_donor) %>%
  dplyr::transmute(
    Rank = as.integer(weight_rank),
    ISO3 = iso3c,
    Donor = country_name,
    Weight = sprintf("%.3f", unit_weight),
    `China share pre` = sprintf("%.3f", mean_china_export_share_pre_treatment),
    `China share post` = sprintf("%.3f", mean_china_export_share_post_treatment),
    `Best post rank` = as.integer(min_china_rank_post_treatment),
    Flag = china_exposure_flag
  )

knitr::kable(
  sdid_donor_china_exposure_table,
  format = "latex",
  booktabs = TRUE,
  caption = "High-weight donor exposure to China's export expansion.",
  align = "rllrrrrl"
) %>%
  kableExtra::kable_styling(
    latex_options = c("HOLD_position", "scale_down"),
    font_size = 8
  ) %>%
  kableExtra::footnote(
    general = paste0(
      "High-weight donors are donors with weights at least twice the uniform donor weight. ",
      "Among them, ",
      sdid_high_weight_exposure_summary$n_high_weight_top_china_post,
      " were cases where China was the top export destination after 2009 and ",
      sdid_high_weight_exposure_summary$n_high_weight_top_two_china_post,
      " were cases where China ranked in the top two, accounting for ",
      sprintf("%.1f", 100 * sdid_high_weight_exposure_summary$weighted_share_high_weight_top_two_china_post),
      " percent of total SDiD weight."
    ),
    general_title = "Note: ",
    escape = FALSE,
    threeparttable = TRUE
  )
```

Below I also present the ten largest donor weights used in the main model as a visual summary.

```{r plot-weights, message=FALSE, warning=FALSE, echo=FALSE, fig.cap=paste0("Ten largest donor weights contributing to the synthetic control in the Brazilian SDiD model. ", caption_note(unit = "Synthetic-control unit weights, normalized over the full donor pool rather than only over the displayed donors", window = "pre-treatment period ending in 2008")), fig.pos="H"}
tar_read(plot_weights_coef)
```

We also fit a model for a donor pool of Latin American states (excluding Caribbean countries), which is substantively attractive because the donor countries are regionally closer to Brazil. This regional donor pool is a sensitivity check rather than the preferred specification: the main design seeks a pre-treatment synthetic counterfactual with adequate donor support and fit, while the Latin American pool is much smaller.


```{r plot-latam, message=FALSE, warning=FALSE, echo=FALSE, fig.cap=paste0("SDiD fit with a Latin American donor pool. ", caption_note(unit = "Absolute UNGA ideal-point distance to China", se = "placebo-based, reported in text", window = "years shown in the plot")), fig.pos="H"}
tar_read(plot_trend_latam)
b_est_main <- tar_read(synth_fit)[1]           # main donor-pool estimate
b_est_latam <- tar_read(synth_fit_latam)[1]          # LATAM donor-pool estimate
se_latam    <- tar_read(se_synth_latam)
```

The Latin America donor-pool estimate has the same sign as the main estimate, but it is less precise and substantially larger in magnitude: `r round(b_est_latam,2)` with a standard error of `r round(se_latam,2)`, compared with `r round(b_est_main,2)` in the main donor pool. I therefore treat it as evidence that the negative sign is not unique to the global donor pool, not as evidence that the two counterfactuals produce nearly identical magnitudes. The countries with the largest Latin American weights are Argentina, Guatemala, Paraguay, and Uruguay -- all but Guatemala are founding members of Mercosur.

```{r plot-weight-latam, message=FALSE, warning=FALSE, echo=FALSE, fig.cap=paste0("Synthetic-control donor weights for the Latin American SDiD specification. ", caption_note(unit = "Synthetic-control unit weights, summing to 1", window = "pre-treatment period ending in 2008")), fig.pos="H"}
tar_read(plot_weights_coef_latam)
```

## Cross-Country Robustness: Duration Thresholds

The main cross-country estimate uses a five-year durability threshold because the theory concerns status that persists long enough to become politically usable. Table \@ref(tab:cross-country-duration-table) reports the same goods-only IFE design under three- and seven-year thresholds, along with the clean single-entry robustness sample. The estimates remain negative across the restricted-risk-set rows, though precision varies with treated-unit support.

```{r cross-country-duration-table, message=FALSE, warning=FALSE, echo=FALSE, results='asis'}
duration_tbl <- m2_model_results %>%
  dplyr::filter(specification %in% c("risk_set_restricted", "clean_single_spell")) %>%
  dplyr::mutate(
    `Minimum duration` = as.character(min_duration_years),
    Specification = specification_label,
    `ATT` = fmt3(att),
    `SE` = fmt3(se),
    `95\\% CI` = paste0("[", fmt3(ci_lo), ", ", fmt3(ci_hi), "]"),
    `p-value` = vapply(p, format_p, character(1)),
    `Units (T/C)` = paste0(n_treated, "/", n_control),
    `Treated periods` = as.character(n_treated_country_years),
    `r*` = as.character(r_cv)
  ) %>%
  dplyr::transmute(
    `Minimum duration`,
    Specification,
    ATT,
    SE,
    `95\\% CI`,
    `p-value`,
    `Units (T/C)`,
    `Treated periods`,
    `r*`
  )

knitr::kable(
  duration_tbl,
  format = "latex",
  booktabs = TRUE,
  escape = FALSE,
  caption = "Duration-threshold robustness for goods-only China top export-destination status.",
  align = "llccccccc"
) %>%
  kableExtra::kable_styling(
    latex_options = c("HOLD_position", "scale_down"),
    font_size = 7
  ) %>%
  kableExtra::footnote(
    general = paste0(
      sub(
        "^Note: ",
        "",
        caption_note(
          unit = "ATT in absolute UNGA ideal-point distance to China; lower values indicate convergence toward China",
          se = paste("bootstrap with", m2_bootstrap_label),
          window = "goods-only ITPD-E rank metric"
        )
      ),
      " The restricted-risk-set rows retain clean pre-entry non-China-top years and qualifying treated years. ",
      "The clean single-entry rows additionally keep only treated countries with one observed qualifying China-top period."
    ),
    general_title = "Note: ",
    escape = FALSE,
    threeparttable = TRUE
  )
```

## Cross-Country Sample Audit: Goods-Only Treatment Coding

The revised rank metric excludes services before partner ranks are computed. Table \@ref(tab:cross-country-sector-table) documents the ITPD-E broad-sector filter. Table \@ref(tab:cross-country-audit-table) then summarizes how the five-year goods-only rule classifies countries in the UNGA/trade panel.

```{r cross-country-sector-table, message=FALSE, warning=FALSE, echo=FALSE, results='asis'}
sector_tbl <- m2_sector_audit %>%
  dplyr::mutate(
    `Included in rank metric` = dplyr::if_else(
      broad_sector %in% goods_sector_values,
      "Yes",
      "No"
    )
  ) %>%
  dplyr::transmute(
    `Broad sector` = broad_sector,
    `Raw rows` = raw_rows,
    `Positive-trade rows` = positive_trade_rows,
    `Included in rank metric`
  )

knitr::kable(
  sector_tbl,
  format = "latex",
  booktabs = TRUE,
  escape = FALSE,
  caption = "ITPD-E sector filter used to construct the goods-only China top export-destination metric.",
  align = "lrrl"
) %>%
  kableExtra::kable_styling(
    latex_options = c("HOLD_position"),
    font_size = 8
  )
```

```{r cross-country-audit-table, message=FALSE, warning=FALSE, echo=FALSE, results='asis'}
audit_tbl <- m2_country_audit %>%
  dplyr::filter(min_duration_years == 5L) %>%
  dplyr::distinct(
    iso3c, country_name, audit_role,
    n_observed_china_top_periods, n_qualifying_periods,
    n_short_eligible_periods, n_pre_min_entry_periods,
    n_no_clean_prior_periods
  ) %>%
  dplyr::count(audit_role, name = "countries") %>%
  dplyr::mutate(
    Role = dplyr::recode(
      audit_role,
      treated_qualifying = "Qualifying treated countries",
      never_china_top_control = "Never-China-top controls",
      excluded_short_duration = "Excluded: short China-top periods",
      excluded_pre_2000 = "Excluded: pre-2000 China-top entry",
      excluded_no_clean_prior = "Excluded: no clean observed prior non-China-top year",
      excluded_other_nonqualifying = "Excluded: other nonqualifying China-top cases",
      .default = audit_role
    )
  ) %>%
  dplyr::select(Role, Countries = countries)

knitr::kable(
  audit_tbl,
  format = "latex",
  booktabs = TRUE,
  escape = FALSE,
  caption = "Country-level sample audit for the five-year goods-only China top export-destination rule.",
  align = "lr"
) %>%
  kableExtra::kable_styling(
    latex_options = c("HOLD_position"),
    font_size = 8
  ) %>%
  kableExtra::footnote(
    general = "The audit uses the goods-only rank metric and classifies countries before the estimator-specific screen for minimum untreated support. Country counts are therefore a coding audit, not a replacement for the treated/control counts in the estimation table.",
    general_title = "Note: ",
    escape = FALSE,
    threeparttable = TRUE
  )
```

## Public Cue and Recoverability Audit

The cross-country media and source audit is used only to discipline the mechanism interpretation. It does not redefine the treatment. The treatment remains durable status in which China is the country's top goods-export destination; the audit asks whether public or official rank language is recoverable around entry and whether non-recovery should be read as meaningful silence. Table \@ref(tab:ex-top1-recoverability-audit) therefore separates observed China status cues, displaced-incumbent recoverability, and metric caveats. Negative cases are interpreted as recoverability diagnostics, not as proof that the status cue was absent.

```{r ex-top1-recoverability-audit, message=FALSE, warning=FALSE, echo=FALSE, results='asis'}
ex_top1_tables <- tar_read(ex_top1_salience_appendix_tables)
recoverability_tbl <- ex_top1_tables$recoverability_table

knitr::kable(
  recoverability_tbl,
  format = "latex",
  booktabs = TRUE,
  longtable = TRUE,
  escape = TRUE,
  caption = "Recoverability and metric caveats in the former-incumbent salience audit.",
  align = "lrllll"
) %>%
  kableExtra::kable_styling(
    font_size = 7
  ) %>%
  kableExtra::column_spec(
    5,
    width = "4.0cm"
  ) %>%
  kableExtra::column_spec(
    6,
    width = "4.2cm"
  ) %>%
  kableExtra::footnote(
    general = paste0(
      "The table is built from the supplemental former-incumbent audit tables. ",
      "The former-incumbent benchmark asks whether the export partner displaced by China was covered in recoverable local or official sources. ",
      "It is a recoverability diagnostic, not a rule for coding China salience as absent."
    ),
    general_title = "Note: ",
    escape = FALSE,
    threeparttable = TRUE
  )
```

```{r ex-top1-australia-afr-context, message=FALSE, warning=FALSE, echo=FALSE, results='asis'}
if (!exists("ex_top1_tables")) {
  ex_top1_tables <- tar_read(ex_top1_salience_appendix_tables)
}
afr_context_tbl <- ex_top1_tables$afr_context_table

knitr::kable(
  afr_context_tbl,
  format = "latex",
  booktabs = TRUE,
  escape = TRUE,
  caption = "Australian Financial Review context source for Australia.",
  align = "llllll"
) %>%
  kableExtra::kable_styling(
    latex_options = c("HOLD_position", "scale_down"),
    font_size = 7
  ) %>%
  kableExtra::footnote(
    general = "The AFR item documents a broad aggregate-trading-partner cue. It is retained as context for metric mismatch and prior salience, not as evidence of newspaper uptake of the strict export-destination treatment.",
    general_title = "Note: ",
    escape = FALSE,
    threeparttable = TRUE
  )
```

\clearpage

## ChatGPT Classification

Below is the coded instruction to the ChatGPT API to classify the subjects of the headlines.

```{r code-chatgpt, echo=TRUE, eval=FALSE}
system_block <- paste(
  'You are a Brazilian expert in International Relations and the political',
  'economy of China–Brazil relations.  Classify the subject of EACH headline',
  'when I say EACH, I mean all of them, even if repetitive or similar. Do classify **all** of them',
  'using **one** label from this list exactly as written:',
  '"diplomacy", "chinese economy", "china-brasil relations",',
  '"disaster and accidents", "chinese politics and policy",',
  '"sports, science, culture, other", "human rights",',
  '"china-brazil trade", "health".\n\n',
  # 13 static few-shot examples — leave them unchanged!
  'Examples:\n',
  '1. Presidente chinês visita o Brasil e participa de reunião com FHC -> china-brasil relations\n',
  '2. Ministro chinês de Defesa faz visita oficial ao Brasil -> china-brasil relations\n',
  '3. Para governo, novo câmbio chinês só beneficia Brasil no médio prazo -> china-brazil trade\n',
  '4. China ultrapassa EUA como maior parceiro comercial do Brasil -> china-brazil trade\n',
  '5. China vai controlar preços de alimentos e combustíveis -> chinese economy\n',
  '6. Vendas de automóveis na China desaceleram em junho -> chinese economy\n',
  '7. China quer devolução das relíquias Yin, patrimônio da Unesco -> sports, science, culture, other\n',
  '8. Dinossauro encontrado na China esclarece origem das aves -> sports, science, culture, other\n',
  '9. OMS confirma mais de 4.600 casos de gripe suína; China registra 2º caso -> health\n',
  '10. Marinha chinesa busca "cansar" barcos japoneses em águas disputadas -> diplomacy\n',
  '11. Acidente após casamento deixa 16 mortos na China -> disaster and accidents\n',
  '12. Jornal chinês sai das bancas por foto da praça Tiananmen -> human rights\n',
  '13. China é parceira estratégica contra crise econômica, diz UE -> chinese economy',
  'return as in the examples, a string with the headline and the classification'
)

```

The classified headline file is treated as an archived derived dataset rather than regenerated during manuscript rendering. This choice reflects the practical behavior of LLM-based coding: early runs sometimes returned incomplete classifications, altered original headlines, or used slightly inconsistent labels, which made exact matching back to the archive difficult. I therefore fixed the prompt, normalized labels, preserved the resulting classified file, and validate the coding below against an independently coded sample. The evidence in the main text is used as a headline-level salience diagnostic.

Below we provide a random sample of the headlines and the classification provided by ChatGPT.

```{r sample-chatgpt, message=FALSE, warning=FALSE, echo=FALSE, fig.cap=paste0("Random sample of China-related Folha de Sao Paulo headlines and ChatGPT classifications. ", caption_note(unit = "headline text and category labels", window = "sample drawn from the 2000-2014 corpus")), fig.pos="H", out.width="\\linewidth"}
knitr::include_graphics(tar_read(sample_headlines_image_file))
```

### Validation of ChatGPT Classification

To assess the reliability of the automated classification, one of the authors independently coded a stratified random sample of 100 headlines (with a minimum of 5 per category). Table \@ref(tab:chatgpt-validation) reports the agreement rate between the ChatGPT labels and the manual coding.

```{r chatgpt-validation, message=FALSE, warning=FALSE, echo=FALSE, results='asis'}
library(knitr)
library(dplyr)

validation_summary <- tar_read(chatgpt_validation_summary)
val_tbl <- validation_summary$table
overall_acc <- validation_summary$overall$accuracy_pct
validation_n <- validation_summary$overall$N
lowest_category <- validation_summary$lowest$category
lowest_acc <- validation_summary$lowest$accuracy_pct
category_acc <- validation_summary$category_accuracy
trade_acc <- category_acc$accuracy_pct[category_acc$category == "china-brazil trade"]
relations_acc <- category_acc$accuracy_pct[category_acc$category == "china-brazil relations"]
diplomacy_acc <- category_acc$accuracy_pct[category_acc$category == "diplomacy"]

kable(
  val_tbl,
  format = "latex",
  booktabs = TRUE,
  caption = paste0("Validation of ChatGPT classification against manual coding (N = ", validation_n, "). Overall accuracy: ", sprintf("%.1f", overall_acc), " percent."),
  align = "lccc"
) %>%
  kableExtra::kable_styling(
    latex_options = c("HOLD_position", "scale_down"),
    font_size = 9
  ) %>%
  kableExtra::footnote(
    general = sub(
      "^Note: ",
      "",
      caption_note(
        unit = "headline-level counts and accuracy (percent) by category",
        window = "stratified validation sample from the 2000-2014 headline corpus"
      )
    ),
    general_title = "Note: ",
    escape = FALSE,
    threeparttable = TRUE
  )
```

The overall accuracy is `r sprintf("%.1f", overall_acc)` percent. The lowest agreement is in the "`r lowest_category`" category (`r sprintf("%.1f", lowest_acc)` percent), where several headlines were manually reclassified as "diplomacy". The category most central to the mechanism, "china-brazil trade," achieved `r sprintf("%.1f", trade_acc)` percent agreement. Related bilateral categories had lower but still usable agreement: `r sprintf("%.1f", relations_acc)` percent for "china-brazil relations" and `r sprintf("%.1f", diplomacy_acc)` percent for "diplomacy."

## Cross-Country: Goods-Only Treated Countries

Table \@ref(tab:table-treated-appendix) describes the `r n_treated_abs` treated countries in the main goods-only restricted-risk-set specification. For each country, the table reports the first qualifying China-top goods-export year and the number of treated and untreated country-years retained in the estimation panel.

```{r table-treated-appendix, message=FALSE, warning=FALSE, echo=FALSE}
library(kableExtra)

treated_tbl <- m2_unit_summary %>%
  dplyr::filter(
    min_duration_years == 5L,
    sample == "risk_set_restricted",
    ever_treated
  ) %>%
  dplyr::arrange(first_treat, country_name) %>%
  dplyr::transmute(
    country = country_name,
    iso3c,
    first_treat_year = as.integer(first_treat),
    treated_years,
    untreated_years,
    first_year_in_panel,
    last_year_in_panel
  )

kable(treated_tbl,
      format = "latex",
      booktabs = TRUE,
      digits = 3,
      caption = "Treated countries in the main goods-only restricted-risk-set cross-country specification.",
      col.names = c("Country", "ISO3c", "Treatment year", "Treated years",
                     "Pre/other untreated years", "First panel year", "Last panel year"),
      align = "llccccc") %>%
  kable_styling(latex_options = c("hold_position", "scale_down"),
                font_size = 9) %>%
  footnote(general = paste0("Ranks are computed from goods exports only. ",
                            "The restricted risk set keeps clean pre-entry non-China-top years and qualifying treated years for treated countries. ",
                            "Treatment year is the first retained year of a qualifying China-top goods-export period."),
           general_title = "Note: ",
           escape = FALSE,
           threeparttable = TRUE)
```
